### 2012-10-17 proposal.tex
Proposal stub

\documentclass[11pt]{article}
\usepackage{amsfonts,amsmath,amssymb,amsthm}
\usepackage{array}
\usepackage{bm}
\usepackage{commath}
\usepackage{enumitem}
\usepackage{fancyhdr,fancyvrb}
\usepackage[letterpaper,text={7in,9in},centering]{geometry}
\usepackage[pdftex]{graphicx}
\usepackage{ifthen}
\usepackage{mathrsfs}
\usepackage{multicol}
\usepackage{pifont}
\usepackage{sectsty}
\usepackage{setspace}
\usepackage{stackrel}
\usepackage{stmaryrd}
\usepackage{tensor}
\usepackage{xspace}

\pagestyle{fancy}
\fancyhead[RO]{Max Wang \& Edward Gan}

\sectionfont{\large}
\subsectionfont{\normalsize}

\begin{document}

Insert proposal here.

\end{document}

### 2012-10-18 proposal.tex
Added basic document outline and bibtex placeholder
added bibliography to document
content vomit pt 1

@@
 \usepackage{xspace}
+\usepackage{natbib}
 
@@
 \begin{document}
+\bibliographystyle{plain}
 
-Insert proposal here.
+\title{Proposal: Loose Grammar Inference for Lossy Compression}
+\author{Edward Gan \& Max Wang}
+\date{10-17-12}
+\maketitle
+
+\begin{abstract}
+Test (\cite{sequitur})
+\end{abstract}
+
+\section{Introduction and Goals}
+
+\section{Background}
+
+Though we have not found very much previous work in the area of lossy grammar based compression,
+our proposed research will try to build upon powerful results in grammar based compression and
+incorporate some of ideas currently used in lossy compression algorithms.
+
+\subsection{Grammar Compression}
+
+As described in the introduction, the proposed research shares many of its goals with the sequitur
+algorithm described in \cite{sequitur}. The authors of \cite{sequitur} demonstrate how 
+context free grammars offer a means of simultaneously revealing the structure of
+data streams and compressing them effectively. They also develop an effective algorithm
+based on iteratively rewriting grammars to keep them small and efficient. Sequitur
+yields intriguing grammar based analyses of texts and musical scores, but an inspection of its
+results yields that it isn't able to detect and compress larger scale structure because
+of small fluctuations that occur in natural streams. This was part of our inspiration for
+trying to achieve both better compression and better structural analysis by ignoring
+small scale fluctuations.
+
+In \cite{sequitur2}, the detailed compression scheme used in the Sequitur algorithm is described.
+By sending the grammar over in an implicit, local-pointer based scheme, they are able to
+represent the grammar with very little overhead. A similar scheme should be easily adaptable to any
+grammar based compression scheme.
+
+Finally, within the realm of Sequitur based compression, \cite{nevillphd} explores a variety
+of ad-hoc extensions to the Sequitur algorithm which improve its compression performance
+on structured data. These ranged from introducing domain-specific constraints to their
+grammar, adding a few steps of backtracking to their normally greedy grammar formation,
+to guessing unifications in attempting to infer recursive grammar rules. Though these
+may invalidate some of their theoretical results on the asymptotic performance of
+Sequitur, in practice they seem to have worked and add on nicely into the grammar inference
+framework.
+
+The intution that grammar inference can support many detailed policies is made more formal
+in \cite{grammarcodes}. The authors classify the properties a grammar needs to function
+as a good compressor, and moreover give a set of reduction rules for putting grammar into
+the appropriate form. Within this context, grammar inference algorithms similar to Sequitur
+can be described as simple applications of their reduction rules to different ways
+of generating base grammars. Even more examples of the variety of grammar inference schemes
+possible under this general framework of reducing grammars can be found in 
+\cite{efficientgreedy}
+
+There has also been very interesting work analyzing how grammar based algorithms compare with
+the theoretical best lossless grammar for a string, in the line of \cite{approximate}. However,
+these analytic bounds for lossless grammars are outside the scope of our proposed research.
+
+In summary, established work on grammar inference algorithms provides a solid set of tools
+for experimenting with the kinds of local modifications we propose to make to improve
+grammar structure by introducing lossiness.
+
+\subsection{Lossy Compression}
+
+\section{Research Plan}
+
+The proposed research project will involve designing new lossy grammar compression
+algorithms, implementing them, and comparing their performance with existing methods.
+If time permits we would also like to prove lower bounds on the runtime of our
+algorithm and upper bounds on the size of the grammar it generates.
+
+\subsection{Algorithm Design}
+
+Starting with an existing grammar inference algorithm, or perhaps with a
+simple variation based
+on the reduction rules in \cite{grammarcodes}, there are 3 places we might introduce
+lossiness into the system. We could
+\begin{itemize}
+\item perform text-transforms on the input before applying a 
+      grammar inference algorithms
+\item modify the grammar inference algorithm so that it generates a lossy
+      grammar on the fly
+\item take a lossless grammar code and apply lossy transforms to it
+\end{itemize}
+
+Of these, the first could provide the biggest potential for taking advantage of
+source-specific knowledge of what kinds of data we can ignore. The second
+allows the most control over precisely how a grammar can be formed, and would not
+be too difficult to build into reduction-rule based frameworks. The third would
+be the most general, since it could be applied to any CFG, but may be difficult
+to get consistent performance from.
+
+After a lossy CFG is obtained, we plan on using standard methods for encoding the CFG
+such as those described in \cite{sequitur2}
+
+\subsection{Evaluation Methodology}
+
+\subsection{Implementation and Resources}
+
+\subsection{Possible Issues}
+
+\nocite{*}
+
+\bibliography{sources}
 

### 2012-10-18 compile.sh
added compilatio script

@@ -1,0 +1,6 @@
+#!/bin/bash
+
+pdflatex proposal
+bibtex proposal
+pdflatex proposal
+pdflatex proposal

### 2012-10-18 proposal.tex
added compilatio script
Reflow

@@
 
-\begin{abstract}
-Test (\cite{sequitur})
-\end{abstract}
-
 \section{Introduction and Goals}
@@
 
-Though we have not found very much previous work in the area of lossy grammar based compression,
-our proposed research will try to build upon powerful results in grammar based compression and
-incorporate some of ideas currently used in lossy compression algorithms.
+Though we have not found very much previous work in the area of lossy grammar
+based compression, our proposed research will try to build upon powerful
+results in grammar based compression and incorporate some of ideas currently
+used in lossy compression algorithms.
 
@@
 
-As described in the introduction, the proposed research shares many of its goals with the sequitur
-algorithm described in \cite{sequitur}. The authors of \cite{sequitur} demonstrate how 
-context free grammars offer a means of simultaneously revealing the structure of
-data streams and compressing them effectively. They also develop an effective algorithm
-based on iteratively rewriting grammars to keep them small and efficient. Sequitur
-yields intriguing grammar based analyses of texts and musical scores, but an inspection of its
-results yields that it isn't able to detect and compress larger scale structure because
-of small fluctuations that occur in natural streams. This was part of our inspiration for
-trying to achieve both better compression and better structural analysis by ignoring
-small scale fluctuations.
+As described in the introduction, the proposed research shares many of its
+goals with the sequitur algorithm described in \cite{sequitur}. The authors of
+\cite{sequitur} demonstrate how context free grammars offer a means of
+simultaneously revealing the structure of data streams and compressing them
+effectively. They also develop an effective algorithm based on iteratively
+rewriting grammars to keep them small and efficient. Sequitur yields intriguing
+grammar based analyses of texts and musical scores, but an inspection of its
+results yields that it isn't able to detect and compress larger scale structure
+because of small fluctuations that occur in natural streams. This was part of
+our inspiration for trying to achieve both better compression and better
+structural analysis by ignoring small scale fluctuations.
 
-In \cite{sequitur2}, the detailed compression scheme used in the Sequitur algorithm is described.
-By sending the grammar over in an implicit, local-pointer based scheme, they are able to
-represent the grammar with very little overhead. A similar scheme should be easily adaptable to any
-grammar based compression scheme.
+In \cite{sequitur2}, the detailed compression scheme used in the Sequitur
+algorithm is described.  By sending the grammar over in an implicit,
+local-pointer based scheme, they are able to represent the grammar with very
+little overhead. A similar scheme should be easily adaptable to any grammar
+based compression scheme.
 
-Finally, within the realm of Sequitur based compression, \cite{nevillphd} explores a variety
-of ad-hoc extensions to the Sequitur algorithm which improve its compression performance
-on structured data. These ranged from introducing domain-specific constraints to their
-grammar, adding a few steps of backtracking to their normally greedy grammar formation,
-to guessing unifications in attempting to infer recursive grammar rules. Though these
-may invalidate some of their theoretical results on the asymptotic performance of
-Sequitur, in practice they seem to have worked and add on nicely into the grammar inference
-framework.
+Finally, within the realm of Sequitur based compression, \cite{nevillphd}
+explores a variety of ad-hoc extensions to the Sequitur algorithm which improve
+its compression performance on structured data. These ranged from introducing
+domain-specific constraints to their grammar, adding a few steps of
+backtracking to their normally greedy grammar formation, to guessing
+unifications in attempting to infer recursive grammar rules. Though these may
+invalidate some of their theoretical results on the asymptotic performance of
+Sequitur, in practice they seem to have worked and add on nicely into the
+grammar inference framework.
 
-The intution that grammar inference can support many detailed policies is made more formal
-in \cite{grammarcodes}. The authors classify the properties a grammar needs to function
-as a good compressor, and moreover give a set of reduction rules for putting grammar into
-the appropriate form. Within this context, grammar inference algorithms similar to Sequitur
-can be described as simple applications of their reduction rules to different ways
-of generating base grammars. Even more examples of the variety of grammar inference schemes
+The intution that grammar inference can support many detailed policies is made
+more formal in \cite{grammarcodes}. The authors classify the properties a
+grammar needs to function as a good compressor, and moreover give a set of
+reduction rules for putting grammar into the appropriate form. Within this
+context, grammar inference algorithms similar to Sequitur can be described as
+simple applications of their reduction rules to different ways of generating
+base grammars. Even more examples of the variety of grammar inference schemes
 possible under this general framework of reducing grammars can be found in
@@
 
-There has also been very interesting work analyzing how grammar based algorithms compare with
-the theoretical best lossless grammar for a string, in the line of \cite{approximate}. However,
-these analytic bounds for lossless grammars are outside the scope of our proposed research.
+There has also been very interesting work analyzing how grammar based
+algorithms compare with the theoretical best lossless grammar for a string, in
+the line of \cite{approximation}. However, these analytic bounds for lossless
+grammars are outside the scope of our proposed research.
 
-In summary, established work on grammar inference algorithms provides a solid set of tools
-for experimenting with the kinds of local modifications we propose to make to improve
-grammar structure by introducing lossiness.
+In summary, established work on grammar inference algorithms provides a solid
+set of tools for experimenting with the kinds of local modifications we propose
+to make to improve grammar structure by introducing lossiness.
 
@@
 
-The proposed research project will involve designing new lossy grammar compression
-algorithms, implementing them, and comparing their performance with existing methods.
-If time permits we would also like to prove lower bounds on the runtime of our
-algorithm and upper bounds on the size of the grammar it generates.
+The proposed research project will involve designing new lossy grammar
+compression algorithms, implementing them, and comparing their performance with
+existing methods.  If time permits we would also like to prove lower bounds on
+the runtime of our algorithm and upper bounds on the size of the grammar it
+generates.
 
@@
 
-Starting with an existing grammar inference algorithm, or perhaps with a
-simple variation based
-on the reduction rules in \cite{grammarcodes}, there are 3 places we might introduce
-lossiness into the system. We could
+Starting with an existing grammar inference algorithm, or perhaps with a simple
+variation based on the reduction rules in \cite{grammarcodes}, there are 3
+places we might introduce lossiness into the system. We could
 \begin{itemize}
-\item perform text-transforms on the input before applying a 
-      grammar inference algorithms
+  \item perform text-transforms on the input before applying a grammar
+    inference algorithms
   \item modify the grammar inference algorithm so that it generates a lossy
@@
 source-specific knowledge of what kinds of data we can ignore. The second
-allows the most control over precisely how a grammar can be formed, and would not
-be too difficult to build into reduction-rule based frameworks. The third would
-be the most general, since it could be applied to any CFG, but may be difficult
-to get consistent performance from.
+allows the most control over precisely how a grammar can be formed, and would
+not be too difficult to build into reduction-rule based frameworks. The third
+would be the most general, since it could be applied to any CFG, but may be
+difficult to get consistent performance from.
 
-After a lossy CFG is obtained, we plan on using standard methods for encoding the CFG
-such as those described in \cite{sequitur2}
+After a lossy CFG is obtained, we plan on using standard methods for encoding
+the CFG such as those described in \cite{sequitur2}
 
 \subsection{Evaluation Methodology}
+
+There are two important metrics in evaluating our
 

### 2012-10-18 proposal.tex
s/[Ss]equitur/\\Sequitur

@@
 
+\newcommand{\Sequitur}{\textsc{Sequitur}\xspace}
+
 \begin{document}
@@
 As described in the introduction, the proposed research shares many of its
-goals with the sequitur algorithm described in \cite{sequitur}. The authors of
+goals with the \Sequitur algorithm described in \cite{sequitur}. The authors of
 \cite{sequitur} demonstrate how context free grammars offer a means of
@@
 effectively. They also develop an effective algorithm based on iteratively
-rewriting grammars to keep them small and efficient. Sequitur yields intriguing
-grammar based analyses of texts and musical scores, but an inspection of its
-results yields that it isn't able to detect and compress larger scale structure
-because of small fluctuations that occur in natural streams. This was part of
-our inspiration for trying to achieve both better compression and better
-structural analysis by ignoring small scale fluctuations.
+rewriting grammars to keep them small and efficient. \Sequitur yields
+intriguing grammar based analyses of texts and musical scores, but an
+inspection of its results yields that it isn't able to detect and compress
+larger scale structure because of small fluctuations that occur in natural
+streams. This was part of our inspiration for trying to achieve both better
+compression and better structural analysis by ignoring small scale
+fluctuations.
 
-In \cite{sequitur2}, the detailed compression scheme used in the Sequitur
+In \cite{sequitur2}, the detailed compression scheme used in the \Sequitur
 algorithm is described.  By sending the grammar over in an implicit,
@@
 
-Finally, within the realm of Sequitur based compression, \cite{nevillphd}
-explores a variety of ad-hoc extensions to the Sequitur algorithm which improve
+Finally, within the realm of \Sequitur based compression, \cite{nevillphd}
+explores a variety of ad-hoc extensions to the \Sequitur algorithm which improve
 its compression performance on structured data. These ranged from introducing
@@
 invalidate some of their theoretical results on the asymptotic performance of
-Sequitur, in practice they seem to have worked and add on nicely into the
+\Sequitur, in practice they seem to have worked and add on nicely into the
 grammar inference framework.
@@
 reduction rules for putting grammar into the appropriate form. Within this
-context, grammar inference algorithms similar to Sequitur can be described as
+context, grammar inference algorithms similar to \Sequitur can be described as
 simple applications of their reduction rules to different ways of generating

### 2012-10-18 proposal.tex
Date
Proofreading fixes

@@
 \pagestyle{fancy}
-\fancyhead[RO]{Max Wang \& Edward Gan}
+\fancyhead[RO]{Edward Gan \& Max Wang}
 
@@
 \author{Edward Gan \& Max Wang}
-\date{10-17-12}
+\date{October 17, 2012}
 \maketitle
@@
 As described in the introduction, the proposed research shares many of its
-goals with the \Sequitur algorithm described in \cite{sequitur}. The authors of
-\cite{sequitur} demonstrate how context free grammars offer a means of
+goals with the \Sequitur algorithm described in \cite{sequitur}.  The authors
+of \cite{sequitur} demonstrate how context free grammars offer a means of
 simultaneously revealing the structure of data streams and compressing them
-effectively. They also develop an effective algorithm based on iteratively
+efficiently.  They also develop an effective algorithm based on iteratively
 rewriting grammars to keep them small and efficient. \Sequitur yields
-intriguing grammar based analyses of texts and musical scores, but an
-inspection of its results yields that it isn't able to detect and compress
-larger scale structure because of small fluctuations that occur in natural
-streams. This was part of our inspiration for trying to achieve both better
-compression and better structural analysis by ignoring small scale
-fluctuations.
+intriguing grammar based analyses of texts and musical scores, inspection of
+its results finds that it unable to detect and compress larger scale structure
+because of small fluctuations that occur in natural streams.  This was part of
+our inspiration for trying to achieve both better compression and better
+structural analysis by ignoring small scale fluctuations.
 
@@
 invalidate some of their theoretical results on the asymptotic performance of
-\Sequitur, in practice they seem to have worked and add on nicely into the
-grammar inference framework.
+\Sequitur, in practice they seem to have worked and add nicely to the grammar
+inference framework.
 
@@
 grammar needs to function as a good compressor, and moreover give a set of
-reduction rules for putting grammar into the appropriate form. Within this
+reduction rules for putting a grammar into the appropriate form.  Within this
 context, grammar inference algorithms similar to \Sequitur can be described as
@@
 Starting with an existing grammar inference algorithm, or perhaps with a simple
-variation based on the reduction rules in \cite{grammarcodes}, there are 3
+variation based on the reduction rules in \cite{grammarcodes}, there are three
 places we might introduce lossiness into the system.  We could
@@
   \item perform text-transforms on the input before applying a grammar
-    inference algorithms
+    inference algorithms;
   \item modify the grammar inference algorithm so that it generates a lossy
-    grammar on the fly
-  \item take a lossless grammar code and apply lossy transforms to it
+    grammar on the fly; or
+  \item take a lossless grammar code and apply lossy transforms to it.
 \end{itemize}
-
 Of these, the first could provide the biggest potential for taking advantage of
@@
 allows the most control over precisely how a grammar can be formed, and would
-not be too difficult to build into reduction-rule based frameworks. The third
-would be the most general, since it could be applied to any CFG, but may be
-difficult to get consistent performance from.
+not be too difficult, we expect, to build into reduction-rule based frameworks.
+The third would be the most general, since it could be applied to any CFG, but
+may be difficult to get consistent performance from.
 

### 2012-10-18 proposal.tex
began evaluation section

--- 2012-10-18 proposal.tex
@@
 
-There are two important metrics in evaluating our
+There are two important metrics we plan on evaluating our algorithm with. First,
+the lossiness it introduces should allow it to compress very well. This can be
+quantified by comparing its compression ratio with other grammar based and 
+general compression schemes on text sources, especially semi-structured text 
+which might best exhibit nearly hierarchical data.
+
+We would also like to make use of our algorithm to reveal the inherent 
+structure in text data. Our hope is that allowing lossiness will alow it 
+to capture larger scale structures, so we plan on subjectively evaluating 
+the CFGs produced for how well they group the components that a human
+might perceive in pieces of text.
+
+Comparing our runtime with existing algorithms can be done in a
+straightforward way.
+
+We plan on evaluating our algorithm on standard english prose texts, but
+we also suspect that the texts which will prove most interesting to 
+analyze are those with slightly more structure imposed on them. Human-made
+tables such as the genealogical trees discussed in \cite{nevillphd} are a
+good example, as are simple musical pieces. Slightly more artificial
+examples we might try include strings generated from actual CFGs but with
+small amounts of noise added in.
 

### 2012-10-18 proposal.tex
Moar reflow
Starting intro

@@
 
+Many sources of data exhibit natural hierarchical structure, but it is
+oftentimes difficult to apprehend this structure at a high level due to small
+fluctuations in the data at the lower levels of the hierarchy.  If one is
+interested primarily in the structure of the data, a lossy compression scheme
+has the potential to mitigate some of this low-level noise in favor of the
+high-level signal, and can reduce the data size to boot.
+
 \section{Background}
@@
 Finally, within the realm of \Sequitur based compression, \cite{nevillphd}
-explores a variety of ad-hoc extensions to the \Sequitur algorithm which improve
-its compression performance on structured data.  These ranged from introducing
-domain-specific constraints to their grammar, adding a few steps of
+explores a variety of ad-hoc extensions to the \Sequitur algorithm which
+improve its compression performance on structured data.  These ranged from
+introducing domain-specific constraints to their grammar, adding a few steps of
 backtracking to their normally greedy grammar formation, to guessing
@@
 
-There are two important metrics we plan on evaluating our algorithm with. First,
-the lossiness it introduces should allow it to compress very well. This can be
-quantified by comparing its compression ratio with other grammar based and 
-general compression schemes on text sources, especially semi-structured text 
-which might best exhibit nearly hierarchical data.
+There are two important metrics we plan on evaluating our algorithm with.
+First, the lossiness it introduces should allow it to compress very well. This
+can be quantified by comparing its compression ratio with other grammar based
+and general compression schemes on text sources, especially semi-structured
+text which might best exhibit nearly hierarchical data.
 
 We would also like to make use of our algorithm to reveal the inherent
-structure in text data. Our hope is that allowing lossiness will alow it 
-to capture larger scale structures, so we plan on subjectively evaluating 
-the CFGs produced for how well they group the components that a human
-might perceive in pieces of text.
+structure in text data. Our hope is that allowing lossiness will alow it to
+capture larger scale structures, so we plan on subjectively evaluating the CFGs
+produced for how well they group the components that a human might perceive in
+pieces of text.
 
-Comparing our runtime with existing algorithms can be done in a
-straightforward way.
+Comparing our runtime with existing algorithms can be done in a straightforward
+way.
 
-We plan on evaluating our algorithm on standard english prose texts, but
-we also suspect that the texts which will prove most interesting to 
-analyze are those with slightly more structure imposed on them. Human-made
-tables such as the genealogical trees discussed in \cite{nevillphd} are a
-good example, as are simple musical pieces. Slightly more artificial
-examples we might try include strings generated from actual CFGs but with
-small amounts of noise added in.
+We plan on evaluating our algorithm on standard english prose texts, but we
+also suspect that the texts which will prove most interesting to analyze are
+those with slightly more structure imposed on them. Human-made tables such as
+the genealogical trees discussed in \cite{nevillphd} are a good example, as are
+simple musical pieces. Slightly more artificial examples we might try include
+strings generated from actual CFGs but with small amounts of noise added in.
 

### 2012-10-18 proposal.tex
Moar intro

@@
 high-level signal, and can reduce the data size to boot.
+
+One method of exploring, and then compressing, the hierarchical structure of
+data is to transform the data into a grammar.  The \Sequitur algorithm of
+\cite{sequitur} does exactly this, demonstrating how context-free grammars in
+particular offer a means of simultaneously revealing the structure of data
+inputs and compressing it efficiently.  However, inspection of \Sequitur's
+results reveals that it is unable to detect and to compress larger scale
+structure because of the aforementioned low-level fluctuations.
+
+We are interested in whether lossy compression techniques can be combined with
+CFG transformation and compression techniques like \Sequitur's in order (a) to
+illuminate the high-level grammatical structure of the data, and (b) to more
+effectively compress the data with some tolerable degree of lossiness.  That
+is, our goals are twofold---we wish to develop a technique which can both
+increase human understanding of a complex document and yield good compression
+ratios.  Notably, if we are interested in the structure rather than the
+details, we surmise that the lossiness we introduce can be highly desirable.
+For instance, introducing lossiness to \Sequitur's transformation on musical
+passages could result in interesting, even music-theoretically sound
+reductions, of a given piece.
 

### 2012-10-18 proposal.tex
~Finishing intro

@@
 reductions, of a given piece.
+
+The scope of our focus will be like that of \Sequitur: we will consider the
+composite application of both grammar transformation and grammar compression to
+an input.
+
+There seems to be little existing work done in the intersection of grammar
+compression and lossy compression.  Thus, we will compare our work to existing
+techniques both in lossless hierarchical compression and lossy non-hierarchical
+compression.  Moreover, we will not concern ourselves with fitting our lossy
+grammar compression into overarching information-theoretic developments.
 

### 2012-10-18 proposal.tex
Crunching first paragraphs of Background

@@
 goals with the \Sequitur algorithm described in \cite{sequitur}.  The authors
-of \cite{sequitur} demonstrate how context free grammars offer a means of
-simultaneously revealing the structure of data streams and compressing them
-efficiently.  They also develop an effective algorithm based on iteratively
-rewriting grammars to keep them small and efficient. \Sequitur yields
-intriguing grammar based analyses of texts and musical scores, inspection of
-its results finds that it unable to detect and compress larger scale structure
-because of small fluctuations that occur in natural streams.  This was part of
-our inspiration for trying to achieve both better compression and better
-structural analysis by ignoring small scale fluctuations.
-
-In \cite{sequitur2}, the detailed compression scheme used in the \Sequitur
-algorithm is described.  By sending the grammar over in an implicit,
-local-pointer based scheme, they are able to represent the grammar with very
-little overhead.  A similar scheme should be easily adaptable to any grammar
-based compression scheme.
+of \cite{sequitur} develop an effective algorithm based on iteratively
+rewriting grammars to keep them small and efficient, and produce interesting
+grammar-based analyses of texts and musical scores.  In \cite{sequitur2}, the
+detailed compression scheme used in the \Sequitur algorithm is described.  By
+sending the grammar over in an implicit, local-pointer based scheme, they are
+able to represent the grammar with very little overhead.  A similar scheme
+should be easily adaptable to any grammar based compression scheme.
 

### 2012-10-18 proposal.tex
Slight revising of intro

@@
 
-The scope of our focus will be like that of \Sequitur: we will consider the
-composite application of both grammar transformation and grammar compression to
-an input.
+As with \Sequitur, the scope of our problem will encompass the full conversion
+pathway of the input data---both the grammar transformation and the grammar
+compression steps---and we will consider introducing lossiness at each point in
+the process.
 

### 2012-10-18 proposal.tex
edits, and a few more transition sentences

@@
 Many sources of data exhibit natural hierarchical structure, but it is
-oftentimes difficult to apprehend this structure at a high level due to small
-fluctuations in the data at the lower levels of the hierarchy.  If one is
+often difficult to apprehend this structure at a high level due to small
+fluctuations at the lower levels of the hierarchy.  If one is
 interested primarily in the structure of the data, a lossy compression scheme
 has the potential to mitigate some of this low-level noise in favor of the
-high-level signal, and can reduce the data size to boot.
+high-level signal, and can greatly reduce the compressed data size.
 
@@
 For instance, introducing lossiness to \Sequitur's transformation on musical
-passages could result in interesting, even music-theoretically sound
-reductions, of a given piece.
+passages could result in interesting, even music-theoretically sound, 
+reductions of a given piece.
 
@@
 techniques both in lossless hierarchical compression and lossy non-hierarchical
-compression.  Moreover, we will not concern ourselves with fitting our lossy
-grammar compression into overarching information-theoretic developments.
+compression.  Moreover, we are not too concerned with fitting our lossy
+grammar compression into information theoretical developments, since
+our main concerns are achieving interesting human comprehensible compressions
+with reasonable fidelity and good ratios as a benchmark.
 
@@
 Though we have not found very much previous work in the area of lossy grammar
-based compression, our proposed research will try to build upon powerful
+based compression, our proposed research will try to build upon established
 results in grammar based compression and incorporate some of ideas currently
@@
 \begin{itemize}
-  \item perform text-transforms on the input before applying a grammar
+  \item perform text-transforms on the input before applying grammar
     inference algorithms;
@@
 Comparing our runtime with existing algorithms can be done in a straightforward
-way.
+way. Though our main focus will be on smaller datasets with human-readable
+structures, we will also experiment to see the effects of lossiness on
+scalability.
 
@@
 strings generated from actual CFGs but with small amounts of noise added in.
+These will reveal best case behavior in compressing a nearly-perfectly
+structured dataset.
 

### 2012-10-18 proposal.tex
Authoring lossy compression background and resources sections

@@
 
+While we were unable to find work in the area of lossy grammar compression, a
+number of existing techniques for lossy hierarchical compression more broadly
+may be applicable.  Kieffer notes that lossless hierarchical compression
+methods, including context-free grammars, can be used interchangeably to
+equivalently represent the structure of a given input \cite{tutorial}.
+However, we found no such claims for lossy hierarchical compression methods.
+
+Kieffer briefly identifies two primary methods for lossy hierarchical
+compression: wavelet-based and fractal-based schemes \cite{tutorial}.  In
+wavelet-based schemas, a signal is compressed by recording the largest
+coefficients of its wavelet decomposition, given a suitable wavelet function.
+Clustered coefficients can be stored in a compact form and used to encode
+hierarchical structure.  The JPEG image format uses a similar technique in
+order to compress images lossily \cite{jpeg}.
+
+In fractal compression, we wish to determine a set of contraction maps $w_i$
+over the domain of the data such that the fixed point of the $W(B) = \bigcup
+w_i(B)$ is the data set itself.  In this way, this fractal transform captures
+the self-similarity of an input by equating it with a covering of contractions
+of itself.  By restricting $w_i$ in, say, image compression, to affine
+transformations, we can easily represent each contraction map by a small
+sequence of coefficients.  Barnsley describes how to ``home in'' on the right
+coefficients to approximately represent a given input.  Interestingly, fractal
+compression on images can sometimes produce results which appear to be of much
+higher resolution than the original, which suggests that the technique provides
+some great measure of insight into the data's structure.
+
+Witten, \emph{et.\ al.}\ apply fractal compression to the area of text
+compression in order to generate texts which approximate a given input.  This
+yields a extraordinarily high compression ratio at the expense of time.  The
+authors do not supply their algorithm for fractal-based generation, but suggest
+that their particular algorithm may only be applicable to a certain style of
+prosaic text.
+
+Work has also been done regarding the lossy compression of particular input
+formats.  For text, methods such as replacing words with synonyms
+\cite{semanticlossy} or reordering the middle letters of a word (i.e., all
+letters save for the first and last) \cite{semilossless} have been explored.
+Applying such methods before transforming text into a grammar might yield
+interesting structural results, although we suspect that, as non-hierarchical
+schemes, they may not.
+
+We will try to adapt this variety of lossy compression techniques to function
+at the different points in the process of converting data into the form of a
+compressed grammar.
+
 \section{Research Plan}
@@
 
-\subsection{Implementation and Resources}
+\subsection{Resources}
+
+The primary resources we will require is source code or a functioning binary
+for the existing lossy compression, grammar transformation, and grammar
+compression algorithms we wish to use for our overall lossy grammar compression
+scheme.
 

### 2012-12-02 src/cfg.rb
Pass next unallocated nonterm symbol when making CFGs

@@ -9,7 +9,8 @@ class CFG
 
   attr_reader :rules
 
-  def initialize(start='~[*]', rules={})
+  def initialize(nonterm, start='~[*]', rules={})
+    @nonterm = nonterm
     @start = start
     @rules = rules
     @cache = {}

### 2012-12-02 src/sequitur.rb

@@ -13,8 +13,7 @@ class Sequitur
   end
 
   def gen_nonterm
-    @nonterm = @nonterm.succ
-    '~[' + @nonterm + ']'
+    '~[' + @nonterm.succ! + ']'
   end
 
   def run
@@ -151,7 +150,7 @@ class Sequitur
     rules = [rule]
     nonterminals = { rule.token => true }
 
-    rules.inject(CFG.new) do |cfg, rule|
+    rules.inject(CFG.new(@nonterm)) do |cfg, rule|
       cfg[rule.token] = rule.inject(List.new) do |list, node|
         symbol = node.value
 

### 2012-12-02 src/util.rb
Util file for lcs, among other things

@@ -1,0 +1,41 @@
+#
+# util.rb - Everybody loves util.
+#
+
+#
+# Longest common substring implementation taken from
+# http://rosettacode.org/wiki/Longest_Common_Subsequence#Ruby
+#
+def lcs(a, b)
+  lengths = Array.new(a.size + 1) { Array.new(b.size + 1) { 0 } }
+
+  # Row 0 and column 0 are initialized to 0 already.
+  a.split('').each_with_index do |x, i|
+    b.split('').each_with_index do |y, j|
+      if x == y
+        lengths[i+1][j+1] = lengths[i][j] + 1
+      else
+        lengths[i+1][j+1] = [lengths[i+1][j], lengths[i][j+1]].max
+      end
+    end
+  end
+
+  # Read the substring out from the matrix.
+  result = ""
+  x, y = a.size, b.size
+
+  while x != 0 and y != 0
+    if lengths[x][y] == lengths[x-1][y]
+      x -= 1
+    elsif lengths[x][y] == lengths[x][y-1]
+      y -= 1
+    else
+      # Assert a[x-1] == b[y-1]
+      result << a[x-1]
+      x -= 1
+      y -= 1
+    end
+  end
+
+  result.reverse
+end

### 2012-12-02 src/cfg.rb
Fully implemented, untested, Reducer class

@@ -43,7 +43,32 @@ class CFG
       rules_copy[lhs] = rhs.copy
       rules_copy
     end
-    CFG.new @start, rules_copy
+    CFG.new @nonterm, @start, rules_copy
+  end
+
+  #
+  # Add a string as a new rule.
+  #
+  def add_rule(str)
+    nonterm = '~[' + @nonterm.succ! + ']'
+
+    # XXX: Dirty hack.
+    @rules[nonterm] = str.split('~').inject(List.new) do |list, s|
+      list << s[0].chr == '[' && s[-1].chr == ']' ? '~' + seq : seq
+    end
+
+    nonterm
+  end
+
+  #
+  # Iterate over all RHS nonterminal nodes.  If a symbol is provided as
+  # argument, iterates only over matching nonterminals.
+  #
+  def each_nonterm(nonterm=nil)
+    @rules.each do |_, rhs|
+      nodes = rhs.select { |node| nonterm.nil? || node.value == nonterm }
+      nodes.each { |node| yield node }
+    end
   end
 
   #
@@ -69,17 +94,6 @@ class CFG
   def expand_all
     expand
     @cache
-  end
-
-  #
-  # Iterate over all RHS nonterminal symbols.  If a symbol is provided as
-  # argument, iterates only over matching nonterminals.
-  #
-  def each_nonterm(nonterm=nil)
-    @rules.each do |_, rhs|
-      nodes = rhs.select { |node| nonterm.nil? || node.value == nonterm }
-      nodes.each { |node| yield node }
-    end
   end
 
   #
@@ -115,6 +129,48 @@ class CFG
 
   def inline(nonterm)
     copy.inline! nonterm
+  end
+
+  #
+  # Factor out symbol-for-symbol occurrences of the RHS of nonterm from the
+  # target rule.
+  #
+  def factor!(target, nonterm)
+    rhs = @rules[target]
+    seq = @rules[nonterm].join
+
+    # Repeatedly factor out the first occurrence of seq until we're done.
+    until (i = rhs.join.index seq).nil?
+      pos = 0
+      rhs = rhs.inject(List.new) do |list, node|
+        val = node.value
+
+        rel = i - pos
+        if rel >= 0 and rel < val.size
+          # The start of seq is in this node.
+          prefix = val[0...rel]
+          list << prefix if prefix.size > 0
+          list << nonterm
+        end
+
+        rel = i + size - pos
+        if rel >= 0 and rel < val.size
+          # The end of seq is in this node.
[…]

### 2012-12-02 src/main.rb

@@ -10,6 +10,7 @@ require 'ostruct'
 require_relative 'cfg.rb'
 require_relative 'list.rb'
 require_relative 'lossify.rb'
+require_relative 'reduce.rb'
 require_relative 'sequitur.rb'
 require_relative 'graphplot.rb'
 

### 2012-12-02 src/reduce.rb

@@ -3,6 +3,7 @@
 #
 
 require_relative 'cfg.rb'
+require_relative 'util.rb'
 
 class Reducer
   attr_reader :cfg
@@ -12,52 +13,120 @@ class Reducer
   end
 
   def run
+    while [
+      eliminate_singletons,
+      unify_internal,
+      unify_pairwise,
+      apply_rules,
+      eliminate_duplicates
+    ].any?; end
   end
 
   private
 
-  def stringify(list)
-    list.inject('') { |s, node| s + node.value }
-  end
-
   #
-  # Eliminate any rules which are used only once in the CFG.
+  # R1: Eliminate any rules which are used only once in the CFG.
   #
   def eliminate_singletons
     @cfg.count.inject(false) do |found, (nonterm, count)|
-      next (found || false) if count != 1
+      next found if count != 1
       @cfg.inline! nonterm
-      found || true
+      true
     end
   end
 
+  class Retry < StandardError; end
+
   #
-  # Unify pairs of substrings (of length at least 2) within a single rule by
-  # making a new rule for them.
+  # R2: Unify pairs of substrings (of length at least 2) within a single rule
+  # by making a new rule for them.
   #
   def unify_internal
     @cfg.rules.each do |(lhs, rhs)|
+      seq = rhs.join[/(.*)(.*)\1/, 1]
+      next if seq.nil?
+
+      nonterm = @cfg.add_rule(seq)
+      @cfg.factor! lhs, nonterm
     end
   end
 
   #
-  # Unify pairs of substrings (of length at least 2) between two different
+  # Loop through pairs of stringified rules, yielding the shorter rule first.
+  #
+  def each_rule_pair
+    rules = Hash[@cfg.rules.map { |lhs, rhs| [lhs, rhs.join('')] }]
+
+    rules.each_index { |rule1, i| rules.each_index { |rule2, j|
+      next unless j < i
+      yield [rule1, rule2].sort_by { |(_, rhs)| rhs.size }
+    } }
+  end
+
+  #
+  # General method for finding substring matches between pairs of RHS
+  # expansions and then reducing.
+  #
+  # @param match    Function which takes two RHS's, returning truthy data on
+  #                 match or falsey on no-match.
+  # @param reduce   Function which takes two LHS's and reduces @cfg.
[…]

### 2012-12-02 src/cfg.rb
Fixing lots of Reducer bugs

@@ -47,15 +47,20 @@ class CFG
   end
 
   #
+  # Returns @rules with RHS's stringified.
+  #
+  def rules_s
+    Hash[rules.map { |lhs, rhs| [lhs, rhs.join('')] }]
+  end
+
+  #
   # Add a string as a new rule.
   #
   def add_rule(str)
     nonterm = '~[' + @nonterm.succ! + ']'
 
-    # XXX: Dirty hack.
-    @rules[nonterm] = str.split('~').inject(List.new) do |list, s|
-      list << s[0].chr == '[' && s[-1].chr == ']' ? '~' + seq : seq
-    end
+    rhs = str.split(/(~\[\w*\])/).select { |s| !s.empty? }
+    @rules[nonterm] = rhs.inject(List.new) { |list, s| list << s }
 
     nonterm
   end
@@ -137,13 +142,14 @@ class CFG
   #
   def factor!(target, nonterm)
     rhs = @rules[target]
-    seq = @rules[nonterm].join
+    seq = @rules[nonterm].join('')
 
     # Repeatedly factor out the first occurrence of seq until we're done.
-    until (i = rhs.join.index seq).nil?
+    until (i = rhs.join('').index seq).nil?
       pos = 0
       rhs = rhs.inject(List.new) do |list, node|
         val = node.value
+        found = false
 
         rel = i - pos
         if rel >= 0 and rel < val.size
@@ -151,17 +157,23 @@ class CFG
           prefix = val[0...rel]
           list << prefix if prefix.size > 0
           list << nonterm
+
+          found = true
         end
 
-        rel = i + size - pos
+        rel = i + seq.size - pos
         if rel >= 0 and rel < val.size
           # The end of seq is in this node.
           suffix = val[rel..-1]
           list << suffix if suffix.size > 0
+
+          found = true
         end
 
         # Increment position pointer.
         pos += val.size
+
+        found ? list : list << node.value
       end
     end
 

### 2012-12-02 src/lossify.rb

@@ -31,7 +31,7 @@ module Lossifier
       1.upto(Float::INFINITY) do |i|
         find, repl = eliminate
         break if find.nil? or repl.nil?
-        puts ">   [#{i.to_s}]".ljust(10) + find + ':' + repl if @verbose
+        puts ">   [#{i}]".ljust(10) + find + ':' + repl if @verbose
       end
       puts "\n" if @verbose
 

### 2012-12-02 src/main.rb

@@ -24,8 +24,9 @@ ALGORITHMS = {
 options = OpenStruct.new
 options.algorithms = ALGORITHMS.values
 options.expand = true
-options.print_grammar = false
-options.plot_grammar = false
+options.print = false
+options.plot = false
+options.reduce = false
 options.verbose = false
 
 OptionParser.new do |opts|
@@ -36,11 +37,11 @@ OptionParser.new do |opts|
   end
 
   opts.on("-g", "--[no-]print-grammar", "Print grammar") do |g|
-    options.print_grammar = g
+    options.print = g
   end
 
   opts.on("-p", "--[no-]plot-grammar", "Plot grammar") do |p|
-    options.plot_grammar = p
+    options.plot = p
   end
 
   opts.on("-l", "--lossifiers [ALGO1, ALGO2, ...]", Array,
@@ -52,6 +53,10 @@ OptionParser.new do |opts|
 
     raise OptionParser::InvalidArgument unless (algos - ALGORITHMS.keys).empty?
     options.algorithms = algos.uniq.map { |algo| ALGORITHMS[algo] }
+  end
+
+  opts.on("-r", "--[no-]reduce", "Apply grammar reductions") do |r|
+    options.reduce = r
   end
 
   opts.on("-v", "--verbose", "Run verbosely") do |v|
@@ -69,24 +74,27 @@ if ARGV[0].nil?
   exit
 end
 
-fname = ARGV[0]
-ftitle = fname[0..-5]
-str = File.read(fname)
-cfg = Sequitur.new(str).run
-
 def puts_cfg(cfg, options)
-  puts cfg if options.print_grammar
+  puts cfg if options.print
   puts cfg.expand if options.expand
   puts "\n"
 end
 
+str = File.read(ARGV[0])
+fprefix = ARGV[0].rpartition('.').first
+
+cfg = Sequitur.new(str).run
+cfg = Reducer.new(cfg).run if options.reduce
+
 puts "Sequitur-----------------------------------------------\n\n"
 puts_cfg cfg, options
-outputCFG(cfg,ftitle+"-orig.png") if options.plot_grammar
+outputCFG(cfg, fprefix + "-orig.png") if options.plot
 
 options.algorithms.each do |algo|
   puts algo.name + '-' * (55 - algo.name.size) + "\n\n"
-  newcfg = algo.new(cfg, options.verbose).run
-  puts_cfg(newcfg, options)
[…]

### 2012-12-02 src/reduce.rb

@@ -8,27 +8,36 @@ require_relative 'util.rb'
 class Reducer
   attr_reader :cfg
 
-  def initialize(cfg)
+  def initialize(cfg, verbose=false)
     @cfg = cfg
+    @verbose = verbose
   end
 
   def run
-    while [
-      eliminate_singletons,
-      unify_internal,
-      unify_pairwise,
-      apply_rules,
-      eliminate_duplicates
-    ].any?; end
+    i = 0
+    puts "> Reduction:" if @verbose
+
+    while (res = [ eliminate_singletons,
+                   unify_internal,
+                   unify_pairwise,
+                   apply_rules,
+                   eliminate_duplicates]).any?
+      puts ">   [#{i}]".ljust(10) + res.map { |e| e ? 1 : 0 }.join(', ')
+    end
+    @cfg
   end
 
   private
+
+  def nsymbs(s)
+    s.gsub(/~\[\w*\]/, '~').size
+  end
 
   #
   # R1: Eliminate any rules which are used only once in the CFG.
   #
   def eliminate_singletons
-    @cfg.count.inject(false) do |found, (nonterm, count)|
+    @cfg.counts.inject(false) do |found, (nonterm, count)|
       next found if count != 1
       @cfg.inline! nonterm
       true
@@ -42,9 +51,9 @@ class Reducer
   # by making a new rule for them.
   #
   def unify_internal
-    @cfg.rules.each do |(lhs, rhs)|
-      seq = rhs.join[/(.*)(.*)\1/, 1]
-      next if seq.nil?
+    @cfg.rules_s.each do |(lhs, rhs)|
+      seq = rhs[/((?>~\[.*\]|.)*)((?>~\[.*\]|.)*)\1/, 1]
+      next if seq.nil? or nsymbs(seq) < 2
 
       nonterm = @cfg.add_rule(seq)
       @cfg.factor! lhs, nonterm
@@ -55,12 +64,12 @@ class Reducer
   # Loop through pairs of stringified rules, yielding the shorter rule first.
   #
   def each_rule_pair
-    rules = Hash[@cfg.rules.map { |lhs, rhs| [lhs, rhs.join('')] }]
-
-    rules.each_index { |rule1, i| rules.each_index { |rule2, j|
-      next unless j < i
-      yield [rule1, rule2].sort_by { |(_, rhs)| rhs.size }
-    } }
[…]

### 2012-12-02 src/graphplot.rb
Using edge weights rather than multiple edges for graphs
Note: edge weights seem to be broken.

@@ -1,27 +1,26 @@
-require 'rubygems'
+#
+# graphplot.rb - Visualization for grammars.
+#
+
 require 'graphviz'
 
 require_relative 'cfg.rb'
 
-def plotCFG(gr)
-# Create a new graph
-  g = GraphViz.new( :G, :type => :digraph, :ordering => :out )
-  nodes = {}
-  gr.rules.each do |curKey,_|
-    nodes[curKey] = g.add_nodes(curKey)
-  end
-  gr.rules.each do |curKey,curRHS|
-    gn1 = nodes[curKey]
-    curRHS.each_value do |sym|
-      if gr.nonterm? sym then
-        g.add_edges(gn1,nodes[sym])
-      end
-    end
-  end
-  return g
-end
+def plot_cfg(cfg, fname)
+  GraphViz.new(:G, :type => :digraph, :ordering => :out) do |g|
+    cfg.rules.each { |lhs, _| g.add_nodes(lhs) }
 
-def outputCFG(gr,fname)
-  g = plotCFG(gr)
-  g.output(:png => fname)
+    cfg.rules.each do |lhs, rhs|
+      lhs_node = g.get_node(lhs)
+
+      rhs.values.select { |sym|
+        cfg.nonterm? sym
+      }.inject(Hash.new(0)) { |counts, sym|
+        counts[sym] += 1
+        counts
+      }.each { |sym, count|
+        g.add_edges(lhs_node, g.get_node(sym), :weight => count)
+      }
+    end
+  end.output(:png => fname + '.png')
 end

### 2012-12-02 src/list.rb

@@ -21,8 +21,8 @@ class List
     end
   end
 
-  def each_value
-    each { |node| yield node.value }
+  def values
+    map(&:value)
   end
 
   #

### 2012-12-02 src/lossify.rb

@@ -2,7 +2,6 @@
 # lossify.rb - CFG lossifier algorithms.
 #
 
-require 'rubygems'
 require 'levenshtein'
 
 module Lossifier

### 2012-12-02 src/main.rb

@@ -74,9 +74,10 @@ if ARGV[0].nil?
   exit
 end
 
-def puts_cfg(cfg, options)
+def output_cfg(cfg, options, plotname)
   puts cfg if options.print
   puts cfg.expand if options.expand
+  plot_cfg cfg, plotname if options.plot
   puts "\n"
 end
 
@@ -87,14 +88,11 @@ cfg = Sequitur.new(str).run
 cfg = Reducer.new(cfg).run if options.reduce
 
 puts "Sequitur-----------------------------------------------\n\n"
-puts_cfg cfg, options
-outputCFG(cfg, fprefix + "-orig.png") if options.plot
+output_cfg cfg, options, fprefix + '-orig'
 
 options.algorithms.each do |algo|
   puts algo.name + '-' * (55 - algo.name.size) + "\n\n"
 
   lossy_cfg = algo.new(cfg, options.verbose).run
-  puts_cfg(lossy_cfg, options)
-
-  outputCFG(lossy_cfg, fprefix + "-" + algo.name + ".png") if options.plot
+  output_cfg lossy_cfg, options, fprefix + '-' + algo.name
 end

### 2012-12-02 src/cfg.rb
More Reducer fixes
Still broken, though.

@@ -149,7 +149,6 @@ class CFG
       pos = 0
       rhs = rhs.inject(List.new) do |list, node|
         val = node.value
-        found = false
 
         rel = i - pos
         if rel >= 0 and rel < val.size
@@ -157,8 +156,6 @@ class CFG
           prefix = val[0...rel]
           list << prefix if prefix.size > 0
           list << nonterm
-
-          found = true
         end
 
         rel = i + seq.size - pos
@@ -166,14 +163,16 @@ class CFG
           # The end of seq is in this node.
           suffix = val[rel..-1]
           list << suffix if suffix.size > 0
+        end
 
-          found = true
+        if pos + val.size <= i and pos >= i + seq.size
+          list << val
         end
 
         # Increment position pointer.
         pos += val.size
 
-        found ? list : list << node.value
+        list
       end
     end
 

### 2012-12-02 src/reduce.rb

@@ -22,7 +22,7 @@ class Reducer
                    unify_pairwise,
                    apply_rules,
                    eliminate_duplicates]).any?
-      puts ">   [#{i}]".ljust(10) + res.map { |e| e ? 1 : 0 }.join(', ')
+      puts ">   [#{i += 1}]".ljust(10) + res.map { |e| e ? 1 : 0 }.join(', ')
     end
     @cfg
   end
@@ -107,7 +107,7 @@ class Reducer
     match_reduce(
       ->(rhs1, rhs2) {
         s = lcs(rhs1, rhs2)
-        nsymbs(s) >= 2 && s
+        nsymbs(s) >= 2 && nsymbs(rhs1) > 2 && nsymbs(rhs2) > 2 && s
       },
       ->(lhs1, lhs2, seq) {
         nonterm = @cfg.add_rule(seq)

### 2012-12-02 src/reduce.rb
Slight reorganization

@@ -44,8 +44,6 @@ class Reducer
     end
   end
 
-  class Retry < StandardError; end
-
   #
   # R2: Unify pairs of substrings (of length at least 2) within a single rule
   # by making a new rule for them.
@@ -71,6 +69,8 @@ class Reducer
       end
     end
   end
+
+  class Retry < StandardError; end
 
   #
   # General method for finding substring matches between pairs of RHS

### 2012-12-02 src/cfg.rb
fixed boolean test

@@ -165,7 +165,7 @@ class CFG
           list << suffix if suffix.size > 0
         end
 
-        if pos + val.size <= i and pos >= i + seq.size
+        if pos + val.size <= i or pos >= i + seq.size
           list << val
         end
 

### 2012-12-02 src/main.rb

@@ -94,5 +94,10 @@ options.algorithms.each do |algo|
   puts algo.name + '-' * (55 - algo.name.size) + "\n\n"
 
   lossy_cfg = algo.new(cfg, options.verbose).run
+  red_lossy_cfg = Reduce.new(lossy_cfg).run if options.reduce
   output_cfg lossy_cfg, options, fprefix + '-' + algo.name
+  if options.reduce then
+    puts "Reducer--\n"
+    output_cfg red_lossy_cfg, options, fprefix + '-r' + algo.name
+  end
 end

### 2012-12-03 src/util.rb
Added actual longest common substring for two raw string inputs

@@ -3,39 +3,26 @@
 #
 
 #
-# Longest common substring implementation taken from
-# http://rosettacode.org/wiki/Longest_Common_Subsequence#Ruby
+# Longest common substring implementation adapted from
+# http://en.wikibooks.org/wiki/Algorithm_implementation/Strings/Longest_common_substring
 #
-def lcs(a, b)
-  lengths = Array.new(a.size + 1) { Array.new(b.size + 1) { 0 } }
 
-  # Row 0 and column 0 are initialized to 0 already.
-  a.split('').each_with_index do |x, i|
-    b.split('').each_with_index do |y, j|
-      if x == y
-        lengths[i+1][j+1] = lengths[i][j] + 1
+# Raw string lcs
+def lcs(s1, s2)
+  m = Array.new(s1.length+1){ [0] * (s2.length+1) }
+  longest, x_longest = 0,0
+  (1 .. s1.length).each do |x|
+    (1 .. s2.length).each do |y|
+      if s1[x-1] == s2[y-1]
+        m[x][y] = m[x-1][y-1] + 1
+        if m[x][y] > longest
+          longest = m[x][y]
+          x_longest = x
+        end
       else
-        lengths[i+1][j+1] = [lengths[i+1][j], lengths[i][j+1]].max
+        m[x][y] = 0
       end
     end
   end
-
-  # Read the substring out from the matrix.
-  result = ""
-  x, y = a.size, b.size
-
-  while x != 0 and y != 0
-    if lengths[x][y] == lengths[x-1][y]
-      x -= 1
-    elsif lengths[x][y] == lengths[x][y-1]
-      y -= 1
-    else
-      # Assert a[x-1] == b[y-1]
-      result << a[x-1]
-      x -= 1
-      y -= 1
-    end
-  end
-
-  result.reverse
+  s1[x_longest - longest .. (x_longest-1)]
 end

### 2012-12-03 src/cfg.rb
More Reducer fixes

@@ -54,14 +54,11 @@ class CFG
   end
 
   #
-  # Add a string as a new rule.
+  # Add a new rule.
   #
-  def add_rule(str)
+  def add_rule(rhs)
     nonterm = '~[' + @nonterm.succ! + ']'
-
-    rhs = str.split(/(~\[\w*\])/).select { |s| !s.empty? }
-    @rules[nonterm] = rhs.inject(List.new) { |list, s| list << s }
-
+    @rules[nonterm] = rhs
     nonterm
   end
 
@@ -144,6 +141,9 @@ class CFG
     rhs = @rules[target]
     seq = @rules[nonterm].join('')
 
+    puts '0: ' + seq + "\n\n"
+    puts '1: ' + rhs.join + "\n\n"
+
     # Repeatedly factor out the first occurrence of seq until we're done.
     until (i = rhs.join('').index seq).nil?
       pos = 0
@@ -178,6 +178,7 @@ class CFG
 
     # Replace the rule with our refactored rule.
     @rules[target] = rhs
+    puts '2: ' + rhs.join + "\n\n"
   end
 
   def factor(target, nonterm)

### 2012-12-03 src/list.rb

@@ -25,6 +25,9 @@ class List
     map(&:value)
   end
 
+  def size; inject(0) { |count, _| count + 1 }; end
+  def length; size; end
+
   #
   # Copy a list.  By default, the new list nodes are assigned the values of
   # the existing node values, but an optional block may be passed which deeply

### 2012-12-03 src/main.rb

@@ -85,7 +85,7 @@ str = File.read(ARGV[0])
 fprefix = ARGV[0].rpartition('.').first
 
 cfg = Sequitur.new(str).run
-cfg = Reducer.new(cfg).run if options.reduce
+cfg = Reducer.new(cfg, options.verbose).run if options.reduce
 
 puts "Sequitur-----------------------------------------------\n\n"
 output_cfg cfg, options, fprefix + '-orig'

### 2012-12-03 src/reduce.rb

@@ -3,7 +3,6 @@
 #
 
 require_relative 'cfg.rb'
-require_relative 'util.rb'
 
 class Reducer
   attr_reader :cfg
@@ -22,55 +21,76 @@ class Reducer
                    unify_pairwise,
                    apply_rules,
                    eliminate_duplicates]).any?
-      puts ">   [#{i += 1}]".ljust(10) + res.map { |e| e ? 1 : 0 }.join(', ')
+      if @verbose
+        puts ">   [#{i += 1}]".ljust(10) + res.map { |e| e ? 1 : 0 }.join(', ')
+      end
     end
     @cfg
   end
 
   private
 
-  def nsymbs(s)
-    s.gsub(/~\[\w*\]/, '~').size
-  end
+  class Retry < StandardError; end
 
   #
-  # R1: Eliminate any rules which are used only once in the CFG.
+  # Convert a string of symbols into a list.
   #
-  def eliminate_singletons
-    @cfg.counts.inject(false) do |found, (nonterm, count)|
-      next found if count != 1
-      @cfg.inline! nonterm
-      true
+  def listify(str)
+    str.split(/(~\[\w*\])/).inject(List.new) do |list, s|
+      if @cfg.nonterm s then
+        list << s
+      else
+        # This also handles the s.empty? case.
+        s.split('').each { |c| list << c }
+      end
+      list
     end
   end
 
   #
-  # R2: Unify pairs of substrings (of length at least 2) within a single rule
-  # by making a new rule for them.
+  # Compute a longest common substring given two lists.
   #
-  def unify_internal
-    @cfg.rules_s.each do |(lhs, rhs)|
-      seq = rhs[/((?>~\[.*\]|.)*)((?>~\[.*\]|.)*)\1/, 1]
-      next if seq.nil? or nsymbs(seq) < 2
+  def lcs(l1, l2)
+    k = z = 0
+    m = Array.new(l1.length + 1){ Array.new(l2.length + 1) { 0 } }
 
-      nonterm = @cfg.add_rule(seq)
-      @cfg.factor! lhs, nonterm
+    l1.values.each_with_index do |symb1, i|
+      l2.values.each_with_index do |symb2, j|
+        if symb1 == symb2
+          m[i + 1][j + 1] = m[i][j] + 1
+          if m[i + 1][j + 1] > z
+            z = m[i + 1][j + 1]
[…]

### 2012-12-03 src/util.rb

@@ -1,28 +1,0 @@
-#
-# util.rb - Everybody loves util.
-#
-
-#
-# Longest common substring implementation adapted from
-# http://en.wikibooks.org/wiki/Algorithm_implementation/Strings/Longest_common_substring
-#
-
-# Raw string lcs
-def lcs(s1, s2)
-  m = Array.new(s1.length+1){ [0] * (s2.length+1) }
-  longest, x_longest = 0,0
-  (1 .. s1.length).each do |x|
-    (1 .. s2.length).each do |y|
-      if s1[x-1] == s2[y-1]
-        m[x][y] = m[x-1][y-1] + 1
-        if m[x][y] > longest
-          longest = m[x][y]
-          x_longest = x
-        end
-      else
-        m[x][y] = 0
-      end
-    end
-  end
-  s1[x_longest - longest .. (x_longest-1)]
-end

### 2012-12-03 src/main.rb
Deleting some Reducing from main.rb

@@ -94,10 +94,5 @@ options.algorithms.each do |algo|
   puts algo.name + '-' * (55 - algo.name.size) + "\n\n"
 
   lossy_cfg = algo.new(cfg, options.verbose).run
-  red_lossy_cfg = Reduce.new(lossy_cfg).run if options.reduce
   output_cfg lossy_cfg, options, fprefix + '-' + algo.name
-  if options.reduce then
-    puts "Reducer--\n"
-    output_cfg red_lossy_cfg, options, fprefix + '-r' + algo.name
-  end
 end

### 2012-12-03 src/lossify.rb
Some cleanup and verbosity

@@ -4,6 +4,8 @@
 
 require 'levenshtein'
 
+require_relative 'verb.rb'
+
 module Lossifier
 
   #
@@ -25,14 +27,11 @@ module Lossifier
     def run
       return @cfg if @cfg.rules.size < 2
 
-      puts "> Elimination:" if @verbose
-
-      1.upto(Float::INFINITY) do |i|
+      Verb.loop('Elimination', @verbose) do
         find, repl = eliminate
         break if find.nil? or repl.nil?
-        puts ">   [#{i}]".ljust(10) + find + ':' + repl if @verbose
+        find + ':' + repl
       end
-      puts "\n" if @verbose
 
       @cfg
     end

### 2012-12-03 src/main.rb

@@ -84,10 +84,11 @@ end
 str = File.read(ARGV[0])
 fprefix = ARGV[0].rpartition('.').first
 
+puts "Sequitur-----------------------------------------------\n\n"
+
 cfg = Sequitur.new(str).run
 cfg = Reducer.new(cfg, options.verbose).run if options.reduce
 
-puts "Sequitur-----------------------------------------------\n\n"
 output_cfg cfg, options, fprefix + '-orig'
 
 options.algorithms.each do |algo|

### 2012-12-03 src/reduce.rb

@@ -3,6 +3,7 @@
 #
 
 require_relative 'cfg.rb'
+require_relative 'verb.rb'
 
 class Reducer
   attr_reader :cfg
@@ -13,18 +14,16 @@ class Reducer
   end
 
   def run
-    i = 0
-    puts "> Reduction:" if @verbose
-
-    while (res = [ eliminate_singletons,
-                   unify_internal,
-                   unify_pairwise,
-                   apply_rules,
-                   eliminate_duplicates]).any?
-      if @verbose
-        puts ">   [#{i += 1}]".ljust(10) + res.map { |e| e ? 1 : 0 }.join(', ')
-      end
+    Verb.loop('Reduction', @verbose) do
+      res = [ eliminate_singletons,
+              unify_internal,
+              unify_pairwise,
+              apply_rules,
+              eliminate_duplicates]
+      break unless res.any?
+      res.map { |e| e ? 1 : 0 }.join(', ')
     end
+
     @cfg
   end
 

### 2012-12-03 src/verb.rb

@@ -1,0 +1,19 @@
+#
+# verb.rb - Be a bit verbose.
+#
+
+class Verb
+  def self.loop(name, v, &blk)
+    inner_loop name, v, &blk
+    puts "\n" if v
+  end
+
+  def self.inner_loop(name, v)
+    puts "> " + name + ':' if v
+    Kernel.loop.each_with_index do |_, i|
+      output = yield
+      puts ">   [#{i}]".ljust(10) + output if v
+    end
+  end
+  private_class_method :inner_loop
+end

### 2012-12-03 src/lossify.rb
Oh, begin/ensure/end

@@ -27,7 +27,7 @@ module Lossifier
     def run
       return @cfg if @cfg.rules.size < 2
 
-      Verb.loop('Elimination', @verbose) do
+      verb_loop('Elimination', @verbose) do
         find, repl = eliminate
         break if find.nil? or repl.nil?
         find + ':' + repl

### 2012-12-03 src/reduce.rb

@@ -14,7 +14,7 @@ class Reducer
   end
 
   def run
-    Verb.loop('Reduction', @verbose) do
+    verb_loop('Reduction', @verbose) do
       res = [ eliminate_singletons,
               unify_internal,
               unify_pairwise,

### 2012-12-03 src/verb.rb

@@ -2,18 +2,14 @@
 # verb.rb - Be a bit verbose.
 #
 
-class Verb
-  def self.loop(name, v, &blk)
-    inner_loop name, v, &blk
-    puts "\n" if v
-  end
-
-  def self.inner_loop(name, v)
+def verb_loop(name, v, &blk)
+  begin
     puts "> " + name + ':' if v
     Kernel.loop.each_with_index do |_, i|
       output = yield
       puts ">   [#{i}]".ljust(10) + output if v
     end
+  ensure
+    puts "\n" if v
   end
-  private_class_method :inner_loop
 end

### 2012-12-03 src/cfg.rb
Killing some puts and adding back algo reduce

@@ -141,9 +141,6 @@ class CFG
     rhs = @rules[target]
     seq = @rules[nonterm].join('')
 
-    puts '0: ' + seq + "\n\n"
-    puts '1: ' + rhs.join + "\n\n"
-
     # Repeatedly factor out the first occurrence of seq until we're done.
     until (i = rhs.join('').index seq).nil?
       pos = 0
@@ -178,7 +175,6 @@ class CFG
 
     # Replace the rule with our refactored rule.
     @rules[target] = rhs
-    puts '2: ' + rhs.join + "\n\n"
   end
 
   def factor(target, nonterm)

### 2012-12-03 src/main.rb

@@ -95,5 +95,7 @@ options.algorithms.each do |algo|
   puts algo.name + '-' * (55 - algo.name.size) + "\n\n"
 
   lossy_cfg = algo.new(cfg, options.verbose).run
+  lossy_cfg = Reducer.new(lossy_cfg, options.verbose).run if options.reduce
+
   output_cfg lossy_cfg, options, fprefix + '-' + algo.name
 end

### 2012-12-03 src/main.rb
Niceify main.rb

@@ -81,21 +81,28 @@ def output_cfg(cfg, options, plotname)
   puts "\n"
 end
 
+def reduce_cfg(cfg, options, plotname)
+  cfg = Reducer.new(cfg, options.verbose).run
+  output_cfg cfg, options, plotname + '-red'
+  cfg
+end
+
+def process_cfg(title, options, fprefix)
+  puts title + '-' * (79 - title.size) + "\n\n"
+  cfg = yield
+  output_cfg cfg, options, fprefix + '-' + title
+  reduce_cfg cfg, options, fprefix + '-' + title
+end
+
 str = File.read(ARGV[0])
 fprefix = ARGV[0].rpartition('.').first
 
-puts "Sequitur-----------------------------------------------\n\n"
-
-cfg = Sequitur.new(str).run
-cfg = Reducer.new(cfg, options.verbose).run if options.reduce
-
-output_cfg cfg, options, fprefix + '-orig'
+cfg = process_cfg('Sequitur', options, fprefix) do
+  Sequitur.new(str).run
+end
 
 options.algorithms.each do |algo|
-  puts algo.name + '-' * (55 - algo.name.size) + "\n\n"
-
-  lossy_cfg = algo.new(cfg, options.verbose).run
-  lossy_cfg = Reducer.new(lossy_cfg, options.verbose).run if options.reduce
-
-  output_cfg lossy_cfg, options, fprefix + '-' + algo.name
+  process_cfg(algo.name, options, fprefix) do
+    algo.new(cfg, options.verbose).run
+  end
 end

### 2012-12-03 src/reduce.rb
Period.

@@ -1,5 +1,5 @@
 #
-# reduce.rb - Yang and Kieffer's grammar reduction algorithm
+# reduce.rb - Yang and Kieffer's grammar reduction algorithm.
 #
 
 require_relative 'cfg.rb'

### 2012-12-03 src/reduce.rb
Reducer#unify_internal probably works

@@ -47,7 +47,7 @@ class Reducer
   end
 
   #
-  # Compute a longest common substring given two lists.
+  # Compute a longest common substring given two symbol lists.
   #
   def lcs(l1, l2)
     k = z = 0
@@ -67,6 +67,42 @@ class Reducer
 
     l1.values.each_with_index.inject(List.new) do |l, (symb, i)|
       if i >= k and i < k + z then l << symb else l end
+    end
+  end
+
+  #
+  # Find the longest disjoint substring pair given a symbol list.
+  #
+  def ldsp(l)
+    left_size = l.size / 2
+
+    left, right =
+      l.values.each_with_index.inject([List.new, List.new]) do |(l, r), (v, i)|
+        if i < left_size then [l << v, r] else [l, r << v] end
+      end
+
+    i = size = 0
+    right_s = right.join('')
+
+    catch :done do
+      (1..left_size).reverse_each do |len|
+        left.each_cons(len).each_with_index do |subseq, k|
+          if right_s.include? subseq.join('')
+            i, size = k, len
+            throw :done
+          end
+        end
+      end
+    end
+
+    l.values.each_with_index.inject(List.new) do |list, (symb, k)|
+      break list if k >= i + size
+
+      if k >= i and k < i + size
+        list << symb
+      else
+        list
+      end
     end
   end
 
@@ -134,12 +170,8 @@ class Reducer
   # by making a new rule for them.
   #
   def unify_internal
-    @cfg.rules_s.inject(false) do |found, (lhs, rhs)|
-      # XXX: Assumes the ~[.*] format of nonterminals.
-      seq = rhs[/.*((?>~\[\w*\]|.)*)(?>~\[\w*\]|.)*\1.*/, 1]
-      next found if seq.nil?
-
-      seq = listify(seq)
+    @cfg.rules.dup.inject(false) do |found, (lhs, rhs)|
+      seq = ldsp(rhs)
       next found if seq.size < 2
 
       nonterm = @cfg.add_rule(seq)
@@ -154,7 +186,7 @@ class Reducer
   #
   def unify_pairwise
     match_reduce(
-      ->((rhs1, a), (rhs2, b)) {
+      ->((rhs1, _), (rhs2, _)) {
         seq = lcs(rhs1, rhs2)
         seq.size >= 2 && rhs1.size > 2 && rhs2.size > 2 && seq
       },

### 2012-12-03 src/main.rb
Oh, we have an option not to reduce?  Oh.

@@ -91,7 +91,7 @@ def process_cfg(title, options, fprefix)
   puts title + '-' * (79 - title.size) + "\n\n"
   cfg = yield
   output_cfg cfg, options, fprefix + '-' + title
-  reduce_cfg cfg, options, fprefix + '-' + title
+  reduce_cfg cfg, options, fprefix + '-' + title if options.reduce
 end
 
 str = File.read(ARGV[0])

### 2012-12-03 src/main.rb
Make main work

@@ -89,9 +89,15 @@ end
 
 def process_cfg(title, options, fprefix)
   puts title + '-' * (79 - title.size) + "\n\n"
+
   cfg = yield
   output_cfg cfg, options, fprefix + '-' + title
-  reduce_cfg cfg, options, fprefix + '-' + title if options.reduce
+
+  if options.reduce
+    reduce_cfg cfg, options, fprefix + '-' + title
+  else
+    cfg
+  end
 end
 
 str = File.read(ARGV[0])

### 2012-12-03 src/lossify.rb
Repeated clustering
Also changed Clustering parameter to 0.5.

@@ -82,21 +82,32 @@ module Lossifier
   #
   class Cluster
 
-    def initialize(cfg, verbose=false, epsilon=0.4)
+    def initialize(cfg, verbose=false, epsilon=0.5)
       @cfg = cfg.copy
       @verbose = verbose
       @epsilon = epsilon
     end
 
     def run
-      clusters.each do |cluster|
-        counts = @cfg.counts
-        repl_sym = cluster.max_by { |nonterm| counts[nonterm] }
+      verb_loop('Clustering', @verbose) do
+        cs = clusters
+        sizes = cs.map(&:size)
 
-        cluster.each do |find_sym|
-          next if find_sym == repl_sym
-          @cfg.replace! find_sym, repl_sym
+        break if sizes.count(1) == cs.size
+
+        cs.each do |cluster|
+          counts = @cfg.counts
+          repl_sym = cluster.max_by { |nonterm| counts[nonterm] }
+
+          cluster.each do |find_sym|
+            next if find_sym == repl_sym
+            @cfg.replace! find_sym, repl_sym
+          end
         end
+
+        sizes.uniq.sort.reverse.inject([]) do |a, i|
+          a << (i.to_s + ':' + sizes.count(i).to_s)
+        end.join(', ')
       end
 
       @cfg

### 2012-12-03 src/lossify.rb
Changing Clustering parameter back to 0.4

@@ -82,7 +82,7 @@ module Lossifier
   #
   class Cluster
 
-    def initialize(cfg, verbose=false, epsilon=0.5)
+    def initialize(cfg, verbose=false, epsilon=0.4)
       @cfg = cfg.copy
       @verbose = verbose
       @epsilon = epsilon

### 2012-12-04 src/cfg.rb
rewrote factor to take advantage of hidden invariant, fixed bugs in process. reduce appears to be working well.

@@ -139,42 +139,32 @@ class CFG
   #
   def factor!(target, nonterm)
     rhs = @rules[target]
-    seq = @rules[nonterm].join('')
-
-    # Repeatedly factor out the first occurrence of seq until we're done.
-    until (i = rhs.join('').index seq).nil?
-      pos = 0
-      rhs = rhs.inject(List.new) do |list, node|
-        val = node.value
-
-        rel = i - pos
-        if rel >= 0 and rel < val.size
-          # The start of seq is in this node.
-          prefix = val[0...rel]
-          list << prefix if prefix.size > 0
-          list << nonterm
-        end
-
-        rel = i + seq.size - pos
-        if rel >= 0 and rel < val.size
-          # The end of seq is in this node.
-          suffix = val[rel..-1]
-          list << suffix if suffix.size > 0
-        end
-
-        if pos + val.size <= i or pos >= i + seq.size
-          list << val
-        end
-
-        # Increment position pointer.
-        pos += val.size
-
-        list
-      end
+    seq = @rules[nonterm]
+    seq_a = seq.to_a
+  # counter to make sure not to replace overlapping occurrences
+    offsetwait = 0
+  # pad the RHS to allow each_cons to get to last symbol
+    new_rhs = List.new
+    rhs_pad = rhs.copy
+    for x in (2..seq_a.size)
+      rhs_pad << " "
     end
 
+  # for each candidate sequence in rhs, check to see if we replace
+    rhs_pad.each_cons(seq_a.size) do |rhsseq|
+      if offsetwait == 0 then
+        if rhsseq == seq_a then
+          new_rhs << nonterm
+          offsetwait = seq_a.size-1
+        else
+          new_rhs << rhsseq[0].value
+        end
+      else
+        offsetwait -= 1
+      end
+    end
     # Replace the rule with our refactored rule.
-    @rules[target] = rhs
+    @rules[target] = new_rhs
   end
 
   def factor(target, nonterm)

### 2012-12-04 src/list.rb

@@ -86,6 +86,10 @@ class List
       @value = value
     end
 
+    def ==(node)
+      @value == node.value
+    end
+
     def <=>(node)
       @value <=> node.value
     end

### 2012-12-04 src/main.rb

@@ -94,6 +94,7 @@ def process_cfg(title, options, fprefix)
   output_cfg cfg, options, fprefix + '-' + title
 
   if options.reduce
+    puts "Reduced Version"
     reduce_cfg cfg, options, fprefix + '-' + title
   else
     cfg

### 2012-12-04 src/reduce.rb

@@ -107,7 +107,8 @@ class Reducer
   end
 
   #
-  # Loop through pairs of rules, yielding the shorter rule first.  We yield the
+  # Loop through pairs of (lhs,rhs-pair) pairs, yielding the shorter rule first.  
+  # We yield the
   # rule as well as a stringified version to avoid unnecessary operations.
   #
   def each_rule_pair
@@ -131,7 +132,7 @@ class Reducer
   # General method for finding substring matches between pairs of RHS
   # expansions and then reducing.
   #
-  # @param match    Function which takes two RHS's, returning truthy data on
+  # @param match    Function which takes two RHS-pair's, returning truthy data on
   #                 match or falsey on no-match.
   # @param reduce   Function which takes two LHS's and reduces @cfg.
   # @return   True if any matches were made, else false.
@@ -140,8 +141,8 @@ class Reducer
     found = false
 
     begin
-      each_rule_pair do |(lhs1, rhs1), (lhs2, rhs2)|
-        if (res = match.call(rhs1, rhs2))
+      each_rule_pair do |(lhs1, rhsp1), (lhs2, rhsp2)|
+        if (res = match.call(rhsp1, rhsp2))
           reduce.call(lhs1, lhs2, res)
           found = true
           raise Retry

### 2012-12-04 src/cfg.rb
Cleaning up new factor!
Also, I think the padding of RHS is a bug---since we only care about
subsequences the same length as @rules[nonterm], we don't want each_cons
to start looking at shorter subsequences padded by spaces.  In
particular, if some suffix of @rules[target] plus the padding happens to
match @rules[nonterm], we'd be in trouble.

@@ -139,32 +139,26 @@ class CFG
   #
   def factor!(target, nonterm)
     rhs = @rules[target]
-    seq = @rules[nonterm]
-    seq_a = seq.to_a
-  # counter to make sure not to replace overlapping occurrences
-    offsetwait = 0
-  # pad the RHS to allow each_cons to get to last symbol
-    new_rhs = List.new
-    rhs_pad = rhs.copy
-    for x in (2..seq_a.size)
-      rhs_pad << " "
-    end
+    seq = @rules[nonterm].to_a
 
-  # for each candidate sequence in rhs, check to see if we replace
-    rhs_pad.each_cons(seq_a.size) do |rhsseq|
-      if offsetwait == 0 then
-        if rhsseq == seq_a then
-          new_rhs << nonterm
-          offsetwait = seq_a.size-1
-        else
-          new_rhs << rhsseq[0].value
-        end
-      else
+    # Counter to make sure not to replace overlapping occurrences.
+    offsetwait = 0
+
+    # Check each subsequence in the target rule against nonterm's RHS,
+    # replacing any non-overlapping instances.
+    @rules[target] = rhs.each_cons(seq.size).inject(List.new) do |l, subseq|
+      if offsetwait > 0 then
         offsetwait -= 1
+        next l
+      end
+
+      if subseq == seq then
+        offsetwait = seq.size - 1
+        l << nonterm
+      else
+        l << subseq.first.value
       end
     end
-    # Replace the rule with our refactored rule.
-    @rules[target] = new_rhs
   end
 
   def factor(target, nonterm)

### 2012-12-04 src/main.rb

@@ -94,7 +94,7 @@ def process_cfg(title, options, fprefix)
   output_cfg cfg, options, fprefix + '-' + title
 
   if options.reduce
-    puts "Reduced Version"
+    puts '-' * 70 + "(Reduced)\n\n"
     reduce_cfg cfg, options, fprefix + '-' + title
   else
     cfg

### 2012-12-04 src/reduce.rb

@@ -107,9 +107,9 @@ class Reducer
   end
 
   #
-  # Loop through pairs of (lhs,rhs-pair) pairs, yielding the shorter rule first.  
-  # We yield the
-  # rule as well as a stringified version to avoid unnecessary operations.
+  # Loop through pairs of (lhs, (rhs, rhs_s)), yielding the shorter rule first.
+  # We yield the RHS paired with its stringification to avoid unnecessary
+  # operations.
   #
   def each_rule_pair
     rules = @cfg.rules.dup
@@ -132,8 +132,8 @@ class Reducer
   # General method for finding substring matches between pairs of RHS
   # expansions and then reducing.
   #
-  # @param match    Function which takes two RHS-pair's, returning truthy data on
-  #                 match or falsey on no-match.
+  # @param match    Function which takes two RHS-pair's, returning truthy data
+  #                 on match or falsey on no-match.
   # @param reduce   Function which takes two LHS's and reduces @cfg.
   # @return   True if any matches were made, else false.
   #

### 2012-12-04 src/cfg.rb
Pad with empty strings in factor!

@@ -144,6 +144,10 @@ class CFG
     # Counter to make sure not to replace overlapping occurrences.
     offsetwait = 0
 
+    # Pad rhs with empty strings so that we append any trailing characters when
+    # we loop over subsequences.
+    (seq.size - 1).times { rhs << '' }
+
     # Check each subsequence in the target rule against nonterm's RHS,
     # replacing any non-overlapping instances.
     @rules[target] = rhs.each_cons(seq.size).inject(List.new) do |l, subseq|

### 2012-12-04 src/cfg.rb
refactored process_cfg to better separate cfg dataflow and pure printing

@@ -181,6 +181,14 @@ class CFG
     end
   end
 
+  def gsize
+    cur_size = @rules.inject(0) do |acc, (lhs,rhs)|
+      acc += (1+rhs.size)
+      acc
+    end
+    return cur_size
+  end
+
   private
 
   #

### 2012-12-04 src/graphplot.rb

@@ -19,7 +19,8 @@ def plot_cfg(cfg, fname)
         counts[sym] += 1
         counts
       }.each { |sym, count|
-        g.add_edges(lhs_node, g.get_node(sym), :weight => count)
+        puts count.to_s
+        g.add_edges(lhs_node, g.get_node(sym), :label => count.to_s)
       }
     end
   end.output(:png => fname + '.png')

### 2012-12-04 src/main.rb

@@ -28,9 +28,14 @@ options.print = false
 options.plot = false
 options.reduce = false
 options.verbose = false
+options.analysis = true
 
 OptionParser.new do |opts|
   opts.banner = USAGE
+
+  opts.on("-a", "--[no-]analysis", "Perform final analysis") do |a|
+    options.analysis = a
+  end
 
   opts.on("-e", "--[no-]expand", "Print expansion") do |e|
     options.expand = e
@@ -44,9 +49,9 @@ OptionParser.new do |opts|
     options.plot = p
   end
 
-  opts.on("-l", "--lossifiers [ALGO1, ALGO2, ...]", Array,
+  opts.on("-l", "--lossifiers algo1,algo2,algon", Array,
           "The lossifier algorithms to use") do |algos|
-    if algos = [ 'none' ]
+    if algos == [ 'none' ]
       options.algorithms = []
       next
     end
@@ -87,29 +92,23 @@ def reduce_cfg(cfg, options, plotname)
   cfg
 end
 
-def process_cfg(title, options, fprefix)
+#process_cfg purely does printing
+def process_cfg(title, options, fprefix, cfg)
   puts title + '-' * (79 - title.size) + "\n\n"
-
-  cfg = yield
   output_cfg cfg, options, fprefix + '-' + title
-
-  if options.reduce
-    puts '-' * 70 + "(Reduced)\n\n"
-    reduce_cfg cfg, options, fprefix + '-' + title
-  else
-    cfg
-  end
 end
 
 str = File.read(ARGV[0])
 fprefix = ARGV[0].rpartition('.').first
 
-cfg = process_cfg('Sequitur', options, fprefix) do
-  Sequitur.new(str).run
-end
+icfg = Sequitur.new(str).run
+process_cfg('Sequitur', options, fprefix, icfg)
 
 options.algorithms.each do |algo|
-  process_cfg(algo.name, options, fprefix) do
-    algo.new(cfg, options.verbose).run
+  lcfg = algo.new(icfg, options.verbose).run
+  process_cfg(algo.name, options, fprefix, lcfg)
+  if options.reduce
+    rlcfg = Reducer.new(lcfg, options.verbose).run
+    process_cfg(algo.name+" Reduced", options, fprefix, rlcfg)
   end
 end

### 2012-12-04 src/cfg.rb
Augh cleanup

@@ -31,6 +31,13 @@ class CFG
     end
   end
 
+  #
+  # Number of symbols in the grammar.
+  #
+  def size
+    @rules.inject(0) { |acc, (lhs, rhs)| acc + rhs.size + 1 }
+  end
+
   def nonterm?(str)
     str[0..1] == '~[' && str[-1].chr == ']'
   end
@@ -179,14 +186,6 @@ class CFG
       rhs.each { |node| counts[node.value] += 1 if nonterm? node.value }
       counts
     end
-  end
-
-  def gsize
-    cur_size = @rules.inject(0) do |acc, (lhs,rhs)|
-      acc += (1+rhs.size)
-      acc
-    end
-    return cur_size
   end
 
   private

### 2012-12-04 src/graphplot.rb

@@ -19,7 +19,6 @@ def plot_cfg(cfg, fname)
         counts[sym] += 1
         counts
       }.each { |sym, count|
-        puts count.to_s
         g.add_edges(lhs_node, g.get_node(sym), :label => count.to_s)
       }
     end

### 2012-12-04 src/fidelity.rb
streamlined final outputs

@@ -10,8 +10,6 @@ require 'bzip2'
 require 'descriptive_statistics'
 require 'levenshtein'
 
-USAGE = "Usage: ./fidelity.rb file1 file2"
-
 module Enumerable
   def stats
     { number: number,
@@ -25,7 +23,7 @@ module Enumerable
 
   def stats_s
     stats.inject('') do |s, (k, v)|
-      s + "%7s " % (k.to_s + ':') + v.to_s + "\n"
+      s + "%2s " % (k.to_s + ':') + v.to_s + "; "
     end.chop
   end
 end
@@ -44,26 +42,16 @@ def ngrams(s, n)
   nmap.values.stats_s
 end
 
-def analyze(fn0, fn1, fnout)
-  File.open(fn0) { |f0| File.open(fn1) { |f1| File.open(fnout, 'w+') { |fout|
-    ss = [f0.read, f1.read]
-    sizes = ss.map { |s| s.size }
-
-    [ "Total Length:      %d, %d" % sizes,
-      "Avg Word Length:   %f, %f" % ss.map { |s| avgwordlen s },
-      "Entropy:           %d, %d" % ss.map { |s| entropy s },
-      "Levenshtein:       %f" % [Levenshtein.distance(*ss) / sizes.max.to_f],
-      *(1..3).map do |i|
-        "%d-grams:\n%s\n%s" % [i, ngrams(ss[0], i), ngrams(ss[1], i)]
-      end
-    ].each { |s| fout.puts s }
-  } } }
+# Print outputs that depend on the final string
+# s is processed string, s2 is optional original string
+def fanalyze(s,s2=nil)
+  puts "Final String Statistics"
+  puts ("Total Length:      %d" % [s.size])
+  puts ("Distance:          %f" % 
+          [Levenshtein.distance(s,s2).to_f/s2.size.to_f]) if s2!=nil
+  puts ("Avg Word Length:   %f" % [avgwordlen(s)])
+  puts ("Entropy:           %d" % [entropy(s)])
+  (1..3).map do |i|
+    puts ("%d-grams:\n%s" % [i, ngrams(s, i)])
+  end
 end
-
-if ARGV[0].nil? or ARGV[1].nil?
-  puts USAGE
-  exit
-end
-outfile = ARGV[2] || 'fid.out'
-
-analyze ARGV[0], ARGV[1], outfile

### 2012-12-04 src/main.rb

@@ -13,6 +13,7 @@ require_relative 'lossify.rb'
 require_relative 'reduce.rb'
 require_relative 'sequitur.rb'
 require_relative 'graphplot.rb'
+require_relative 'fidelity.rb'
 
 USAGE = "Usage: ./main.rb [options] input-file"
 ALGORITHMS = {
@@ -23,7 +24,7 @@ ALGORITHMS = {
 # Default option values.
 options = OpenStruct.new
 options.algorithms = ALGORITHMS.values
-options.expand = true
+options.expand = false
 options.print = false
 options.plot = false
 options.reduce = false
@@ -82,6 +83,7 @@ end
 def output_cfg(cfg, options, plotname)
   puts cfg if options.print
   puts cfg.expand if options.expand
+  puts "Size: %d " % [cfg.size] if options.analysis
   plot_cfg cfg, plotname if options.plot
   puts "\n"
 end
@@ -92,8 +94,8 @@ def reduce_cfg(cfg, options, plotname)
   cfg
 end
 
-#process_cfg purely does printing
-def process_cfg(title, options, fprefix, cfg)
+#process_cfg prints out data on the CFG structure
+def process_cfg(title, options, fprefix, cfg, origstr=nil)
   puts title + '-' * (79 - title.size) + "\n\n"
   output_cfg cfg, options, fprefix + '-' + title
 end
@@ -101,14 +103,24 @@ end
 str = File.read(ARGV[0])
 fprefix = ARGV[0].rpartition('.').first
 
+#output format:
+#Sequitur Grammar Properties
+#Original String Properties
+#Foreach Algorithm:
+#  Grammar Properties
+#  (Reduced Grammar Properties)
+#  Final String Properties
+
 icfg = Sequitur.new(str).run
 process_cfg('Sequitur', options, fprefix, icfg)
+fanalyze(str)
 
 options.algorithms.each do |algo|
   lcfg = algo.new(icfg, options.verbose).run
-  process_cfg(algo.name, options, fprefix, lcfg)
+  process_cfg(algo.name, options, fprefix, lcfg, str)
   if options.reduce
     rlcfg = Reducer.new(lcfg, options.verbose).run
-    process_cfg(algo.name+" Reduced", options, fprefix, rlcfg)
+    process_cfg(algo.name+" Reduced", options, fprefix, rlcfg, str)
   end
+  fanalyze(lcfg.expand,str)
 end

### 2012-12-04 src/main.rb
removed unused arg

@@ -83,7 +83,8 @@ end
 def output_cfg(cfg, options, plotname)
   puts cfg if options.print
   puts cfg.expand if options.expand
-  puts "Size: %d " % [cfg.size] if options.analysis
+  puts "Rules: %d" % [cfg.rules.size] if options.analysis
+  puts "Size: %d" % [cfg.size] if options.analysis
   plot_cfg cfg, plotname if options.plot
   puts "\n"
 end
@@ -95,7 +96,7 @@ def reduce_cfg(cfg, options, plotname)
 end
 
 #process_cfg prints out data on the CFG structure
-def process_cfg(title, options, fprefix, cfg, origstr=nil)
+def process_cfg(title, options, fprefix, cfg)
   puts title + '-' * (79 - title.size) + "\n\n"
   output_cfg cfg, options, fprefix + '-' + title
 end
@@ -117,10 +118,10 @@ fanalyze(str)
 
 options.algorithms.each do |algo|
   lcfg = algo.new(icfg, options.verbose).run
-  process_cfg(algo.name, options, fprefix, lcfg, str)
+  process_cfg(algo.name, options, fprefix, lcfg)
   if options.reduce
     rlcfg = Reducer.new(lcfg, options.verbose).run
-    process_cfg(algo.name+" Reduced", options, fprefix, rlcfg, str)
+    process_cfg(algo.name+" Reduced", options, fprefix, rlcfg)
   end
   fanalyze(lcfg.expand,str)
 end

### 2012-12-04 src/cfg.rb
timing data

@@ -35,7 +35,7 @@ class CFG
   # Number of symbols in the grammar.
   #
   def size
-    @rules.inject(0) { |acc, (lhs, rhs)| acc + rhs.size + 1 }
+    @rules.inject(0) { |acc, (lhs, rhs)| acc + rhs.size }
   end
 
   def nonterm?(str)

### 2012-12-04 src/main.rb

@@ -29,7 +29,7 @@ options.print = false
 options.plot = false
 options.reduce = false
 options.verbose = false
-options.analysis = true
+options.analysis = false
 
 OptionParser.new do |opts|
   opts.banner = USAGE
@@ -96,8 +96,9 @@ def reduce_cfg(cfg, options, plotname)
 end
 
 #process_cfg prints out data on the CFG structure
-def process_cfg(title, options, fprefix, cfg)
+def process_cfg(title, options, fprefix, time, cfg)
   puts title + '-' * (79 - title.size) + "\n\n"
+  puts "Time: %f" % [time]
   output_cfg cfg, options, fprefix + '-' + title
 end
 
@@ -112,16 +113,22 @@ fprefix = ARGV[0].rpartition('.').first
 #  (Reduced Grammar Properties)
 #  Final String Properties
 
+t1 = Time.now
 icfg = Sequitur.new(str).run
-process_cfg('Sequitur', options, fprefix, icfg)
-fanalyze(str)
+t2 = Time.now
+process_cfg('Sequitur', options, fprefix, t2-t1, icfg)
+fanalyze(str) if options.analysis
 
 options.algorithms.each do |algo|
+  t1 = Time.now
   lcfg = algo.new(icfg, options.verbose).run
-  process_cfg(algo.name, options, fprefix, lcfg)
+  t2 = Time.now
+  process_cfg(algo.name, options, fprefix, t2-t1, lcfg)
   if options.reduce
+    t1 = Time.now
     rlcfg = Reducer.new(lcfg, options.verbose).run
-    process_cfg(algo.name+" Reduced", options, fprefix, rlcfg)
+    t2 = Time.now
+    process_cfg(algo.name+" Reduced", options, fprefix, t2-t1, rlcfg)
   end
-  fanalyze(lcfg.expand,str)
+  fanalyze(lcfg.expand,str) if options.analysis
 end

### 2012-12-04 src/lossify.rb
changed algorithm to replace with rule with shortest expansion to prevent all cycles

@@ -18,7 +18,7 @@ module Lossifier
   #
   class Similarity
 
-    def initialize(cfg, verbose=false, threshold=0.5)
+    def initialize(cfg, verbose=false, threshold=0.4)
       @cfg = cfg.copy
       @verbose = verbose
       @threshold = threshold
@@ -92,12 +92,16 @@ module Lossifier
       verb_loop('Clustering', @verbose) do
         cs = clusters
         sizes = cs.map(&:size)
+        sym_expand_len = @cfg.rules.inject(Hash.new(0)) do |lens, (lhs,_)|
+          lens[lhs] = @cfg.expand(lhs).size
+          lens
+        end
 
         break if sizes.count(1) == cs.size
 
         cs.each do |cluster|
           counts = @cfg.counts
-          repl_sym = cluster.max_by { |nonterm| counts[nonterm] }
+          repl_sym = cluster.min_by { |nonterm| sym_expand_len[nonterm] }
 
           cluster.each do |find_sym|
             next if find_sym == repl_sym

### 2012-12-04 src/main.rb

@@ -128,7 +128,7 @@ options.algorithms.each do |algo|
     t1 = Time.now
     rlcfg = Reducer.new(lcfg, options.verbose).run
     t2 = Time.now
-    process_cfg(algo.name+" Reduced", options, fprefix, t2-t1, rlcfg)
+    process_cfg(algo.name+"-Reduced", options, fprefix, t2-t1, rlcfg)
   end
   fanalyze(lcfg.expand,str) if options.analysis
 end

### 2012-12-04 src/lossify.rb
make similarity follow same rule as cluster

@@ -54,15 +54,9 @@ module Lossifier
           if dist < @threshold
             counts = @cfg.counts
 
-            # Replace the less common variable, or, in the case of a tie,
-            # the variable with the longer rule.
-            if counts[nonterm1] >= counts[nonterm2]
-              @cfg.replace! nonterm2, nonterm1
-              ret = [nonterm2, nonterm1]
-            else
-              @cfg.replace! nonterm1, nonterm2
-              ret = [nonterm1, nonterm2]
-            end
+            # replace longer rule with shorter rule
+            @cfg.replace! nonterm2, nonterm1
+            ret = [nonterm2, nonterm1]
 
             return ret
           end

### 2012-12-04 src/lossify.rb
added file for automatically introducing errors into other files. Added some clean,large inputs for use in final results

@@ -18,7 +18,7 @@ module Lossifier
   #
   class Similarity
 
-    def initialize(cfg, verbose=false, threshold=0.4)
+    def initialize(cfg, verbose=false, threshold=0.3)
       @cfg = cfg.copy
       @verbose = verbose
       @threshold = threshold
@@ -76,7 +76,7 @@ module Lossifier
   #
   class Cluster
 
-    def initialize(cfg, verbose=false, epsilon=0.4)
+    def initialize(cfg, verbose=false, epsilon=0.3)
       @cfg = cfg.copy
       @verbose = verbose
       @epsilon = epsilon

### 2012-12-04 src/mkerror.rb

@@ -1,0 +1,15 @@
+pchange = 30
+
+str = File.read(ARGV[0])
+out_f = File.new("e_"+ARGV[0], "w+")
+
+newstr = str.chars.inject("") do |str, c|
+  if rand(pchange)==0
+    str << '*'
+  else
+    str << c
+  end
+end
+
+out_f.write newstr
+out_f.close

### 2012-12-05 tex/paper.tex
first sections of final paper, make sure you have the latex package algorithmicx installed

\documentclass[11pt]{article}
\usepackage{amsfonts,amsmath,amssymb,amsthm}
\usepackage{array}
\usepackage{bm}
\usepackage{commath}
\usepackage{enumitem}
\usepackage{fancyhdr,fancyvrb}
\usepackage[letterpaper,text={7in,9in},centering]{geometry}
\usepackage[pdftex]{graphicx}
\usepackage{ifthen}
\usepackage{mathrsfs}
\usepackage{multicol}
\usepackage{pifont}
\usepackage{sectsty}
\usepackage{setspace}
\usepackage{stackrel}
\usepackage{stmaryrd}
\usepackage{tensor}
\usepackage{xspace}
\usepackage{natbib}
\usepackage{algorithm}
\usepackage{algpseudocode}

\pagestyle{fancy}
\fancyhead[RO]{Edward Gan \& Max Wang}

\sectionfont{\large}
\subsectionfont{\normalsize}

\newcommand{\Sequitur}{\textsc{Sequitur}\xspace}

\begin{document}
\bibliographystyle{plain}

\title{Squint: Lossy Hierarchical Compression on Symbol Streams}
\author{Edward Gan \& Max Wang}
\date{December 11, 2012}
\maketitle

\begin{abstract}
Grammar-based compression schemes like \Sequitur exploit repeated structures
in streams of symbolic data to achieve good compression ratios.
We develop a set of algorithms for improving upon Grammar-based compression 
schemes by allowing some loss in the compression. Although precise 
subsequences might be altered, in return we can simplify 
the grammars produced to obtain better ratios while still preserving
important syntatic patterns at both the micro and macro level.
\end{abstract}

\section{Introduction and Goals}
Many sources of data exhibit natural hierarchical structure. For instance,
english text often can be grouped well into sections, paragraphs, sentences, ...
all the way down to words and phonemes. Music too, consists of repeated motifs,
phrases, themes, and so on. Finally, computer code is so hierarchical that its
syntax can usually be put in BNF. This allows grammar-based compression schemes
to build up Context Free Grammars (CFGs) and achieve very good compression
ratios.

Taking a step back, in the image compression world there are both lossless and
lossy compression schemes. In certain situations, such as for low-pixel count
logos, one could use a lossless image compression scheme. In other situations
however, such as for textures and some photographs, the exact pixel values
are almost irrelevant compared to the overall gradients, colors, and structures
in the image, and one could use a lossy compression scheme like jpeg.

We believe it is valuable to bring some of the attitudes prevalent in image/
video compression to the world of hierarchical text compression. Better
compression ratios for text are possible.

It may seem strange to allow a string such as "mind" to be lossily 
corrupted into ``rind'', but if one is more interested in recurrent 
structure in the text than in its precise meaning, these changes 
aren't important. For instance, if one is looking at a text in a foreign
language, one is more likely to pick up the texture, the repeated sounds,
some phrases that come up and the breaks between larger scale sections.
A change in a single letter probably wouldn't be noticed. Similarly,
when skimming through huge files, perhaps execution logs, one usually
first starts looking for patterns or things that stand out. Lorem ipsum
is a perfect example of a piece of ``text'' whose value lies in its texture
and sentence/paragraph structure, and we will come back to it again.

In this paper we develop 2 algorithms and implement a system for lossily
compressing text while preserving its hierarchical structures. The goal
was to preserve the feel of a text at both micro and macro levels. 
Since grammar-based compression schemes were designed to capture
these structures, we built
our system on top of the existing \Sequitur grammar inference algorithm. 
Starting with the CFG generated by a lossless grammar
inference algorithm, we repeatedly simplify it, throwing out details
along the way to emphasize repetitions. Our main contribution lies in the
development of 2 algorithms for ``simplifying'' CFG grammars while 
preserving their structure.

\section{Background}

TODO

\subsection{Grammar Compression}

TODO

\subsection{Lossy Compression}

TODO

Jpeg, Mpeg: using parts of previous frame to encode next frame, 
Fractal: Using self-similarity to encode

\section{Lossifier Algorithms}

\subsection{Motivation}

The scope of what we shall call a \emph{lossifier} algorithm is to take
an admissible CFG and simplify it. The output CFG should be smaller
than the input CFG but generate a string which preserves the
structure in the original string as much as possible.

Building off the ideas in jpeg compression discussed
earlier, one idea for simplifying data is to look at a part
of the data and try to match it as closely as possible with an
existing set of primitives. In particular, just as with Mpeg and
fractal compression, to approximate one part of the grammar
we can use \emph{other parts of the grammar}. This is the seed for
both of our algorithms. Given a rule $S \rightarrow rhs$, we will
try to replace occurences of $S$ with other similar variables.

For example, consider the grammars in figure \ref{sred}.

\begin{figure}[t]
\centering
\includegraphics[scale=0.6]{sred1.pdf}
\includegraphics[scale=0.6]{sred2.pdf}
\caption{Grammar Simplification Example}
\label{sred}
\end{figure}

To simplify the grammar, it would make sense to replace B with A, since
they form two almost identical halves of S and are similar.

\subsection{Pair-Similarity}

To define whether two variables are \emph{similar}, we can look at their
expansions. Let $expand(rhs)$ denote the string of terminal symbols we get
when we recursively substitute all of the variables inside $rhs$. Then
$V_1 \rightarrow rhs_1$ and $V_2 \rightarrow rhs_2$ are similar when

$$\frac{levenshtein(expand(rhs_1),expand(rhs_2))}
        {max(len(expand(rhs_1)),len(expand(rhs_2)))} < \epsilon$$.

where $levenshtein$ is the levenstein edit distance between two strings.
For convenience we can define.

$$lsim(s1,s2) = \frac{levenshtein(s1,s2)}{max(len(s1),len(s2))}$$

The first algorithm takes the idea of variable similarity and uses it
to simplify the grammar sequentially as much as possible.

\begin{algorithm}
\caption{Similarity Lossifier Algorithm}
\label{sim_alg}
\begin{algorithmic}[1]
\Procedure{Lossify-Sim}{$g$} \Comment{Simplify grammar g}
  \While{there are similar variables to replace}
    \For{$var_1 \rightarrow rhs_1,\ var_2 \rightarrow rhs_2 \in g$}
      \State $str_{i} \gets expand(rhs_i)$
      \If{$lsim(str_{1},str_{2})<\epsilon$}
        \State replace the longer var with the shorter throughout $g$
        \State break
      \EndIf
    \EndFor
  \EndWhile
\EndProcedure
\end{algorithmic}
\end{algorithm}

\subsection{Clustering}

The previous algorithm updates the grammar sequentially, since each
variable replacement can have an effect on large parts of the grammar,
we will have to go through comparing all pairs again after the change.
An alternative algorithm would be to try to replace similar variables in
parallel. This would avoid many of the repeat pair-comparisons done in
\emph{Similarity}. In other words, the algorithm below makes puts 
similar variables
into equivalence class clusters, and then replaces them with a representative
all at once each iteration.

\begin{algorithm}
\caption{Cluster Lossifier Algorithm}
\label{cluster_alg}
\begin{algorithmic}[1]
\Procedure{Cluster-Sim}{$g$} \Comment{Simplify grammar g}
  \While{there are similar variables to replace}
    \State $clusters \gets ()$
    \For{$var_1 \rightarrow rhs_1 \in g$}
      \State blah
    \EndFor
  \EndWhile
\EndProcedure
\end{algorithmic}
\end{algorithm}

\subsection{Mathematical Analysis}
Runtime Analysis
No-loops proof

\section{Implementation}
Overview

\subsection{Grammar Reductions}
What is their purpose? Why do we want them? What are they?

\subsection{Data Structures and Optimizations}
Grammar Data Structure
Cluster Data Structure
Invariants Maintained
Core utility methods??
Linked Lists vs Arrays, Caching Expansions, Strings vs Symbols for expansion, Native Levenshtein

\section{Results}
\subsection{Illustrative Examples}
Qualitative, show off diagrams and detailed metrics for small examples
\subsection{Similarity vs Clustering}
Fixed compression ratio + fidelity comparison for different ratios
Show the runtime difference
\subsection{Grammar Reduction Impact}
\subsection{Larger Tests}

\section{Conclusions}
\subsection{Ratios and Fidelity}
\subsection{Future Work}

\nocite{*}

\bibliography{sources}

\end{document}

### 2012-12-05 compile.sh

@@ -1,6 +1,0 @@
-#!/bin/bash
-
-pdflatex proposal
-bibtex proposal
-pdflatex proposal
-pdflatex proposal

### 2012-12-05 tex/compile.sh

@@ -1,0 +1,6 @@
+#!/bin/bash
+
+pdflatex paper
+bibtex paper
+pdflatex paper
+pdflatex paper

### 2012-12-06 tex/paper.tex
Began filling in results section, collected some final results

@@
 similar variables
-into equivalence class clusters, and then replaces them with a representative
+into equivalence class clusters, and then replaces each variable
+in a cluster them with a representative
 all at once each iteration.
@@
 \subsection{Illustrative Examples}
-Qualitative, show off diagrams and detailed metrics for small examples
+It is useful to look at how the algorithm functions on a small piece of
+data with both plenty of hierarchical structure, and also some messiness
+that most lossless compression schemes would not deal well with. Take
+the string below, similar to what one might come across in a math
+puzzle book.
+
+\begin{verbatim}
+123,124,123 + 321,323,321 = ???
+124,123,123 + 321,3231,321 = ????
+--
+124,123,123 + 321,3231,321 = ????
+123,124,123 + 321,323,321 = ????
+\end{verbatim}
+
+Disregarding its meaning, it consists of four very similar equations of the
+form $a + b = ?$, and on a smaller scale each term consists of
+3 repetitions of approximately the same string of either $123$ or $321$.
+
+Sequitur yields a relatively complicated inferred grammar with 13 rules and
+a size of 51. Both the Similarity and Clustering Algorithms with $\epsilon = .4$
+, yield the following grammar after reduction. It has 5 rules and is almost
+half the size, at 30.
+
+\begin{verbatim}
+~[*] => ~[BC]--~[BC]?
+
+~[AZ] => ~[A]3~[A]~[A] + ~[C]~[C]3~[C]1 = ???
+~[A] => 12
+~[BC] => ~[AZ]~[AZ]
+~[C] => 32
+-----------------------
+1231212 + 32323321 = ???1231212 + 32323321 = ???--(cont.)
+1231212 + 32323321 = ???1231212 + 32323321 = ????
+\end{verbatim}
+
+The lossifier algorithm reduced the string to a pair of
+pairs of the same equations, where each equation consists of the
+sum of triples of "12" or "32". Smaller discrepencies
+in the numbers involved were dropped, while the filler
+commas and linebreaks also ended
+up being lost. Note that the structure described is easily
+observable from the grammar generated.
+
 \subsection{Similarity vs Clustering}
-Fixed compression ratio + fidelity comparison for different ratios
-Show the runtime difference
+For most of our benchmark tests, we focused on two extreme kinds of data:
+one with large amounts of repetition at all levels, and one with a word
+and sentence level structure but with little obvious large-scale repetition.
+One set of test data was constructed by repeating the same letter `a' n-times
+but then adding some
+texture to the data by changing 1/30 of the characters uniformly
+to `a'. Call these test-cases \emph{rep-n}
+Another set of test data was constructed from the first n characters of the
+book of Genesis from the KJV. Call these test-cases \emph{gen-n}
+
+For our first suite of tests, we compare the effectiveness of the
+\emph{Similarity} and \emph{Cluster} algorithms. We might expect
+Similarity to achieve better quality at the same compression ratio
+since it updates the state of the grammar between each variable
+replacement, on the other hand since it must update the grammar
+it has the potential to run much more slowly than Cluster.
+
+We don't report the running times for just \Sequitur since its linear
+running time is negligible compared to our algorithms for all but the
+smallest input data.
+
+\begin{table}[t]
+\begin{tabular}{l|l|l|l|l|l|l}
+Dataset & S-T(ime) & S-R(atio) & S-L(evdistance) & C-T & C-R & C-L \\
+\hline
+rep-100   & 0.006 & 18/34    & 0.713 & 0.002 & 20/34    & 0.535\\
+rep-400   & 0.028 & 45/76    & 0.269 & 0.009 & 41/76    & 0.743\\
+rep-1600  & 0.355 & 103/194  & 0.873 & 0.042 & 103/194  & 0.804\\
+rep-6400  & 3.978 & 258/448  & 0.706 & 0.203 & 256/448  & 0.869\\
+rep-25600 & 135.0 & 900/1482 & 0.829 & 2.906 & 898/1482 & 0.725\\
+\end{tabular}
+\caption{$\epsilon = .4$}
+\end{table}
+
+\begin{table}[t]
+\begin{tabular}{l|l|l|l|l|l|l}
+Dataset & S-T(ime) & S-R(atio) & S-L(evdistance) & C-T & C-R & C-L \\
+\hline
+rep-100   & 0.002 & 30/34    & 0.040 & 0.002 & 30/34    & 0.040\\
+rep-400   & 0.019 & 60/76    & 0.035 & 0.009 & 60/76    & 0.035\\
+rep-1600  & 0.350 & 123/194  & 0.355 & 0.055 & 123/194  & 0.206\\
+rep-6400  & 3.942 & 280/448  & 0.387 & 0.295 & 274/448  & 0.543\\
+rep-25600 & 124.5 & 922/1482 & 0.565 & 2.597 & 926/1482 & 0.437\\
+\end{tabular}
+\caption{$\epsilon = .2$}
+\end{table}
+
+TODO: Fill in results for $\epsilon = .1$
+
+TODO: Runtime graphs and regression
+
+Qualitatively the output strings generally are strings of a's with interspersed
+`*' as expected, but the length of the final output varies widely.
+
+Things to mention: Compression ratios and fidelity comparable for similarity
+and cluster. Cluster has much better growth rate. Smaller epsilon provide
+much better value/cost ratio as the input files get bigger for highly
+repetitive files.
+
+Thus, we see that similarity provides no consistent advantage over cluster
+and has much less scalable performance, with an estimated running time of 
+BLAH. For the rest of the paper we will only test Cluster.
+
 \subsection{Grammar Reduction Impact}
+benchmark two extreme test cases again, with and without reducer
+see how much reducer helps (it shouldn't help too much)
+notice how slow it is
+
 \subsection{Larger Tests}
+Final compression ratios on benchmark data
+ASCII ART =p
+Our Code
 
@@
 \subsection{Future Work}
+> Better ways to set epsilon to maximize value / cost for different size and types
+> Better ways to quantify preservation of hierarchies
+> How to find near `self-similarities' that aren't captured by a rule
+> Iterating lossify and reduce in a cycle
 

### 2012-12-06 src/lossify.rb

@@ -18,7 +18,7 @@ module Lossifier
   #
   class Similarity
 
-    def initialize(cfg, verbose=false, threshold=0.3)
+    def initialize(cfg, verbose=false, threshold=0.4)
       @cfg = cfg.copy
       @verbose = verbose
       @threshold = threshold
@@ -76,7 +76,7 @@ module Lossifier
   #
   class Cluster
 
-    def initialize(cfg, verbose=false, epsilon=0.3)
+    def initialize(cfg, verbose=false, epsilon=0.4)
       @cfg = cfg.copy
       @verbose = verbose
       @epsilon = epsilon

### 2012-12-06 tex/paper.tex
added parameter for controliling threshold programmatically

@@
 texture to the data by changing 1/30 of the characters uniformly
-to `a'. Call these test-cases \emph{rep-n}
+to `*'. Call these test-cases \emph{rep-n}
 Another set of test data was constructed from the first n characters of the
@@
 rep-6400  & 3.978 & 258/448  & 0.706 & 0.203 & 256/448  & 0.869\\
+rep-12800
 rep-25600 & 135.0 & 900/1482 & 0.829 & 2.906 & 898/1482 & 0.725\\
@@
 rep-6400  & 3.942 & 280/448  & 0.387 & 0.295 & 274/448  & 0.543\\
+rep-12800  
 rep-25600 & 124.5 & 922/1482 & 0.565 & 2.597 & 926/1482 & 0.437\\
@@
 
-TODO: Runtime graphs and regression
+TODO: Runtime graphs and regression. The exponents are about 1.8 and 1.3
+respectively, note that the rulesize of sequitur grows at about sqrt(n)
 
@@
 repetitive files.
+
+TODO: Analogous for genesis
+> Growth rate for similarity is far worse for genesis
 
@@
 \subsection{Larger Tests}
-Final compression ratios on benchmark data
+Final compression ratios on various benchmark data
 ASCII ART =p

### 2012-12-06 src/lossify.rb

@@ -18,7 +18,7 @@ module Lossifier
   #
   class Similarity
 
-    def initialize(cfg, verbose=false, threshold=0.4)
+    def initialize(cfg, verbose=false, threshold)
       @cfg = cfg.copy
       @verbose = verbose
       @threshold = threshold
@@ -76,7 +76,7 @@ module Lossifier
   #
   class Cluster
 
-    def initialize(cfg, verbose=false, epsilon=0.4)
+    def initialize(cfg, verbose=false, epsilon)
       @cfg = cfg.copy
       @verbose = verbose
       @epsilon = epsilon

### 2012-12-06 src/main.rb

@@ -30,9 +30,14 @@ options.plot = false
 options.reduce = false
 options.verbose = false
 options.analysis = false
+options.thresh = 0.4
 
 OptionParser.new do |opts|
   opts.banner = USAGE
+
+  opts.on("--t N", Float, "Threshold level") do |t|
+    options.thresh = t
+  end
 
   opts.on("-a", "--[no-]analysis", "Perform final analysis") do |a|
     options.analysis = a
@@ -121,7 +126,7 @@ fanalyze(str) if options.analysis
 
 options.algorithms.each do |algo|
   t1 = Time.now
-  lcfg = algo.new(icfg, options.verbose).run
+  lcfg = algo.new(icfg, options.verbose, options.thresh).run
   t2 = Time.now
   process_cfg(algo.name, options, fprefix, t2-t1, lcfg)
   if options.reduce

### 2012-12-08 src/fidelity.rb
Code cleanup and output prettifying

@@ -1,14 +1,13 @@
-#!/usr/local/bin/ruby
-
 #
 # fidelity.rb - Analysis framework for comparing the fidelity between a string
 # and its compressed output.
 #
 
-require 'rubygems'
 require 'bzip2'
 require 'descriptive_statistics'
 require 'levenshtein'
+
+require_relative 'util.rb'
 
 module Enumerable
   def stats
@@ -23,7 +22,7 @@ module Enumerable
 
   def stats_s
     stats.inject('') do |s, (k, v)|
-      s + "%2s " % (k.to_s + ':') + v.to_s + "; "
+      s + "%2s " % (k.to_s + ':') + v.to_s + "\n"
     end.chop
   end
 end
@@ -42,16 +41,18 @@ def ngrams(s, n)
   nmap.values.stats_s
 end
 
-# Print outputs that depend on the final string
-# s is processed string, s2 is optional original string
-def fanalyze(s,s2=nil)
+#
+# Perform some statistical analysis on a string.  If a second string is
+# supplied as an argument, we compare `s' against it.
+#
+def fanalyze(s, s2=nil)
   puts "Final String Statistics"
   puts ("Total Length:      %d" % [s.size])
-  puts ("Distance:          %f" % 
-          [Levenshtein.distance(s,s2).to_f/s2.size.to_f]) if s2!=nil
+  puts ("Distance:          %f" %
+        [Levenshtein.distance(s, s2).to_f / s2.size.to_f]) if s2 != nil
   puts ("Avg Word Length:   %f" % [avgwordlen(s)])
   puts ("Entropy:           %d" % [entropy(s)])
   (1..3).map do |i|
-    puts ("%d-grams:\n%s" % [i, ngrams(s, i)])
+    puts ("%d-grams:\n%s" % [i, ngrams(s, i).indent(4)])
   end
 end

### 2012-12-08 src/lossify.rb

@@ -54,11 +54,9 @@ module Lossifier
           if dist < @threshold
             counts = @cfg.counts
 
-            # replace longer rule with shorter rule
+            # Replace the longer rule with the shorter rule.
             @cfg.replace! nonterm2, nonterm1
-            ret = [nonterm2, nonterm1]
-
-            return ret
+            return [nonterm2, nonterm1]
           end
         end
       end
@@ -86,16 +84,19 @@ module Lossifier
       verb_loop('Clustering', @verbose) do
         cs = clusters
         sizes = cs.map(&:size)
-        sym_expand_len = @cfg.rules.inject(Hash.new(0)) do |lens, (lhs,_)|
-          lens[lhs] = @cfg.expand(lhs).size
-          lens
-        end
 
         break if sizes.count(1) == cs.size
 
+        rule_expand_lengths = Hash[
+          @cfg.rules.map { |lhs, _| [lhs, @cfg.expand(lhs).size] }
+        ]
+
+        # Replace each cluster with the nonterminal in the cluster with the
+        # shortest rule.  This ensures that we avoid creating cycles as a
+        # result of replacement.
         cs.each do |cluster|
           counts = @cfg.counts
-          repl_sym = cluster.min_by { |nonterm| sym_expand_len[nonterm] }
+          repl_sym = cluster.min_by { |nonterm| rule_expand_lengths[nonterm] }
 
           cluster.each do |find_sym|
             next if find_sym == repl_sym

### 2012-12-08 src/main.rb

@@ -22,56 +22,56 @@ ALGORITHMS = {
 }
 
 # Default option values.
-options = OpenStruct.new
-options.algorithms = ALGORITHMS.values
-options.expand = false
-options.print = false
-options.plot = false
-options.reduce = false
-options.verbose = false
-options.analysis = false
-options.thresh = 0.4
+@options = OpenStruct.new
+@options.algorithms = ALGORITHMS.values
+@options.analysis = false
+@options.expand = false
+@options.print = false
+@options.plot = false
+@options.reduce = false
+@options.verbose = false
+@options.thresh = 0.4
 
 OptionParser.new do |opts|
   opts.banner = USAGE
 
-  opts.on("--t N", Float, "Threshold level") do |t|
-    options.thresh = t
-  end
-
   opts.on("-a", "--[no-]analysis", "Perform final analysis") do |a|
-    options.analysis = a
+    @options.analysis = a
   end
 
   opts.on("-e", "--[no-]expand", "Print expansion") do |e|
-    options.expand = e
+    @options.expand = e
   end
 
   opts.on("-g", "--[no-]print-grammar", "Print grammar") do |g|
-    options.print = g
+    @options.print = g
   end
 
   opts.on("-p", "--[no-]plot-grammar", "Plot grammar") do |p|
-    options.plot = p
+    @options.plot = p
+  end
+
+  opts.on("-r", "--[no-]reduce", "Apply grammar reductions") do |r|
+    @options.reduce = r
+  end
+
+  opts.on("-t", "--threshold N", Float, "Similarity threshold level") do |t|
+    @options.thresh = t
   end
 
   opts.on("-l", "--lossifiers algo1,algo2,algon", Array,
           "The lossifier algorithms to use") do |algos|
     if algos == [ 'none' ]
-      options.algorithms = []
+      @options.algorithms = []
       next
     end
 
     raise OptionParser::InvalidArgument unless (algos - ALGORITHMS.keys).empty?
-    options.algorithms = algos.uniq.map { |algo| ALGORITHMS[algo] }
-  end
-
-  opts.on("-r", "--[no-]reduce", "Apply grammar reductions") do |r|
-    options.reduce = r
+    @options.algorithms = algos.uniq.map { |algo| ALGORITHMS[algo] }
   end
 
[…]

### 2012-12-08 src/util.rb

@@ -1,0 +1,9 @@
+#
+# util.rb - Everybody loves util.
+#
+
+class String
+  def indent(n, char=' ')
+    (char * n) + gsub(/(\n+)/) { $1 + (char * n) }
+  end
+end

### 2012-12-08 src/fidelity.rb
Prettifying

@@ -45,14 +45,13 @@ end
 # Perform some statistical analysis on a string.  If a second string is
 # supplied as an argument, we compare `s' against it.
 #
-def fanalyze(s, s2=nil)
-  puts "Final String Statistics"
-  puts ("Total Length:      %d" % [s.size])
-  puts ("Distance:          %f" %
-        [Levenshtein.distance(s, s2).to_f / s2.size.to_f]) if s2 != nil
-  puts ("Avg Word Length:   %f" % [avgwordlen(s)])
-  puts ("Entropy:           %d" % [entropy(s)])
-  (1..3).map do |i|
-    puts ("%d-grams:\n%s" % [i, ngrams(s, i).indent(4)])
+def fanalyze(s1, s2=nil)
+  res =  "Total Length:      %d\n" % [s1.size]
+  res += "Distance:          %f\n" %
+         [Levenshtein.distance(s1, s2).to_f / s2.size.to_f] if s2 != nil
+  res += "Avg Word Length:   %f\n" % [avgwordlen(s1)]
+  res += "Entropy:           %d\n" % [entropy(s1)]
+  (1..3).inject(res) do |res, i|
+    res + "%d-grams:\n%s\n" % [i, ngrams(s1, i).indent(4)]
   end
 end

### 2012-12-08 src/main.rb

@@ -85,12 +85,14 @@ if ARGV[0].nil?
   exit
 end
 
+#
+# Print out a CFG based on the supplied options.
+#
 def output_cfg(title, time, cfg)
   time = '[' + time.to_s + 's]'
   puts title + '-' * (79 - (title.size + time.size)) + time + "\n\n"
 
-  puts cfg if @options.print
-  puts "\n"
+  puts cfg, "\n" if @options.print
 
   if @options.analysis
     puts ("Rules:             %d" % [cfg.rules.size])
@@ -98,12 +100,26 @@ def output_cfg(title, time, cfg)
     puts "\n"
   end
 
-  puts cfg.expand if @options.expand
-  puts "\n"
+  puts cfg.expand, "\n" if @options.expand
 
   plot_cfg cfg, @fprefix + '-' + title if @options.plot
 end
 
+#
+# Print out fidelity analysis.
+#
+def output_stats(title, s1, s2=nil)
+  return unless @options.analysis
+
+  title += '-Stats'
+  puts title + '-' * (79 - title.size) + "\n\n"
+
+  puts fanalyze(s1, s2), "\n"
+end
+
+#
+# Time a computation.
+#
 def time
   t1 = Time.now
   res = yield
@@ -116,7 +132,7 @@ str = File.read(ARGV[0])
 
 icfg, t = time { Sequitur.new(str).run }
 output_cfg 'Sequitur', t, icfg
-fanalyze str if @options.analysis
+output_stats 'Sequitur', str
 
 @options.algorithms.each do |algo|
   lcfg, t = time { algo.new(icfg, @options.verbose, @options.thresh).run }
@@ -124,8 +140,8 @@ fanalyze str if @options.analysis
 
   if @options.reduce
     rlcfg, t = time { Reducer.new(lcfg, @options.verbose).run }
-    output_cfg algo.name + "-Reduced", t, rlcfg
+    output_cfg algo.name + '-Reduced', t, rlcfg
   end
 
-  fanalyze lcfg.expand, str if @options.analysis
+  output_stats algo.name, lcfg.expand, str
 end

### 2012-12-08 src/main.rb
Require rubygems in main.rb

@@ -4,6 +4,8 @@
 # main.rb - Main executable for running experiments.
 #
 
+require 'rubygems'
+
 require 'optparse'
 require 'ostruct'
 

### 2012-12-09 tex/paper.tex
added script for running gen-n tests, more work on results section

@@
 
-\begin{algorithm}
+\begin{algorithm}[h]
 \caption{Similarity Lossifier Algorithm}
@@
       \If{$lsim(str_{1},str_{2})<\epsilon$}
-        \State replace the longer var with the shorter throughout $g$
+        \State replace the longer var with the smaller throughout $g$
         \State break
@@
 parallel. This would avoid many of the repeat pair-comparisons done in
-\emph{Similarity}. In other words, the algorithm below makes puts 
+\emph{Similarity}. In other words, the algorithm below puts 
 similar variables
@@
 
-\begin{algorithm}
+\begin{algorithm}[h]
 \caption{Cluster Lossifier Algorithm}
@@
     \For{$var_1 \rightarrow rhs_1 \in g$}
-      \State blah
+      \For{$c \in clusters$}
+        \State $(var_2 \rightarrow rhs_2) \gets $ first variable in $c$
+        \State $str_{i} \gets expand(rhs_i)$
+        \If{$lsim(str_{1},str_{2})<\epsilon$}
+          \State add $var_1$ to $c$
+          \State break
+        \EndIf
+      \EndFor
+      \If{$var_1$ hasn't been added to a cluster}
+        \State Add $var_1$ to a new cluster in $clusters$
+      \EndIf
+    \EndFor
+    \For{Cluster $c \in clusters$}
+      \State replace all $var \in c$ with the smallest.
     \EndFor
@@
 \end{algorithm}
+
+Both of the algorithms depend on the concept of the "smallest / smaller"
+variable. We can define the size of a variable to be the length of
+the expansion of its rhs. In this way, if we think of our CFG's as a DAG, then
+variables higher up on the DAG will always have larger expansions than
+those lower down, and thus will have a larger size. By always replacing
+with the smallest variable, we avoid creating cyclic references in our CFG.
+We will prove this fact in the next section.
 
@@
 \subsection{Illustrative Examples}
-It is useful to look at how the algorithm functions on a small piece of
+It is useful to look at how the algorithms function on a small piece of
 data with both plenty of hierarchical structure, and also some messiness
@@
 Disregarding its meaning, it consists of four very similar equations of the
-form $a + b = ?$, and on a smaller scale each term consists of
-3 repetitions of approximately the same string of either $123$ or $321$.
+form $a + b = ?$, and on a smaller scale each summand consists of
+3 repetitions of approximately the same string of either $123,$ or $321,$.
 
+When comparing grammars, we chose to define the \emph{size} of a grammar
+as the sum of the lengths of the rhs's of the production rules.
 Sequitur yields a relatively complicated inferred grammar with 13 rules and
@@
 texture to the data by changing 1/30 of the characters uniformly
-to `*'. Call these test-cases \emph{rep-n}
+at random to `*'. Call these test-cases \emph{rep-n}
 Another set of test data was constructed from the first n characters of the
@@
 
-\begin{table}[t]
+Below are the results on running Sequitur and Cluster, without
+the grammar reductions, on the rep-n datasets. Time is the wall clock
+running as measured on a 2.26 Ghz Intel Core 2 Duo Macbook Pro.
+Ratio is the relative size of the grammar produced compared with
+Sequitur, where the size of the grammar is the sum of the lengths of the
+rhs's of the rules. Levdistance is the levenshtein distance between
+the output of the lossified grammar and the original input string, it
+is a kind of measure of the fidelity of our algorithms but doesn't
+take into account how well patterns were preserved.
+
+\begin{table}[h]
 \begin{tabular}{l|l|l|l|l|l|l}
@@
 rep-6400  & 3.978 & 258/448  & 0.706 & 0.203 & 256/448  & 0.869\\
-rep-12800
+rep-12800 & & & & & & \\
 rep-25600 & 135.0 & 900/1482 & 0.829 & 2.906 & 898/1482 & 0.725\\
 \end{tabular}
-\caption{$\epsilon = .4$}
+\caption{rep-n: $\epsilon = .4$}
 \end{table}
 
-\begin{table}[t]
+\begin{table}[h]
 \begin{tabular}{l|l|l|l|l|l|l}
@@
 rep-6400  & 3.942 & 280/448  & 0.387 & 0.295 & 274/448  & 0.543\\
-rep-12800  
+rep-12800 & & & & & & \\
 rep-25600 & 124.5 & 922/1482 & 0.565 & 2.597 & 926/1482 & 0.437\\
 \end{tabular}
-\caption{$\epsilon = .2$}
+\caption{rep-n: $\epsilon = .2$}
 \end{table}
 
-TODO: Fill in results for $\epsilon = .1$
+\begin{table}[h]
+\begin{tabular}{l|l|l|l|l|l|l}
+Dataset & S-T(ime) & S-R(atio) & S-L(evdistance) & C-T & C-R & C-L \\
+\hline
+rep-100   & & /34    & & & /34    & \\
+rep-400   & & /76    & & & /76    & \\
+rep-1600  & & /194  & & & /194  & \\
+rep-6400  & & /448  & & & /448  & \\
+rep-12800 & & & & & & \\
+rep-25600 & & /1482 & & & /1482 & \\
+\end{tabular}
+\caption{$rep-n: \epsilon = .1$}
+\end{table}
 
@@
 
-TODO: Analogous for genesis
-> Growth rate for similarity is far worse for genesis
+Repeating the above tests on the \emph{gen-n} datasets yields the following:
+
+\begin{table}[t]
+\begin{tabular}{l|l|l|l|l|l|l}
+Dataset & S-T(ime) & S-R(atio) & S-L(evdistance) & C-T & C-R & C-L \\
+\hline
+gen-100   & & /& & & /& \\
+gen-400   & & /& & & /& \\
+gen-1600  & & /& & & /& \\
+gen-6400  & & /& & & /& \\
+gen-12800 & & /& & & /& \\
+gen-25600 & & /& & & /& \\
+\end{tabular}
+\caption{$gen-n: \epsilon = .4$}
+\end{table}
+
+\begin{table}[t]
+\begin{tabular}{l|l|l|l|l|l|l}
+Dataset & S-T(ime) & S-R(atio) & S-L(evdistance) & C-T & C-R & C-L \\
+\hline
+gen-100   & & /& & & /& \\
+gen-400   & & /& & & /& \\
+gen-1600  & & /& & & /& \\
+gen-6400  & & /& & & /& \\
+gen-12800 & & /& & & /& \\
+gen-25600 & & /& & & /& \\
+\end{tabular}
+\caption{$gen-n: \epsilon = .2$}
+\end{table}
+
+\begin{table}[t]
+\begin{tabular}{l|l|l|l|l|l|l}
+Dataset & S-T(ime) & S-R(atio) & S-L(evdistance) & C-T & C-R & C-L \\
+\hline
+gen-100   & & /& & & /& \\
+gen-400   & & /& & & /& \\
+gen-1600  & & /& & & /& \\
+gen-6400  & & /& & & /& \\
+gen-12800 & & /& & & /& \\
+gen-25600 & & /& & & /& \\
+\end{tabular}
+\caption{$gen-n: \epsilon = .1$}
+\end{table}
 
@@
 \subsection{Grammar Reduction Impact}
-benchmark two extreme test cases again, with and without reducer
-see how much reducer helps (it shouldn't help too much)
-notice how slow it is
+Although the grammars generated by Sequitur are irreducible,
+our lossifiers produce grammars which could potentially be
+compressed further (losslessly) using Reduction Rules without
+changing the output string. To test the cost and
+effectiveness of applying reduction rules, we looked at the
+rep-n and gen-n test cases again, this time focusing on the
+clustering algorithm for $\epsilon = .2$.
+
+\begin{table}[t]
+\begin{tabular}{l|l|l|l|l}
+Dataset & S-T(ime) & S-R(atio) & C-T & C-R \\
+\hline
+rep-100   & & /& & /\\
+rep-400   & & /& & /\\
+rep-1600  & & /& & /\\
+rep-6400  & & /& & /\\
+rep-12800 & & /& & /\\
+rep-25600 & & /& & /\\
+\end{tabular}
+\caption{$rep-n: \epsilon = .2$}
+\end{table}
+
+\begin{table}[t]
+\begin{tabular}{l|l|l|l|l}
+Dataset & S-T(ime) & S-R(atio) & C-T & C-R \\
+\hline
+gen-100   & & /& & /\\
+gen-400   & & /& & /\\
+gen-1600  & & /& & /\\
+gen-6400  & & /& & /\\
+gen-12800 & & /& & /\\
+gen-25600 & & /& & /\\
+\end{tabular}
+\caption{$gen-n: \epsilon = .2$}
+\end{table}
 
@@
 \section{Conclusions}
+\subsection{Algorithm Tradeoffs}
+Runtimes for each algorithm
+Considerations in choosing epsilon
 \subsection{Ratios and Fidelity}
 \subsection{Future Work}
-> Better ways to set epsilon to maximize value / cost for different size and types
-> Better ways to quantify preservation of hierarchies
-> How to find near `self-similarities' that aren't captured by a rule
-> Iterating lossify and reduce in a cycle
+* Better ways to set epsilon to maximize value / cost for different size and types
+* Better ways to quantify preservation of hierarchies
+* How to find near `self-similarities' that aren't captured by a rule
+* Iterating lossify and reduce in a cycle
 

### 2012-12-09 src/finaltests/runTest2.sh

@@ -1,0 +1,21 @@
+#!/bin/bash
+s[1]=100
+s[2]=400
+s[3]=1600
+s[4]=6400
+s[5]=12800
+s[6]=25600
+t[1]=0.1
+t[2]=0.2
+t[3]=0.4
+to[1]=1
+to[2]=2
+to[3]=4
+for i in {1..6}
+do
+  for j in {1..3}
+  do
+  echo "ruby ../main.rb -ae -t ${t[j]} gen${s[i]}.txt > results/gen${s[i]}_res_e${to[j]}.txt"
+  ruby ../main.rb -ae -t ${t[j]} gen${s[i]}.txt > results/gen${s[i]}_res_e${to[j]}.txt
+  done
+done

### 2012-12-10 tex/paper.tex
OMG typing in the results into the latex tables

@@
 \usepackage[pdftex]{graphicx}
+\usepackage[lofdepth,lotdepth]{subfig}
 \usepackage{ifthen}
@@
 is a kind of measure of the fidelity of our algorithms but doesn't
-take into account how well patterns were preserved.
+take into account how well patterns were preserved. For the rep-n tests,
+proportion is the proportion of the final string produced that was
+`*'. The closer to 1/30~0.033, the more accurate the lossifier was
+in recreating the original texture of the input.
 
-\begin{table}[h]
-\begin{tabular}{l|l|l|l|l|l|l}
-Dataset & S-T(ime) & S-R(atio) & S-L(evdistance) & C-T & C-R & C-L \\
+The columns of the data are prefixed by S or T depending on whether
+they are measurements for sequitur or cluster respectively.
+
+\begin{table}[t]
+\subfloat[$\epsilon = .4$]{
+\begin{tabular}{l|l|l|l|l||l|l|l|l}
+Dataset & S-T(ime) & S-R(atio) & S-L(evdistance) & S-P(roportion) & C-T & C-R & C-L & C-P \\
 \hline
-rep-100   & 0.006 & 18/34    & 0.713 & 0.002 & 20/34    & 0.535\\
-rep-400   & 0.028 & 45/76    & 0.269 & 0.009 & 41/76    & 0.743\\
-rep-1600  & 0.355 & 103/194  & 0.873 & 0.042 & 103/194  & 0.804\\
-rep-6400  & 3.978 & 258/448  & 0.706 & 0.203 & 256/448  & 0.869\\
-rep-12800 & & & & & & \\
-rep-25600 & 135.0 & 900/1482 & 0.829 & 2.906 & 898/1482 & 0.725\\
+rep-100   & 0.006 & 18/34    & 0.713 & 0.07 & 0.002 & 20/34    & 0.535 & 0.04\\
+rep-400   & 0.028 & 45/76    & 0.269 & 0.01 & 0.009 & 41/76    & 0.743 & 0.04\\
+rep-1600  & 0.355 & 103/194  & 0.873 & 0.03 & 0.042 & 103/194  & 0.804 & 0.02\\
+rep-6400  & 3.978 & 258/448  & 0.706 & 0.005& 0.203 & 256/448  & 0.869 & 0.50\\
+rep-12800 & 44.63 & 604/1064 & 0.791 & 0.01 & 1.136 & 600/1064 & 0.856 & 0.01\\
+rep-25600 & 135.0 & 900/1482 & 0.829 & 0.005& 2.906 & 898/1482 & 0.725 & 0.25\\
 \end{tabular}
-\caption{rep-n: $\epsilon = .4$}
+}
+
+\subfloat[$\epsilon = .2$]{
+\begin{tabular}{l|l|l|l|l||l|l|l|l}
+Dataset & S-T(ime) & S-R(atio) & S-L(evdistance) & S-P(roportion) & C-T & C-R & C-L & C-P \\
+\hline
+rep-100   & 0.002 & 30/34    & 0.040 & 0.05 & 0.002 & 30/34    & 0.040 & 0.05\\
+rep-400   & 0.019 & 60/76    & 0.035 & 0.02 & 0.009 & 60/76    & 0.035 & 0.02\\
+rep-1600  & 0.350 & 123/194  & 0.355 & 0.03 & 0.055 & 123/194  & 0.206 & 0.01\\
+rep-6400  & 3.942 & 280/448  & 0.387 & 0.08 & 0.295 & 274/448  & 0.543 & 0.09\\
+rep-12800 & 42.21 & 638/1064 & 0.491 & 0.03 & 1.105 & 632/1064 & 0.457 & 0.01\\
+rep-25600 & 124.5 & 922/1482 & 0.565 & 0.12 & 2.597 & 926/1482 & 0.437 & 0.007\\
+\end{tabular}
+}
+\caption{rep-n Tests}
 \end{table}
 
-\begin{table}[h]
-\begin{tabular}{l|l|l|l|l|l|l}
-Dataset & S-T(ime) & S-R(atio) & S-L(evdistance) & C-T & C-R & C-L \\
-\hline
-rep-100   & 0.002 & 30/34    & 0.040 & 0.002 & 30/34    & 0.040\\
-rep-400   & 0.019 & 60/76    & 0.035 & 0.009 & 60/76    & 0.035\\
-rep-1600  & 0.350 & 123/194  & 0.355 & 0.055 & 123/194  & 0.206\\
-rep-6400  & 3.942 & 280/448  & 0.387 & 0.295 & 274/448  & 0.543\\
-rep-12800 & & & & & & \\
-rep-25600 & 124.5 & 922/1482 & 0.565 & 2.597 & 926/1482 & 0.437\\
-\end{tabular}
-\caption{rep-n: $\epsilon = .2$}
-\end{table}
-
-\begin{table}[h]
-\begin{tabular}{l|l|l|l|l|l|l}
-Dataset & S-T(ime) & S-R(atio) & S-L(evdistance) & C-T & C-R & C-L \\
-\hline
-rep-100   & & /34    & & & /34    & \\
-rep-400   & & /76    & & & /76    & \\
-rep-1600  & & /194  & & & /194  & \\
-rep-6400  & & /448  & & & /448  & \\
-rep-12800 & & & & & & \\
-rep-25600 & & /1482 & & & /1482 & \\
-\end{tabular}
-\caption{$rep-n: \epsilon = .1$}
-\end{table}
-
-TODO: Runtime graphs and regression. The exponents are about 1.8 and 1.3
-respectively, note that the rulesize of sequitur grows at about sqrt(n)
-
-Qualitatively the output strings generally are strings of a's with interspersed
-`*' as expected, but the length of the final output varies widely.
-
-Things to mention: Compression ratios and fidelity comparable for similarity
-and cluster. Cluster has much better growth rate. Smaller epsilon provide
-much better value/cost ratio as the input files get bigger for highly
-repetitive files.
+Runtime Comparison, esp. w.r.t num of rules.
+Quality Comparison.
+Epsilon Comparison.
 
@@
 \begin{table}[t]
-\begin{tabular}{l|l|l|l|l|l|l}
+\subfloat[$\epsilon = .4$]{
+\begin{tabular}{l|l|l|l||l|l|l}
 Dataset & S-T(ime) & S-R(atio) & S-L(evdistance) & C-T & C-R & C-L \\
 \hline
-gen-100   & & /& & & /& \\
-gen-400   & & /& & & /& \\
-gen-1600  & & /& & & /& \\
-gen-6400  & & /& & & /& \\
-gen-12800 & & /& & & /& \\
-gen-25600 & & /& & & /& \\
+gen-100   & 0.006 & 87/89    & 0.04 &0.006 & 87/89    & 0.04\\
+gen-400   & 0.175 & 224/258  & 0.12 &0.069 & 219/258  & 0.19\\
+gen-1600  & 4.863 & 539/697  & 0.31 &0.732 & 529/697  & 0.33\\
+gen-6400  & 216.9 & 1740/2348& 0.41 &9.127 & 1694/2348& 0.43\\
+gen-12800 & 1812  & 3296/4523& 0.40 &29.17 & 3181/4523& 0.44\\
+gen-25600 & 12989 & 5823/8060& 0.43 &176.7 & 5540/8060& 0.50\\
 \end{tabular}
-\caption{$gen-n: \epsilon = .4$}
-\end{table}
+}
 
-\begin{table}[t]
-\begin{tabular}{l|l|l|l|l|l|l}
+\subfloat[$\epsilon = .2$]{
+\begin{tabular}{l|l|l|l||l|l|l}
 Dataset & S-T(ime) & S-R(atio) & S-L(evdistance) & C-T & C-R & C-L \\
 \hline
-gen-100   & & /& & & /& \\
-gen-400   & & /& & & /& \\
-gen-1600  & & /& & & /& \\
-gen-6400  & & /& & & /& \\
-gen-12800 & & /& & & /& \\
-gen-25600 & & /& & & /& \\
+gen-100   & 0.004 & 89/89    & 0.00 &0.002 & 89/89    & 0.00 \\
+gen-400   & 0.085 & 245/258  & 0.03 &0.053 & 245/258  & 0.03 \\
+gen-1600  & 1.292 & 664/697  & 0.03 &0.874 & 664/697  & 0.03 \\
+gen-6400  & 73.83 & 2173/2348& 0.05 &14.02 & 2171/2348& 0.05\\
+gen-12800 & 549.2 & 4197/4253& 0.05 &45.64 & 4186/4253& 0.05\\
+gen-25600 & 3939  & 7405/8060& 0.06 &224.9 & 7374/8060& 0.06\\
 \end{tabular}
-\caption{$gen-n: \epsilon = .2$}
-\end{table}
-
-\begin{table}[t]
-\begin{tabular}{l|l|l|l|l|l|l}
-Dataset & S-T(ime) & S-R(atio) & S-L(evdistance) & C-T & C-R & C-L \\
-\hline
-gen-100   & & /& & & /& \\
-gen-400   & & /& & & /& \\
-gen-1600  & & /& & & /& \\
-gen-6400  & & /& & & /& \\
-gen-12800 & & /& & & /& \\
-gen-25600 & & /& & & /& \\
-\end{tabular}
-\caption{$gen-n: \epsilon = .1$}
+}
+\caption{gen-n}
 \end{table}
@@
 \begin{table}[t]
+\subfloat[rep-n tests]{
 \begin{tabular}{l|l|l|l|l}
@@
 rep-12800 & & /& & /\\
-rep-25600 & & /& & /\\
 \end{tabular}
-\caption{$rep-n: \epsilon = .2$}
-\end{table}
+}
 
-\begin{table}[t]
+\subfloat[gen-n tests]{
 \begin{tabular}{l|l|l|l|l}
@@
 gen-12800 & & /& & /\\
-gen-25600 & & /& & /\\
 \end{tabular}
-\caption{$gen-n: \epsilon = .2$}
+}
+\caption{Reducer Tests for $\epsilon = .2$}
 \end{table}

### 2012-12-10 src/finaltests/runTest3.sh

@@ -1,0 +1,12 @@
+#!/bin/bash
+s[1]=100
+s[2]=400
+s[3]=1600
+s[4]=6400
+for i in {1..4}
+do
+  echo "ruby ../main.rb -ar -t 0.2 e_rep${s[i]}.txt > results/rep${s[i]}_redres_e2.txt"
+  ruby ../main.rb -ar -t 0.2 e_rep${s[i]}.txt > results/rep${s[i]}_redres_e2.txt
+  echo "ruby ../main.rb -ar -t 0.2 gen${s[i]}.txt > results/gen${s[i]}_redres_e2.txt"
+  ruby ../main.rb -ar -t 0.2 gen${s[i]}.txt > results/gen${s[i]}_redres_e2.txt
+done

### 2012-12-10 tex/paper.tex
finished similarity vs cluster and reduction rule analysis, language needs polishing

@@
 
+\subsubsection{Rep-n results}
+
 Below are the results on running Sequitur and Cluster, without
@@
 
-\begin{table}[t]
+\begin{table}[p]
 \subfloat[$\epsilon = .4$]{
@@
 
-Runtime Comparison, esp. w.r.t num of rules.
-Quality Comparison.
-Epsilon Comparison.
+The compression ratios achieved for both algorithms are roughly equal for the same
+value for $\epsilon$, so it is fair to compare the algorithms for the same $\epsilon$.
+
+From plotting the data, we see that runtime of the algorithms are roughly polynomial
+in the size of the data, so we can perform a linear regression on the log-log transform
+of the data to find the exponents $\alpha$ for the runtime $r(n)=n^\alpha$ where $n$ is the
+length of the input string. In the chart below, all of the linear regressions had
+$R^2$ values of over $0.98$.
+
+\begin{tabular}{l|l|l|l|l}
+Data     & Sim: $\epsilon=.4$ & Sim: $\epsilon=.2$ & Clust: $\epsilon = .4$ & Clust: $\epsilon = .2$ \\
+Exponent & $1.859$ & $2.026$ & $1.306$ & $1.305$
+\end{tabular}
+
+For both values of $\epsilon$, 
+The clustering algorithm is nearly linear, while the similarity algorithm is roughly
+quadratic. On this dataset, the leveshtein distances were similar for both algorithms
+but fairly high since the length of the final string could change drastically.
+The proportion of '*'s in the input was usually the right order of magnitude,
+with both algorithms making big mistakes occassionally. In most cases,
+the two algorithms maintained similarly acceptable fidelity levels in recreating
+the texture of the original rep-n input file.
+
+Setting higher values for $\epsilon$ had diminishing returns for compression
+for larger and larger file sizes, at the cost of drastically lowering the fidelity
+of the output.
+
+\subsubsection{Gen-n results}
 
@@
 
-Thus, we see that similarity provides no consistent advantage over cluster
-and has much less scalable performance, with an estimated running time of 
-BLAH. For the rest of the paper we will only test Cluster.
+Just as with the rep-n data, log-log regression tests yield the following
+exponents for the runtime on the gen-n data.
+
+\begin{tabular}{l|l|l|l|l}
+Data     & Sim: $\epsilon=.4$ & Sim: $\epsilon=.2$ & Clust: $\epsilon = .4$ & Clust: $\epsilon = .2$ \\
+Exponent & $2.629$ & $2.493$ & $1.814$ & $2.061$
+\end{tabular}
+
+Again, the two algorithms have almost identical compression ratios and
+their fidelity is about the same both subjectively and based on levenshtein
+distance.
+
+INCLUDE SAMPLE TEXT
+
+The runtimes are noticeably worse than for the rep-n dataset. The
+similarity algorithm runs in roughly $n^{2.5}$ time while cluster
+runs in roughly $n^2$ time.
+
+Changing epsilon in this case had a large effect on both the compression
+ratio and the fidelity of the final output string.
+
+\subsubsection{Summary}
+
+Both of the algorithms were able to achieve great additional compression
+ratios on top of \Sequitur of up to about 40 percent for the highly
+structured rep-n data, and 30 percent for the more loosely structured
+gen-n data without using the grammar reducer. The fidelity between
+the original and lossified strings wasn't great as measured by
+levenshtein distance, but the texture in both the rep-n and
+gen-n datasets was preserved well. Decreasing the $\epsilon$ threshold
+improved fidelity greatly for larger very structured data at little
+cost to the compression ratio, but hurt the compression ratio
+a lot for less structured data.
+
+The differences between the two algorithms lie almost entirely in their
+runtime. Both algorithms perform much worse on the less structued
+gen-n data, we believe because simplifications are made more incrementally
+and the grammar does not get dramatically simpler after a few iterations
+of our algorithm like it does for the rep-n data.
+Unfortunately the Similarity algorithm tends to consistently incur
+an overhead of about $O(\sqrt{n})$ on top of the runtime for the
+Cluster algorithm. Because of this we dismiss the Similarity algorithm
+as unscalable and do not test it further.
 
@@
 rep-n and gen-n test cases again, this time focusing on the
-clustering algorithm for $\epsilon = .2$.
+clustering algorithm for $\epsilon = .2$. This time we give the ratio
+on top of the simplified grammar produced by the clustering algorithm,
+and the time taken to run just the reduction rules.
 
-\begin{table}[t]
+\begin{table}[p]
 \subfloat[rep-n tests]{
-\begin{tabular}{l|l|l|l|l}
-Dataset & S-T(ime) & S-R(atio) & C-T & C-R \\
+\begin{tabular}{l|l|l}
+Dataset & Time & Ratio \\
 \hline
-rep-100   & & /& & /\\
-rep-400   & & /& & /\\
-rep-1600  & & /& & /\\
-rep-6400  & & /& & /\\
-rep-12800 & & /& & /\\
+rep-100   & 0.008 & 29/30  \\
+rep-400   & 0.028 & 56/60  \\
+rep-1600  & 0.132 & 96/123 \\
+rep-6400  & 1.145 & 153/186\\
+rep-12800 & 19.43 & 403/632\\
 \end{tabular}
 }
-
 \subfloat[gen-n tests]{
-\begin{tabular}{l|l|l|l|l}
-Dataset & S-T(ime) & S-R(atio) & C-T & C-R \\
+\begin{tabular}{l|l|l}
+Dataset & Time & Ratio \\
 \hline
-gen-100   & & /& & /\\
-gen-400   & & /& & /\\
-gen-1600  & & /& & /\\
-gen-6400  & & /& & /\\
-gen-12800 & & /& & /\\
+gen-100   & 0.016 & 89/89\\
+gen-400   & 0.126 & 245/245\\
+gen-1600  & 4.833 & 660/664\\
+gen-6400  & 308.7 & 2139/2171\\
 \end{tabular}
@@
 \end{table}
+
+When plotted on a log-log scale, the runtime for our straightforward 
+implementation of the reduction
+rules appears to be superpolynomial. While modest additional gains of about 30 percent
+are possible for highly self-similar data, very few additional gains are
+possible for more diverse datasets such as gen-n. This appears to be because
+the only simplifications that are possible on diverse datasets don't lead to
+the wide-ranging repetitive substitutions which allow reduction rules to be allowed.
+The simplifications in gen-n tend to be more local.
+
+Because we were unable to get reduction rules to scale to larger datasets,
+we will not consider them further.
 

### 2012-12-10 tex/paper.tex
tried to do a runtime analysis, more difficult to make fit data than I thought

@@
 
-\subsection{Mathematical Analysis}
-Runtime Analysis
-No-loops proof
+\subsection{Runtime Estimates}
+Given an input of length $n$, let $r$ be the number of rules generated
+by Sequitur from the input. Let $R$ be the total number of variable
+replacements we perform. For each replacement, we had to have searched though
+up to all pairs of variables, and for each pair of variables expanding
+them and calculating the levenshtein distance could take time linear in $n$.
+Thus the runtime is $\boxed{O(R*r^2*n)}$.
+
+From the \Sequitur paper \cite{sequitur} we
+know that the number of rules in the grammar generated by Sequitur
+is between $O(\sqrt{n})$ and $O(n)$. Thus the worst case runtime on the similarity
+algorithm is $\boxed{O(n^4)}$. This is because we could have up to $r$ replacements.
+However, in practice we will see that the runtime is much better
+than $O(n^4)$. This is because we found $r$ for our test data to be slightly sublinear
+even for complicated english text, and we found that we didn't quite make a linear
+number of replacements $R$ with respect to $r$.
+so our optimistic runtime could be anywhere 
+from $O(n^4)$ to $O(n^{2})$ depending on how $r$ grows.
+
+The inner loops of the Cluster are similar to those of the Similarity algorithm.
+We iterate the entire clustering and replacement procedure $C$ times. Going through
+each variable-cluster pair takes up to $O(r^2)$ time and calculating the levenshtein
+distance takes $O(n)$ time. The time taken to actually perform the replacements each
+iteration is absorbed into the $O(r^2)$ term. So, again we have a runtime of
+$\boxed{O(C*r^2*n)}$. The big gain over the similarity algorithm is that $C$ turns out
+to be nearly constant, since there are very few huge chains of reductions that are
+only possible after other reductions have been performed, an in fact we are able to 
+perform a very large number of the remaining reductions each time we perform an iteration
+of the clustering algorithm. Thus since $C$ is nearly constant, the optimistic runtime
+could be anywhere from $O(n^3)$ to $O(n^2)$.
+
+CANT THINK OF PROOF OF NOLOOPS
 
@@
 \begin{verbatim}
+Grammar:
 ~[*] => ~[BC]--~[BC]?
@@
 -----------------------
+Generated String:
 1231212 + 32323321 = ???1231212 + 32323321 = ???--(cont.)
@@
 \subsection{Larger Tests}
-Final compression ratios on various benchmark data
-ASCII ART =p
-Our Code
+Genesis-256000
+Obfuscated C Code
 

### 2012-12-10 tex/paper.tex
some bullet points for conclusions

@@
 these structures, we built
+Music Final Composition: Dec 10
 our system on top of the existing \Sequitur grammar inference algorithm. 
@@
 Genesis-256000
-Obfuscated C Code
+Obfuscated C Code (looks really cool!)
 
 \section{Conclusions}
-\subsection{Algorithm Tradeoffs}
-Runtimes for each algorithm
-Considerations in choosing epsilon
+\subsection{Summary of Results}
+* General concept of simplifying grammars, two proposed algorithms
+* reasonably efficient implementation with runtime better than expected
+* compared algorithms, evaluated effectiveness of reducer
+* looked at effect of epsilon
+
 \subsection{Ratios and Fidelity}
+
+* How useful are these algorithms??
+* What is the ratio we get
+* What kind of fidelity to we get
+
 \subsection{Future Work}

### 2012-12-11 tex/paper.tex
Moved all results to the end, added big interesting subjective data

@@
 \usepackage{algpseudocode}
+\usepackage{listings}
+\usepackage{hyperref}
+\lstset{basicstyle=\ttfamily\footnotesize,breaklines=true}
 
@@
 \end{abstract}
+
+\tableofcontents
 
@@
 
-\begin{table}[p]
-\subfloat[$\epsilon = .4$]{
-\begin{tabular}{l|l|l|l|l||l|l|l|l}
-Dataset & S-T(ime) & S-R(atio) & S-L(evdistance) & S-P(roportion) & C-T & C-R & C-L & C-P \\
-\hline
-rep-100   & 0.006 & 18/34    & 0.713 & 0.07 & 0.002 & 20/34    & 0.535 & 0.04\\
-rep-400   & 0.028 & 45/76    & 0.269 & 0.01 & 0.009 & 41/76    & 0.743 & 0.04\\
-rep-1600  & 0.355 & 103/194  & 0.873 & 0.03 & 0.042 & 103/194  & 0.804 & 0.02\\
-rep-6400  & 3.978 & 258/448  & 0.706 & 0.005& 0.203 & 256/448  & 0.869 & 0.50\\
-rep-12800 & 44.63 & 604/1064 & 0.791 & 0.01 & 1.136 & 600/1064 & 0.856 & 0.01\\
-rep-25600 & 135.0 & 900/1482 & 0.829 & 0.005& 2.906 & 898/1482 & 0.725 & 0.25\\
-\end{tabular}
-}
-
-\subfloat[$\epsilon = .2$]{
-\begin{tabular}{l|l|l|l|l||l|l|l|l}
-Dataset & S-T(ime) & S-R(atio) & S-L(evdistance) & S-P(roportion) & C-T & C-R & C-L & C-P \\
-\hline
-rep-100   & 0.002 & 30/34    & 0.040 & 0.05 & 0.002 & 30/34    & 0.040 & 0.05\\
-rep-400   & 0.019 & 60/76    & 0.035 & 0.02 & 0.009 & 60/76    & 0.035 & 0.02\\
-rep-1600  & 0.350 & 123/194  & 0.355 & 0.03 & 0.055 & 123/194  & 0.206 & 0.01\\
-rep-6400  & 3.942 & 280/448  & 0.387 & 0.08 & 0.295 & 274/448  & 0.543 & 0.09\\
-rep-12800 & 42.21 & 638/1064 & 0.491 & 0.03 & 1.105 & 632/1064 & 0.457 & 0.01\\
-rep-25600 & 124.5 & 922/1482 & 0.565 & 0.12 & 2.597 & 926/1482 & 0.437 & 0.007\\
-\end{tabular}
-}
-\caption{rep-n Tests}
-\end{table}
-
 The compression ratios achieved for both algorithms are roughly equal for the same
@@
 Repeating the above tests on the \emph{gen-n} datasets yields the following:
-
-\begin{table}[p]
-\subfloat[$\epsilon = .4$]{
-\begin{tabular}{l|l|l|l||l|l|l}
-Dataset & S-T(ime) & S-R(atio) & S-L(evdistance) & C-T & C-R & C-L \\
-\hline
-gen-100   & 0.006 & 87/89    & 0.04 &0.006 & 87/89    & 0.04\\
-gen-400   & 0.175 & 224/258  & 0.12 &0.069 & 219/258  & 0.19\\
-gen-1600  & 4.863 & 539/697  & 0.31 &0.732 & 529/697  & 0.33\\
-gen-6400  & 216.9 & 1740/2348& 0.41 &9.127 & 1694/2348& 0.43\\
-gen-12800 & 1812  & 3296/4523& 0.40 &29.17 & 3181/4523& 0.44\\
-gen-25600 & 12989 & 5823/8060& 0.43 &176.7 & 5540/8060& 0.50\\
-\end{tabular}
-}
-
-\subfloat[$\epsilon = .2$]{
-\begin{tabular}{l|l|l|l||l|l|l}
-Dataset & S-T(ime) & S-R(atio) & S-L(evdistance) & C-T & C-R & C-L \\
-\hline
-gen-100   & 0.004 & 89/89    & 0.00 &0.002 & 89/89    & 0.00 \\
-gen-400   & 0.085 & 245/258  & 0.03 &0.053 & 245/258  & 0.03 \\
-gen-1600  & 1.292 & 664/697  & 0.03 &0.874 & 664/697  & 0.03 \\
-gen-6400  & 73.83 & 2173/2348& 0.05 &14.02 & 2171/2348& 0.05\\
-gen-12800 & 549.2 & 4197/4253& 0.05 &45.64 & 4186/4253& 0.05\\
-gen-25600 & 3939  & 7405/8060& 0.06 &224.9 & 7374/8060& 0.06\\
-\end{tabular}
-}
-\caption{gen-n}
-\end{table}
 
@@
 
-\begin{table}[p]
+When plotted on a log-log scale, the runtime for our straightforward 
+implementation of the reduction
+rules appears to be superpolynomial. While modest additional gains of about 30 percent
+are possible for highly self-similar data, very few additional gains are
+possible for more diverse datasets such as gen-n. This appears to be because
+the only simplifications that are possible on diverse datasets don't lead to
+the wide-ranging repetitive substitutions which allow reduction rules to be allowed.
+The simplifications in gen-n tend to be more local.
+
+Because we were unable to get reduction rules to scale to larger datasets,
+we will not consider them further.
+
+\subsection{Larger Tests}
+
+Though we were ultimately not able to come up with a good way to quantify how much of the
+structure of a text was preserved, we have included the final output strings we obtain
+after running the Clustering Algorithm with $\epsilon = .4$ in the Appendices. We
+believe these tests serve as a good (subjective) summary of the strengths and
+weaknesses of our
+algorithm in compressing data while preserving
+its character.
+
+The first piece of text is from the gen-n dataset again, the first 256000 characters
+of the book of Genesis. As listed in table \ref{seqclus_g_t}, the compressed grammar
+was size 5540 compared with Sequitur's size of 8060. Without knowledge of English 
+spelling, the Cluster algorithm produces output that is hard to read. However,
+key words are clearly recognizable, especially important words such as `God', and
+the overall structure in the book of Genesis is retained.
+
+The second piece of text is one of the entries to this year's International
+Obfuscated C code contest, at \url{http://www.ioccc.org/2012/deckmyn/deckmyn.c} .
+The size of the lossified grammar is 1576 compared with the original size of
+1973. The intricate whitespacing was lost, but the clear delineation into
+sections was retained, key words such as `define' and `for' are very visible,
+and the texture of the obfuscated C is very visible.
+
+\section{Conclusions}
+\subsection{Summary of Results}
+* General concept of simplifying grammars, two proposed algorithms
+* reasonably efficient implementation with runtime better than expected
+* compared algorithms, evaluated effectiveness of reducer
+* looked at effect of epsilon
+
+\subsection{Ratios and Fidelity}
+
+* How useful are these algorithms??
+* What is the ratio we get
+* What kind of fidelity to we get
+
+\subsection{Future Work}
+* Better ways to set epsilon to maximize value / cost for different size and types
+* Better ways to quantify preservation of hierarchies
+* How to find near `self-similarities' that aren't captured by a rule
+* Iterating lossify and reduce in a cycle
+
+\pagebreak
+
+\nocite{*}
+
+\bibliography{sources}
+
+\pagebreak
+
+\appendix
+
+\section{Data}
+
+\begin{table}[h]
+\subfloat[$\epsilon = .4$]{
+\begin{tabular}{l|l|l|l|l||l|l|l|l}
+Dataset & S-T(ime) & S-R(atio) & S-L(evdistance) & S-P(roportion) & C-T & C-R & C-L & C-P \\
+\hline
+rep-100   & 0.006 & 18/34    & 0.713 & 0.07 & 0.002 & 20/34    & 0.535 & 0.04\\
+rep-400   & 0.028 & 45/76    & 0.269 & 0.01 & 0.009 & 41/76    & 0.743 & 0.04\\
+rep-1600  & 0.355 & 103/194  & 0.873 & 0.03 & 0.042 & 103/194  & 0.804 & 0.02\\
+rep-6400  & 3.978 & 258/448  & 0.706 & 0.005& 0.203 & 256/448  & 0.869 & 0.50\\
+rep-12800 & 44.63 & 604/1064 & 0.791 & 0.01 & 1.136 & 600/1064 & 0.856 & 0.01\\
+rep-25600 & 135.0 & 900/1482 & 0.829 & 0.005& 2.906 & 898/1482 & 0.725 & 0.25\\
+\end{tabular}
+}
+
+\subfloat[$\epsilon = .2$]{
+\begin{tabular}{l|l|l|l|l||l|l|l|l}
+Dataset & S-T(ime) & S-R(atio) & S-L(evdistance) & S-P(roportion) & C-T & C-R & C-L & C-P \\
+\hline
+rep-100   & 0.002 & 30/34    & 0.040 & 0.05 & 0.002 & 30/34    & 0.040 & 0.05\\
+rep-400   & 0.019 & 60/76    & 0.035 & 0.02 & 0.009 & 60/76    & 0.035 & 0.02\\
+rep-1600  & 0.350 & 123/194  & 0.355 & 0.03 & 0.055 & 123/194  & 0.206 & 0.01\\
+rep-6400  & 3.942 & 280/448  & 0.387 & 0.08 & 0.295 & 274/448  & 0.543 & 0.09\\
+rep-12800 & 42.21 & 638/1064 & 0.491 & 0.03 & 1.105 & 632/1064 & 0.457 & 0.01\\
+rep-25600 & 124.5 & 922/1482 & 0.565 & 0.12 & 2.597 & 926/1482 & 0.437 & 0.007\\
+\end{tabular}
+}
+\caption{rep-n Tests}
+\label{seqclus_r_t}
+\end{table}
+
+\begin{table}[h]
+\subfloat[$\epsilon = .4$]{
+\begin{tabular}{l|l|l|l||l|l|l}
+Dataset & S-T(ime) & S-R(atio) & S-L(evdistance) & C-T & C-R & C-L \\
+\hline
+gen-100   & 0.006 & 87/89    & 0.04 &0.006 & 87/89    & 0.04\\
+gen-400   & 0.175 & 224/258  & 0.12 &0.069 & 219/258  & 0.19\\
+gen-1600  & 4.863 & 539/697  & 0.31 &0.732 & 529/697  & 0.33\\
+gen-6400  & 216.9 & 1740/2348& 0.41 &9.127 & 1694/2348& 0.43\\
+gen-12800 & 1812  & 3296/4523& 0.40 &29.17 & 3181/4523& 0.44\\
+gen-25600 & 12989 & 5823/8060& 0.43 &176.7 & 5540/8060& 0.50\\
+\end{tabular}
+}
+
+\subfloat[$\epsilon = .2$]{
+\begin{tabular}{l|l|l|l||l|l|l}
+Dataset & S-T(ime) & S-R(atio) & S-L(evdistance) & C-T & C-R & C-L \\
+\hline
+gen-100   & 0.004 & 89/89    & 0.00 &0.002 & 89/89    & 0.00 \\
+gen-400   & 0.085 & 245/258  & 0.03 &0.053 & 245/258  & 0.03 \\
+gen-1600  & 1.292 & 664/697  & 0.03 &0.874 & 664/697  & 0.03 \\
+gen-6400  & 73.83 & 2173/2348& 0.05 &14.02 & 2171/2348& 0.05\\
+gen-12800 & 549.2 & 4197/4253& 0.05 &45.64 & 4186/4253& 0.05\\
+gen-25600 & 3939  & 7405/8060& 0.06 &224.9 & 7374/8060& 0.06\\
+\end{tabular}
+}
+\caption{gen-n Tests}
+\label{seqclus_g_t}
+\end{table}
+
+\begin{table}[h]
 \subfloat[rep-n tests]{
@@
 }
+
 \caption{Reducer Tests for $\epsilon = .2$}
+\label{red_t}
 \end{table}
 
-When plotted on a log-log scale, the runtime for our straightforward 
-implementation of the reduction
-rules appears to be superpolynomial. While modest additional gains of about 30 percent
-are possible for highly self-similar data, very few additional gains are
-possible for more diverse datasets such as gen-n. This appears to be because
-the only simplifications that are possible on diverse datasets don't lead to
-the wide-ranging repetitive substitutions which allow reduction rules to be allowed.
-The simplifications in gen-n tend to be more local.
+\pagebreak
 
-Because we were unable to get reduction rules to scale to larger datasets,
-we will not consider them further.
+\subsection*{Deckmyn.c Input / Output:}
+\subsubsection*{Input:}
 
-\subsection{Larger Tests}
-Genesis-256000
-Obfuscated C Code (looks really cool!)
+\begin{lstlisting}
+#include<stdio.h>
+#define c(C) printf("%c",C)
+#define C(c) ((int*)(C[1]+6))[c]
+main(int c,char    
+*C[])         {(C[c]=C[
+1]+2   )[0]=   c(52*c(\
+'C'+  '4'/4)    );for(c
+=0;  c<491;++   c)for(*
+*C=  C[1]['c'    +c]  =
+0;*   C[0]<8;(    **  C
+)++    )C[1][c+   'c']=
+*(C[  1]+c+'c')+  C[1][
+99+    c]+(C[1    ][**C
++8*c  +99]==32    );  (
+*C)[4]=*C[2]==    75  ?
+*((C[2]+=3)-2    )==70?
+1:0:0;C(0)=C(    1)=c=0
+;while(*C[2]?   C[2][1]
+?*(C[2]+2)?1    :0:0:0)
+{if( *C [2      ]>'w'){
+C(1)=0;C[1]    [2]++;*C
+[2]=0;}else   C(1)+=*C[
+2]==58?2+(    C[2][3]&&
+*(C[2]+3)<   'x'):*C[2]
+=='s'?(C[   2][1]-=48):
+*C[2]>=65  ?3-(*C[2]==\
+'m'?1:0)  :1;C(0)=C(1)>
+C(0)?C(1  ):C(0);c+=3;*
+(C+2)+=3;}printf("  %d\
+          %d\n",        56+8*C(    0),80**(C[3]    ++))
+         ;*C[2]=0       ;C[2]       -=c;*C[3]       =0;
+        while(C[3]      [1,-         1]--){;   for(  **
+        C=0     ;*      *C<          80;(**   C)++)  {C
+       [2]      -=3     **           C[3];   *C[3]   ++
+       =0;      *C[     3]          =**C>=  51||*   *C<
+       18       ||*     *C         %8!=2?0  :255   ;c(1
+       -1       );c     (*C       [3]);for(       (*C)[
+       1]      =0;(     *C)[    1]<3;(*C)[1]    ++)c(*C
+       [3      ]|((     *C)[  4]?**C>18&&* *C<42 ?C[1][
+       42     +*(*      C+1)    +3***C]:0: **C>=  11&&*
+       *C     <64?      ~C[1 ][   7***C+97 +(*C)[  1]]:
+        0)   );c(       *C[3 ]++)   ;for(C (1)=0;   (C(
+        2)   =C(1       ))<C (0);)   {(*C) [2]=C   [2][
+        1]  -49;        c=(* C[2]<=   63); c=(*   C)[0]
+        -4  *(C[        3][0 ]=105-  C[2][ c]   -7*(*(C
+        [2]+c)<         'c') -18*(  C[2][c    ]<77)+2*(
+         *C)[4          ]-7* (C[2] [c]<'C'  ))-6;for(C(
+        3)=0;                                          
+       C(3)<                                (*C[2]?*C[2
+      ]>'r'                             ?C[    2][1]:(1
+     +2*(*C             [2]>   64)    +(2-!C[2    ][3])
+    *(58 ==             (*(C         +2))[0])-         
+    (C[2] [0            ]=='m'   )  ):C(2)?C(0     )-C(
+  2):0  );C(3)++)       C(1)+=c(C  [1][4]|(*     C[2]&&
+ *C[2  ]<'s'?*C[2]      ==58?C(3)  ==1?**C     >17&&**C
+ <51  ?C[ 2]    [1]     ==59?39:  C[2][1     ]==58?9:1:
+0:0  :63  >*     C[2    ]?(c<7&&  c>-9?C[1    ][(*C[2]<
+45?  'c'  *5     +2*    '%':*C[   2]<  61?570    :571)+
+3*c  ]:   0)      :*C   [2]>'o'  ?**C>26&&**C<29    &&!
+(*C  )[    2]     ||(   *C)[2]  ==1&&(&*C[0])       [0]
+ <34  &&   31     <**   C?C(3)  <2?15+225*C(3     ):0:(
+ *C)[  2]  ==     3?C   (3)<2  &&**C>22&&**C     <45&&C
+ (3)<      2?     C[    1][7*  *(*C)+151+C      (3)]:0:
+  7==(     *C    )[2    ]&&*  *C>26&&**C       <42&&C(3
+  )<2?C    [1   ][7     ***C  +135+C(3)]       :0:*C[2]
+   <'k'?    (c >-5                      &&      c<5?C(3
+    )<2?C   [1][(                       (*C)     [2]<3?
+     207:205)+7*                        c+C(3     )]:C[
+      2][2]==46               &&        (c==-2||   c==-
+            1-                          2*(*C[3]%2  ))?
+            96                          :0:0)|((*C)[2  
+             ]?                                        
+             C(         3      )        ?C(3)< 2       
+             &&                        *C[3]>  7   &&  
+             c<           1&&c  >-24  ?8:0:*   C       
+             [3         ]<8&&c  >1   &&c<24    ?   ' ' 
+        :0   :0)        |(C(3) <2     &&(  **  C       
+       ==66   &&        *C[3   ]>14   ||*  C[  3       
+       ]>12   &&        58   ==**C||   *               
+       C[3    ]<        2  &&**C==10   )            ?  
+        5*  51:                       0)|(7==  (       
+         *C)[                        2]?*C[3]  <       
+8?c>13&& c<23&&C(3)<2?C
+[1][144+ 7*c+C(3)]:0:C(
+3) && c< -14&&c>- 24?C[
+1] [7 *c +400+C(3 )]:0:
+0) :! C  (3)?**C> 21&&C
+[0 ][    0]   <32      
+?C     [ 1 ]   [(      
+*   C )    [2   ] +323 
+   +7     **   *C ]:36 
+ + 1   < **   C&& '0'> 
+*     *C ?  C[1][      
+   C[ 2]  [2]+162      
+ + 7* **C]:0:0:0));C[1 
+][ 3] ++;C[2]+=3;}c(0) 
+;C [3 ]-=2;}*C[3]=0;}}
 
-\section{Conclusions}
-\subsection{Summary of Results}
-* General concept of simplifying grammars, two proposed algorithms
-* reasonably efficient implementation with runtime better than expected
-* compared algorithms, evaluated effectiveness of reducer
-* looked at effect of epsilon
+\end{lstlisting}
 
-\subsection{Ratios and Fidelity}
+\subsection*{Output:}
 
-* How useful are these algorithms??
-* What is the ratio we get
-* What kind of fidelity to we get
+\begin{lstlisting}
+#include<stdio.h>
+#define c(C) prinf("%c",C)
+#define C(c) (in*)C[]+6))[c]
+mainin c,char  
+*C[])     {C[c]=C
+]+  )[0]=  c(52*c(\
+C'  '4'/4)   );for(c
+=0; c<491;   c)for(*
+*  C[]+'c'  c]  =
+0;*  0]<8;  **  C
+)    C[]+c  'c']=
+C[  ]+c+'c'  C[]+99   c]+C[  ][**C
++8*  +99]=3  );  
+*)[]=*C
+]=  7  
+*C[]+=3)-  )==70?
+1):;C()=C(  )=c=0
+;whileC
+?  2][]+
+?C[]+2)?  ):0)
+{if(*C [    >'w'){C()=0;C[]+  ]++*C
+[2]0;else   C(1)+=C[
+2]=58?2+  C
+3]&&
+*C
++3]  'x'):C
 
-\subsection{Future Work}
-* Better ways to set epsilon to maximize value / cost for different size and types
-* Better ways to quantify preservation of hierarchies
-* How to find near `self-similarities' that aren't captured by a rule
-* Iterating lossify and reduce in a cycle
+==''?C[  C[]+-=48):
+*3]>=6  3-C
+]=\
+m'?10)  1;C()=C(1)>C(0)?C(  :C();c+=3;*
+C[+)+=3;}prinf("  \
+      \n",    56+8*C(  0),80**C
+  +))
+     *C3]=0    C
+    -=c*C3]    0;
+     whileC
+  [1,-    ]+--){  for  **
+    =0  ;*  *C      80;(*  C)++)  {C
+    3]  -=  **      3]  C[3]  ++    0;      3]      =**>=  51||  *C<    8    ||    *C      8!=2?0  25  ;c(1    -    );        3]);for    (*)[    ]  0;      <3;(*)[]+  +)c(*C    [   ]|(      4]**18&&*C<42 ?C[]+    42   (*  C+1)  3***C]0): **  11&&    *C  64?
+~C
+  7***C+97 +(*)[  ]+]:    0)  );c(     ]++  ;forC[ (1)0;  C[    +)  C(    ))<C (0);  {(*) 3]=C  ][    ]+  -49    c=(* C
+<   63); c=(*   )[0]    -4  C[    3][02]105-  ][ c  -7(*C[    ][c)<    'c' -18*  C[]+    <77)2*    *C)[     ]-7* (C
+ [c]<'C'  )-6;forC[    3)0;              C(3]       (C
+?C
 
-\nocite{*}
+  ]>'r'     C[  C[]+:(1
+  2*      ][  64   +(2-!C
+  3])
+  (58]=      (*C[    +))[0])-    
+  C[3] [0    ]='m'  ):C(+)?C(  )-C(  +)0)  ;C(3)++    C()+=c(  C[]+4]|(    3]&&*C2  ]<''?C
+    =58?C(3)  ]=1?**  >17&&*C <51  C[ ]  ]+    =59?39  C[]+  2]=58?9:1:
+00)  6  >    2   ?(c<7&& c>-9?C
+  (*][
+45  'c'    2*  '%':C[  2]  61?570  571)+
+3*  ]:  0)  *C  ]['o'  **&&*C<29  &&!
+(*     ]  ||  (*)[  ]=&&(&C[0]    [0] <34  &&  3   <**  ?C(3)  2?15+225*C(  ):0:(*C)[    ]=  3?  C(3)<2  &&**22&&*C  45&&C
+C(3)    2?  C[  [7*  (*)+151+    3)]::
+  ]=      )[2   &&*C>&&*C    <42&&C(  )<2?  [1  [7  ***C  +135+C(3)    *C3]   <'k'?  c >-          c<5?C(    2?   []+         (*C  ][3?  207:205)+7        c+C(  ):C
+  ][2]]=46          c]=-2||  c]=-    1-        2*(*3]%  ))?
+96      0)|((*)[       ]              *C            ?C(3)<                   ]>           c<     &&  >-  ?8):           [      <8&&  >1  &&c<       ' '      +)  +)    |(C(3) <    &&  **          ]=66           ]>1   ||*              >1        58  ]=**||                <       &&*C]=10            5*  51      0)|(7]=          (*)[      2]?C[  <     8?c>1&& c<2&&C(3)2?C
+C[]+144+ 7*c+C():0:C(
+3) && c< -14&&c>- 24C[
+1] [7 *c +400+C( ):0:
+) :!*C  3)** 2&&C
+[0   0]  3    
+?   [ 1 ]       
+*   )   2    +323    +7     *C ]:36 
+ + 1   **  && '0'> 
+  *C   C
+][      [ 2  ][16    
+ + 7* **]:):0));C
 
-\bibliography{sources}
+][ 3] ++C
++=3;}c() 
+;C [3 ]-=2;}C[]=0;}
+
+\end{lstlisting}
+
+\pagebreak
+
+\subsection*{gen256000 Output String, Cluster, $\epsilon=.4$:}
+
+\subsubsection*{Input:}
+
+\begin{lstlisting}
+1:1 In the beginning God created the heaven and the earth.
+
+1:2 And the earth was without form, and void; and darkness was upon
+the face of the deep. And the Spirit of God moved upon the face of the
+waters.
+
+1:3 And God said, Let there be light: and there was light.
+
+1:4 And God saw the light, that it was good: and God divided the light
+from the darkness.
+
+1:5 And God called the light Day, and the darkness he called Night.
+And the evening and the morning were the first day.
+
+1:6 And God said, Let there be a firmament in the midst of the waters,
+and let it divide the waters from the waters.
+
+1:7 And God made the firmament, and divided the waters which were
+under the firmament from the waters which were above the firmament:
+and it was so.
+
+1:8 And God called the firmament Heaven. And the evening and the
+morning were the second day.
+
+1:9 And God said, Let the waters under the heaven be gathered together
+unto one place, and let the dry land appear: and it was so.
+
+1:10 And God called the dry land Earth; and the gathering together of
+the waters called he Seas: and God saw that it was good.
+
+1:11 And God said, Let the earth bring forth grass, the herb yielding
+seed, and the fruit tree yielding fruit after his kind, whose seed is
+in itself, upon the earth: and it was so.
+
+1:12 And the earth brought forth grass, and herb yielding seed after
+his kind, and the tree yielding fruit, whose seed was in itself, after
+his kind: and God saw that it was good.
+
+1:13 And the evening and the morning were the third day.
+
+1:14 And God said, Let there be lights in the firmament of the heaven
+to divide the day from the night; and let them be for signs, and for
+seasons, and for days, and years: 1:15 And let them be for lights in
+the firmament of the heaven to give light upon the earth: and it was
+so.
+
+1:16 And God made two great lights; the greater light to rule the day,
+and the lesser light to rule the night: he made the stars also.
+
+1:17 And God set them in the firmament of the heaven to give light
+upon the earth, 1:18 And to rule over the day and over the night, and
+to divide the light from the darkness: and God saw that it was good.
+
+1:19 And the evening and the morning were the fourth day.
+
+1:20 And God said, Let the waters bring forth abundantly the moving
+creature that hath life, and fowl that may fly above the earth in the
+open firmament of heaven.
+
+1:21 And God created great whales, and every living creature that
+moveth, which the waters brought forth abundantly, after their kind,
+and every winged fowl after his kind: and God saw that it was good.
+
+1:22 And God blessed them, saying, Be fruitful, and multiply, and fill
+the waters in the seas, and let fowl multiply in the earth.
+
+1:23 And the evening and the morning were the fifth day.
+
+1:24 And God said, Let the earth bring forth the living creature after
+his kind, cattle, and creeping thing, and beast of the earth after his
+kind: and it was so.
+
+1:25 And God made the beast of the earth after his kind, and cattle
+after their kind, and every thing that creepeth upon the earth after
+his kind: and God saw that it was good.
+
+1:26 And God said, Let us make man in our image, after our likeness:
+and let them have dominion over the fish of the sea, and over the fowl
+of the air, and over the cattle, and over all the earth, and over
+every creeping thing that creepeth upon the earth.
+\end{lstlisting}
+
+\subsubsection*{Output:}
+
+\begin{lstlisting}
+1:1 n e beininGod Hae mae ane ar
+
+ine ars itht frm,n void;n darkness aupon
+t ofe en.ine Spirt ofGod moend upont ofwate
+
+3e said, htn e s ht
+
+ine w ht, tht waodn God dividhtfrm darknes
+
+5caledht Dayn darkness e caledNht
+n e e frinwee e frt
+
+6e said,  amament ine midt ofwate
+ndt t dividwatewate
+
+7ine fealn dividwatehichs in
+underfeal wates hichs inabenfeal:n t wa
+
+8caledmamenHaen.e e e 
+frinwee e secon
+
+9e said, undere mae bemae e mae entoone plcen t e dry ln apart wa
+
+0caleddry ln Earn e mae inmae  fwatecalede Sarine tht waod
+
+ine id, Lt e frargras, e ding
+seedn e rut dingrut fes inwhseseed i
+in tel, e frt wa
+
+n e frrurhtargras, n dingseedfes inn e dingrut , whseseeds in tel, fe
+s inine tht waod3e e frinwee e thir
+
+1e said, e behts ine mamene e 
+todividthayfrm e htn e e e frsignd an
+seasond anayn ar: 1:n e e e frhts in
+e frmamene e  toinhte fr: n t wa
+
+6ine  tcrathts;e grHatoule thay
+nde estoule e hte e e e stars al
+
+7ine t e ine mamene e toinht
+e fr1:n toule e e ayn e e htd an
+todivide htfrm darknes: God tht waod9n e e n frinwee e fr
+
+0e said, binar aundanolye moilngcaledthhath in frwlthay flyabene frentoopen firmamenofmae
+
+1 n God Haedcratwhalesn enry lcalede 
+moen, hiche wate brurhtar aundanoly, fe
+s inn enrys inedfrwlfes in: God th ts no
+
+n God es te m,saying, e rut fuln multiplyllwateine seasn t frwlmultiply in e fr
+
+3n e e n frinwee e fi
+
+e said, e frare lcaledfe
+s in, cattln crepn e frfehi
+int wa
+
+2n God Haede frfehis inn cattl
+fe
+s inn enry thinge frfes in: God th ts no
+
+6e said, usmae mae ur imae feur ienes:n e e mae dominon e e fish fe saan arfrwl
+e e n arcattln ar alle frn ar
+enrycrepe fr
+
+(Rest Omitted)
+
+\end{lstlisting}
 

### 2012-12-11 tex/paper.tex
Added in prelim background section

@@
 these structures, we built
-Music Final Composition: Dec 10
 our system on top of the existing \Sequitur grammar inference algorithm. 
@@
 along the way to emphasize repetitions. Our main contribution lies in the
-development of 2 algorithms for ``simplifying'' CFG grammars while 
-preserving their structure.
+development, implementation, and evaluation of 2 algorithms for 
+``simplifying'' CFG grammars while preserving their structure.
 
@@
 
-TODO
+The work in this paper builds upon existing research in lossless
+grammar based compression, and also draws inspiration from many
+techniques used in lossy image and video compression.
 
 \subsection{Grammar Compression}
+As described in the introduction, the proposed research shares many of its
+goals with the \Sequitur algorithm described in \cite{sequitur}.  The authors
+of \cite{sequitur} develop an effective algorithm based on iteratively
+rewriting grammars to keep them small and efficient, and produce interesting
+grammar-based analyses of texts and musical scores.
 
-TODO
+The same author in \cite{nevillphd}
+explores a variety of ad-hoc extensions to the \Sequitur algorithm which
+improve its compression performance on structured data.  These ranged from
+introducing domain-specific constraints to their grammar, adding a few steps of
+backtracking to their normally greedy grammar formation, to guessing
+unifications in attempting to infer recursive grammar rules.  Though these may
+invalidate some of their theoretical results on the asymptotic performance of
+\Sequitur, in practice they seem to have worked and add nicely to the grammar
+inference framework.
+
+The intution that grammar inference can support many detailed policies is made
+more formal in \cite{grammarcodes}.  The authors classify the properties a
+grammar needs to function as a good compressor, and moreover give a set of
+reduction rules for putting a grammar into an appropriate
+``irreducible'' form.  Within this
+context, grammar inference algorithms similar to \Sequitur can be described as
+simple applications of their reduction rules to different ways of generating
+base grammars.  Even more examples of the variety of grammar inference schemes
+possible under this general framework of reducing grammars can be found in
+\cite{efficientgreedy}
+
+There has also been very interesting work analyzing how grammar based
+algorithms compare with the theoretical best lossless grammar for a string, in
+the line of \cite{approximation}.  However, these analytic bounds for lossless
+grammars are outside the scope of our proposed research.
+
+In summary, established work on grammar inference algorithms provides a solid
+set of tools for experimenting with the kinds of local modifications we propose
+to make to improve grammar structure by introducing lossiness.
 
@@
 
-TODO
+While we were unable to find work in the area of lossy grammar compression, a
+number of existing techniques for lossy hierarchical compression more broadly
+may be applicable.
 
-Jpeg, Mpeg: using parts of previous frame to encode next frame, 
-Fractal: Using self-similarity to encode
+Kieffer briefly identifies two primary methods for lossy hierarchical
+compression: wavelet-based and fractal-based schemes \cite{tutorial}.  In
+wavelet-based schemas, a signal is compressed by recording the largest
+coefficients of its wavelet decomposition, in effect
+trying to represent a signal as closely as possible using wavelet
+building blocks. The JPEG image format uses a similar technique in
+order to compress images lossily \cite{jpeg}. Fractal based compression
+takes this a step further and represents parts of a signal using self-similarity,
+encoding an image for instance as a fixed point of contraction maps. This is
+similar in spirit to what is done in MPEG-1, where parts of one frame can be
+transformed to encode parts of the next frame.
+
+Work has also been done regarding the lossy compression of particular input
+formats.  For text, methods such as replacing words with synonyms
+\cite{semanticlossy} or reordering the middle letters of a word (i.e., all
+letters save for the first and last) \cite{semilossless} have been explored.
+However, these are very specific tricks to optimize text, and not more broadly
+applicable to compressing hierarchical streams.
 
@@
 
-The scope of what we shall call a \emph{lossifier} algorithm is to take
-an admissible CFG and simplify it. The output CFG should be smaller
+In order to make our algorithm for lossily compressing streams
+as general as possible, we decided
+to approach the problem of lossy compression by starting with a lossless
+grammar inference algorithm. 
+By building off of a grammar inference algorithm, we
+hope to take advantage of some of the partial structures
+that are visible in a CFG representation of a stream. 
+In this project, we chose \Sequitur, but
+any way of obtaining a grammar could be used. Then, we would transform
+the grammar, introducing loss and perform compressions by simplifying
+the grammar. Thus our compression system is a multi-step process:
+
+\begin{itemize}
+\item Infer a CFG representation of a stream (possibly using Sequitur)
+\item Simplify the CFG, possibly introducing loss
+\item Output the CFG
+\end{itemize}
+
+Our core new contribution comes in step two, in what we shall call a \emph{lossifier} 
+algorithm, i.e. an algorithm that takes
+an admissible CFG and simplifies it. The output CFG should be smaller
 than the input CFG but generate a string which preserves the
@@
 
-Building off the ideas in jpeg compression discussed
+Building off the ideas in lossy compression discussed
 earlier, one idea for simplifying data is to look at a part

### 2012-12-11 tex/paper.tex
redid math section, improved some small sentences in results section

@@
 In this project, we chose \Sequitur, but
-any way of obtaining a grammar could be used. Then, we would transform
+any way of obtaining a grammar could be used. Then, transformed
 the grammar, introducing loss and perform compressions by simplifying
@@
 into equivalence class clusters, and then replaces each variable
-in a cluster them with a representative
+in a cluster them with the smallest representative
 all at once each iteration.
@@
     \EndFor
+    \State save a mapping $m$ from variables to sizes
     \For{Cluster $c \in clusters$}
-      \State replace all $var \in c$ with the smallest.
+      \State replace all $var \in c$ with the smallest (smallest w.r.t. $m$).
     \EndFor
@@
 
-\subsection{Runtime Estimates}
+\subsection{Proof of Grammar Consistency}
+
+One risk in replacing variables in a CFG is that one could introduce
+cyclical expansion in the CFG, such as if 
+$A\rightarrow aB$ and $B\rightarrow bA$. In this 
+section we will argue that neither of our algorithms do so if the
+original grammar was well formed.
+
+We can view the CFG as a DAG where for variables $V_1$, $V_2$, we add an edge
+whenever $V_1$ contains $V_2$ in its rhs. When sizes are compuated
+on a CFG before we modify it, if $V_1 \rightarrow V_2$ then $size(V_1) > size(V_2)$.
+Replacing a variable $V_1$ with
+$V_2$ has the same effect as redirecting all in-edges to $V_1$ onto $V_2$.
+
+Suppose for the sake of contradiction that the Similarity algorithm
+created a cycle at one step by redirecting an edge originally pointing 
+at $A$ to $B$. This would mean that there was a path going from 
+$B$ to $A$. However, by comparing the sizes of $A$ and $B$ we know that
+$A$ must have been bigger than $B$. Thus there cannot have been a path from $B$
+to $A$, and we have a contradiction.
+
+For the Cluster algorithm the analysis is a bit more involved, since
+to avoid recomputation we only calculate the sizes of the variables once
+before doing all of the replacements. First note that even as we modify
+the DAG, all of the edges we create in the DAG will respect the ordering of
+sizes that we calculated before we did any replacements. This is because
+each time we redirect an edge from $X\rightarrow A$ to $X\rightarrow B$
+$size(X)>size(A)>size(B)$ and transitivity will preserve the ordering.
+
+Suppose then that we created a cycle during Cluster
+by redirecting an edge from $X \rightarrow A$ to $X \rightarrow B$.
+This means that there must have been a path from $B$ to $X$. However,
+even in its intermediate state we know that the DAG respects the
+$size$ ordering, and $size(X) > size(A) > size(B)$, so there cannot
+be a path from $B$ to $X$. This is a contradiction and we are done.
+
+\subsection{Runtime Analysis}
 Given an input of length $n$, let $r$ be the number of rules generated
-by Sequitur from the input. Let $R$ be the total number of variable
+by Sequitur from the input. 
+In the analysis of the Similarity algorithm, let $R$ be the total number of variable
 replacements we perform. For each replacement, we had to have searched though
@@
 them and calculating the levenshtein distance could take time linear in $n$.
+Performing the replacement only takes time linear in $r$.
 Thus the runtime is $\boxed{O(R*r^2*n)}$.
@@
 know that the number of rules in the grammar generated by Sequitur
-is between $O(\sqrt{n})$ and $O(n)$. Thus the worst case runtime on the similarity
+is between $O(1)$ and $O(n)$. Thus the worst case runtime on the similarity
 algorithm is $\boxed{O(n^4)}$. This is because we could have up to $r$ replacements.
 However, in practice we will see that the runtime is much better
-than $O(n^4)$. This is because we found $r$ for our test data to be slightly sublinear
-even for complicated english text, and we found that we didn't quite make a linear
-number of replacements $R$ with respect to $r$.
-so our optimistic runtime could be anywhere 
-from $O(n^4)$ to $O(n^{2})$ depending on how $r$ grows.
+than $O(n^4)$. There are a variety of possible reasons for this:
+in our test data, $r$ didn't grow quite linearly in $n$, $R$ could grow much
+slower than $r$, and the use of native code in calculating the levenshtein distance 
+made the its contribution much less noticeable for the input sizes we tested.
 
@@
 iteration is absorbed into the $O(r^2)$ term. So, again we have a runtime of
-$\boxed{O(C*r^2*n)}$. The big gain over the similarity algorithm is that $C$ turns out
+$\boxed{O(C*r^2*n)}$. Just as with the similarity algorithm,
+a variety of factors make the runtime much better than the
+$\boxed{O(n^4)}$ worst case scenario. 
+The big gain over the similarity algorithm is that $C$ turns out
 to be nearly constant, since there are very few huge chains of reductions that are
@@
 perform a very large number of the remaining reductions each time we perform an iteration
-of the clustering algorithm. Thus since $C$ is nearly constant, the optimistic runtime
-could be anywhere from $O(n^3)$ to $O(n^2)$.
-
-CANT THINK OF PROOF OF NOLOOPS
+of the clustering algorithm. Thus a more reasonable upper bound for the runtime
+of cluster would be $\boxed{O(n^3)}$
 
@@
 What is their purpose? Why do we want them? What are they?
+Grammar reductions are a way of taking a CFG, generated possibly through
+ad-hoc techniques, and systematically removing internal redundancies
+like duplicate bigrams in Sequitur. They make the CFG representation smaller
+but don't introduce any loss. We want them since our lossifier algorithms
+often make our grammars redundant, even though the each step we take
+in the lossifier makes the grammars smaller.
 
@@
 Another set of test data was constructed from the first n characters of the
-book of Genesis from the KJV. Call these test-cases \emph{gen-n}
+book of Genesis from the KJV version of the Bible. 
+Call these test-cases \emph{gen-n}
 
@@
 
-Below are the results on running Sequitur and Cluster, without
+In table \ref{simclus_r_t} we report the results on running Sequitur 
+and Cluster, without
 the grammar reductions, on the rep-n datasets. Time is the wall clock
@@
 The columns of the data are prefixed by S or T depending on whether
-they are measurements for sequitur or cluster respectively.
+they are measurements for Similarity or Cluster respectively.
 
@@
 
+The actual compression ratio was respectable, with both
+Similarity and Cluster coming in at nearly half the size of the Sequitur CFG.
 Setting higher values for $\epsilon$ had diminishing returns for compression
@@
 
-Repeating the above tests on the \emph{gen-n} datasets yields the following:
+Repeating the above tests on the \emph{gen-n} datasets yields the results in table
+\ref{simclus_g_t}
 
 Just as with the rep-n data, log-log regression tests yield the following
-exponents for the runtime on the gen-n data.
+exponents for the runtime on the gen-n data, again with very high values of
+$R^2$.
 
@@
 
-Again, the two algorithms have almost identical compression ratios and
+Just as in the rep-n results, the two algorithms have 
+almost identical compression ratios and
 their fidelity is about the same both subjectively and based on levenshtein
-distance.
-
-INCLUDE SAMPLE TEXT
+distance. For an example of the text generated on one of the gen-n datasets,
+see the appendix.
 
@@
 \caption{rep-n Tests}
-\label{seqclus_r_t}
+\label{simclus_r_t}
 \end{table}
@@
 \caption{gen-n Tests}
-\label{seqclus_g_t}
+\label{simclus_g_t}
 \end{table}
@@
 every creeping thing that creepeth upon the earth.
+
+(Rest Omitted)
+
 \end{lstlisting}

### 2012-12-11 tex/paper.tex
Reflow and kill trailing whitespace

@@
 \documentclass[11pt]{article}
+\usepackage[letterpaper,text={7in,9in},centering]{geometry}
+\usepackage{algorithm}
+\usepackage{algpseudocode}
 \usepackage{amsfonts,amsmath,amssymb,amsthm}
 \usepackage{array}
-\usepackage{bm}
-\usepackage{commath}
-\usepackage{enumitem}
 \usepackage{fancyhdr,fancyvrb}
-\usepackage[letterpaper,text={7in,9in},centering]{geometry}
 \usepackage[pdftex]{graphicx}
+\usepackage{hyperref}
+\usepackage{listings}
+\usepackage{natbib}
 \usepackage[lofdepth,lotdepth]{subfig}
-\usepackage{ifthen}
-\usepackage{mathrsfs}
-\usepackage{multicol}
-\usepackage{pifont}
 \usepackage{sectsty}
@@
 \usepackage{stackrel}
-\usepackage{stmaryrd}
-\usepackage{tensor}
 \usepackage{xspace}
-\usepackage{natbib}
-\usepackage{algorithm}
-\usepackage{algpseudocode}
-\usepackage{listings}
-\usepackage{hyperref}
-\lstset{basicstyle=\ttfamily\footnotesize,breaklines=true}
 
@@
 \newcommand{\Sequitur}{\textsc{Sequitur}\xspace}
+
+\lstset{basicstyle=\ttfamily\footnotesize,breaklines=true}
 
@@
 \begin{abstract}
-Grammar-based compression schemes like \Sequitur exploit repeated structures
-in streams of symbolic data to achieve good compression ratios.
-We develop a set of algorithms for improving upon Grammar-based compression 
-schemes by allowing some loss in the compression. Although precise 
-subsequences might be altered, in return we can simplify 
-the grammars produced to obtain better ratios while still preserving
-important syntatic patterns at both the micro and macro level.
+
+Grammar-based compression schemes like \Sequitur exploit repeated structures in
+streams of symbolic data to achieve good compression ratios.  We develop a set
+of algorithms for improving upon Grammar-based compression schemes by allowing
+some loss in the compression. Although precise subsequences might be altered,
+in return we can simplify the grammars produced to obtain better ratios while
+still preserving important syntatic patterns at both the micro and macro
+levels.
+
 \end{abstract}
@@
 \section{Introduction and Goals}
+
 Many sources of data exhibit natural hierarchical structure. For instance,
-english text often can be grouped well into sections, paragraphs, sentences, ...
-all the way down to words and phonemes. Music too, consists of repeated motifs,
-phrases, themes, and so on. Finally, computer code is so hierarchical that its
-syntax can usually be put in BNF. This allows grammar-based compression schemes
-to build up Context Free Grammars (CFGs) and achieve very good compression
-ratios.
+english text often can be grouped well into sections, paragraphs, sentences,
+...  all the way down to words and phonemes. Music too, consists of repeated
+motifs, phrases, themes, and so on. Finally, computer code is so hierarchical
+that its syntax can usually be put in BNF. This allows grammar-based
+compression schemes to build up Context Free Grammars (CFGs) and achieve very
+good compression ratios.
 
@@
 logos, one could use a lossless image compression scheme. In other situations
-however, such as for textures and some photographs, the exact pixel values
-are almost irrelevant compared to the overall gradients, colors, and structures
-in the image, and one could use a lossy compression scheme like jpeg.
+however, such as for textures and some photographs, the exact pixel values are
+almost irrelevant compared to the overall gradients, colors, and structures in
+the image, and one could use a lossy compression scheme like jpeg.
 
@@
 
-It may seem strange to allow a string such as "mind" to be lossily 
-corrupted into ``rind'', but if one is more interested in recurrent 
-structure in the text than in its precise meaning, these changes 
-aren't important. For instance, if one is looking at a text in a foreign
-language, one is more likely to pick up the texture, the repeated sounds,
-some phrases that come up and the breaks between larger scale sections.
-A change in a single letter probably wouldn't be noticed. Similarly,
-when skimming through huge files, perhaps execution logs, one usually
-first starts looking for patterns or things that stand out. Lorem ipsum
-is a perfect example of a piece of ``text'' whose value lies in its texture
-and sentence/paragraph structure, and we will come back to it again.
+It may seem strange to allow a string such as "mind" to be lossily corrupted
+into ``rind'', but if one is more interested in recurrent structure in the text
+than in its precise meaning, these changes aren't important. For instance, if
+one is looking at a text in a foreign language, one is more likely to pick up
+the texture, the repeated sounds, some phrases that come up and the breaks
+between larger scale sections.  A change in a single letter probably wouldn't
+be noticed. Similarly, when skimming through huge files, perhaps execution
+logs, one usually first starts looking for patterns or things that stand out.
+Lorem ipsum is a perfect example of a piece of ``text'' whose value lies in its
+texture and sentence/paragraph structure, and we will come back to it again.
 
 In this paper we develop 2 algorithms and implement a system for lossily
-compressing text while preserving its hierarchical structures. The goal
-was to preserve the feel of a text at both micro and macro levels. 
-Since grammar-based compression schemes were designed to capture
-these structures, we built
-our system on top of the existing \Sequitur grammar inference algorithm. 
-Starting with the CFG generated by a lossless grammar
-inference algorithm, we repeatedly simplify it, throwing out details
-along the way to emphasize repetitions. Our main contribution lies in the
-development, implementation, and evaluation of 2 algorithms for 
-``simplifying'' CFG grammars while preserving their structure.
+compressing text while preserving its hierarchical structures. The goal was to
+preserve the feel of a text at both micro and macro levels.  Since
+grammar-based compression schemes were designed to capture these structures, we
+built our system on top of the existing \Sequitur grammar inference algorithm.
+Starting with the CFG generated by a lossless grammar inference algorithm, we
+repeatedly simplify it, throwing out details along the way to emphasize
+repetitions. Our main contribution lies in the development, implementation, and
+evaluation of 2 algorithms for ``simplifying'' CFG grammars while preserving
+their structure.
 
@@
 
-The work in this paper builds upon existing research in lossless
-grammar based compression, and also draws inspiration from many
-techniques used in lossy image and video compression.
+The work in this paper builds upon existing research in lossless grammar based
+compression, and also draws inspiration from many techniques used in lossy
+image and video compression.
 
 \subsection{Grammar Compression}
+
 As described in the introduction, the proposed research shares many of its
@@
 
-The same author in \cite{nevillphd}
-explores a variety of ad-hoc extensions to the \Sequitur algorithm which
-improve its compression performance on structured data.  These ranged from
-introducing domain-specific constraints to their grammar, adding a few steps of
-backtracking to their normally greedy grammar formation, to guessing
-unifications in attempting to infer recursive grammar rules.  Though these may
-invalidate some of their theoretical results on the asymptotic performance of
-\Sequitur, in practice they seem to have worked and add nicely to the grammar
-inference framework.
+The same author in \cite{nevillphd} explores a variety of ad-hoc extensions to
+the \Sequitur algorithm which improve its compression performance on structured
+data.  These ranged from introducing domain-specific constraints to their
+grammar, adding a few steps of backtracking to their normally greedy grammar
+formation, to guessing unifications in attempting to infer recursive grammar
+rules.  Though these may invalidate some of their theoretical results on the
+asymptotic performance of \Sequitur, in practice they seem to have worked and
+add nicely to the grammar inference framework.
 
@@
 grammar needs to function as a good compressor, and moreover give a set of
-reduction rules for putting a grammar into an appropriate
-``irreducible'' form.  Within this
-context, grammar inference algorithms similar to \Sequitur can be described as
-simple applications of their reduction rules to different ways of generating
-base grammars.  Even more examples of the variety of grammar inference schemes
-possible under this general framework of reducing grammars can be found in
-\cite{efficientgreedy}
+reduction rules for putting a grammar into an appropriate ``irreducible'' form.
+Within this context, grammar inference algorithms similar to \Sequitur can be
+described as simple applications of their reduction rules to different ways of
+generating base grammars.  Even more examples of the variety of grammar
+inference schemes possible under this general framework of reducing grammars
+can be found in \cite{efficientgreedy}
 
@@
 wavelet-based schemas, a signal is compressed by recording the largest
-coefficients of its wavelet decomposition, in effect
-trying to represent a signal as closely as possible using wavelet
-building blocks. The JPEG image format uses a similar technique in
-order to compress images lossily \cite{jpeg}. Fractal based compression
-takes this a step further and represents parts of a signal using self-similarity,
-encoding an image for instance as a fixed point of contraction maps. This is
-similar in spirit to what is done in MPEG-1, where parts of one frame can be
-transformed to encode parts of the next frame.
+coefficients of its wavelet decomposition, in effect trying to represent a
+signal as closely as possible using wavelet building blocks. The JPEG image
+format uses a similar technique in order to compress images lossily
+\cite{jpeg}. Fractal based compression takes this a step further and represents
+parts of a signal using self-similarity, encoding an image for instance as a
+fixed point of contraction maps. This is similar in spirit to what is done in
+MPEG-1, where parts of one frame can be transformed to encode parts of the next
+frame.
 
@@
 
-In order to make our algorithm for lossily compressing streams
-as general as possible, we decided
-to approach the problem of lossy compression by starting with a lossless
-grammar inference algorithm. 
-By building off of a grammar inference algorithm, we
-hope to take advantage of some of the partial structures
-that are visible in a CFG representation of a stream. 
-In this project, we chose \Sequitur, but
-any way of obtaining a grammar could be used. Then, transformed
-the grammar, introducing loss and perform compressions by simplifying
-the grammar. Thus our compression system is a multi-step process:
+In order to make our algorithm for lossily compressing streams as general as
+possible, we decided to approach the problem of lossy compression by starting
+with a lossless grammar inference algorithm.  By building off of a grammar
+inference algorithm, we hope to take advantage of some of the partial
+structures that are visible in a CFG representation of a stream.  In this
+project, we chose \Sequitur, but any way of obtaining a grammar could be used.
+Then, transformed the grammar, introducing loss and perform compressions by
+simplifying the grammar. Thus our compression system is a multi-step process:
 
 \begin{itemize}
-\item Infer a CFG representation of a stream (possibly using Sequitur)
+  \item Infer a CFG representation of a stream (possibly using \Sequitur)
   \item Simplify the CFG, possibly introducing loss
@@
 
-Our core new contribution comes in step two, in what we shall call a \emph{lossifier} 
-algorithm, i.e. an algorithm that takes
-an admissible CFG and simplifies it. The output CFG should be smaller
-than the input CFG but generate a string which preserves the
-structure in the original string as much as possible.
-
-Building off the ideas in lossy compression discussed
-earlier, one idea for simplifying data is to look at a part
-of the data and try to match it as closely as possible with an
-existing set of primitives. In particular, just as with Mpeg and
-fractal compression, to approximate one part of the grammar
-we can use \emph{other parts of the grammar}. This is the seed for
-both of our algorithms. Given a rule $S \rightarrow rhs$, we will
-try to replace occurences of $S$ with other similar variables.
+Our core new contribution comes in step two, in what we shall call a
+\emph{lossifier} algorithm, i.e. an algorithm that takes an admissible CFG and
+simplifies it. The output CFG should be smaller than the input CFG but generate
+a string which preserves the structure in the original string as much as
+possible.
+
+Building off the ideas in lossy compression discussed earlier, one idea for
+simplifying data is to look at a part of the data and try to match it as
+closely as possible with an existing set of primitives. In particular, just as
+with Mpeg and fractal compression, to approximate one part of the grammar we
+can use \emph{other parts of the grammar}. This is the seed for both of our
+algorithms. Given a rule $S \rightarrow rhs$, we will try to replace occurences
+of $S$ with other similar variables.
 
@@
 \centering
-\includegraphics[scale=0.6]{sred1.pdf}
-\includegraphics[scale=0.6]{sred2.pdf}
+\includegraphics[scale=0.6]{include/sred1.pdf}
+\includegraphics[scale=0.6]{include/sred2.pdf}
 \caption{Grammar Simplification Example}
@@
 
-To simplify the grammar, it would make sense to replace B with A, since
-they form two almost identical halves of S and are similar.
+To simplify the grammar, it would make sense to replace B with A, since they
+form two almost identical halves of S and are similar.
 
@@
 To define whether two variables are \emph{similar}, we can look at their
-expansions. Let $expand(rhs)$ denote the string of terminal symbols we get
-when we recursively substitute all of the variables inside $rhs$. Then
-$V_1 \rightarrow rhs_1$ and $V_2 \rightarrow rhs_2$ are similar when
-
-$$\frac{levenshtein(expand(rhs_1),expand(rhs_2))}
-        {max(len(expand(rhs_1)),len(expand(rhs_2)))} < \epsilon$$.
-
-where $levenshtein$ is the levenstein edit distance between two strings.
-For convenience we can define.
-
-$$lsim(s1,s2) = \frac{levenshtein(s1,s2)}{max(len(s1),len(s2))}$$
-
-The first algorithm takes the idea of variable similarity and uses it
-to simplify the grammar sequentially as much as possible.
+expansions. Let $expand(rhs)$ denote the string of terminal symbols we get when
+we recursively substitute all of the variables inside $rhs$. Then $V_1
+\rightarrow rhs_1$ and $V_2 \rightarrow rhs_2$ are similar when
+
+\[
+  \frac{levenshtein(expand(rhs_1),expand(rhs_2))}
+      {max(len(expand(rhs_1)),len(expand(rhs_2)))} < \epsilon
+\]
+
+where $levenshtein$ is the levenstein edit distance between two strings.  For
+convenience we can define.
+
+\[ lsim(s1,s2) = \frac{levenshtein(s1,s2)}{max(len(s1),len(s2))} \]
+
+The first algorithm takes the idea of variable similarity and uses it to
+simplify the grammar sequentially as much as possible.
 
@@
 
-The previous algorithm updates the grammar sequentially, since each
-variable replacement can have an effect on large parts of the grammar,
-we will have to go through comparing all pairs again after the change.
-An alternative algorithm would be to try to replace similar variables in
-parallel. This would avoid many of the repeat pair-comparisons done in
-\emph{Similarity}. In other words, the algorithm below puts 
-similar variables
-into equivalence class clusters, and then replaces each variable
-in a cluster them with the smallest representative
-all at once each iteration.
+The previous algorithm updates the grammar sequentially, since each variable
+replacement can have an effect on large parts of the grammar, we will have to
+go through comparing all pairs again after the change.  An alternative
+algorithm would be to try to replace similar variables in parallel. This would
+avoid many of the repeat pair-comparisons done in \emph{Similarity}. In other
+words, the algorithm below puts similar variables into equivalence class
+clusters, and then replaces each variable in a cluster them with the smallest
+representative all at once each iteration.
 
@@
 Both of the algorithms depend on the concept of the "smallest / smaller"
-variable. We can define the size of a variable to be the length of
-the expansion of its rhs. In this way, if we think of our CFG's as a DAG, then
-variables higher up on the DAG will always have larger expansions than
-those lower down, and thus will have a larger size. By always replacing
-with the smallest variable, we avoid creating cyclic references in our CFG.
-We will prove this fact in the next section.
+variable. We can define the size of a variable to be the length of the
+expansion of its rhs. In this way, if we think of our CFG's as a DAG, then
+variables higher up on the DAG will always have larger expansions than those
+lower down, and thus will have a larger size. By always replacing with the
+smallest variable, we avoid creating cyclic references in our CFG.  We will
+prove this fact in the next section.
 
@@
 
-One risk in replacing variables in a CFG is that one could introduce
-cyclical expansion in the CFG, such as if 
-$A\rightarrow aB$ and $B\rightarrow bA$. In this 
-section we will argue that neither of our algorithms do so if the
-original grammar was well formed.
+One risk in replacing variables in a CFG is that one could introduce cyclical
+expansion in the CFG, such as if $A\rightarrow aB$ and $B\rightarrow bA$. In
+this section we will argue that neither of our algorithms do so if the original
+grammar was well formed.
 
 We can view the CFG as a DAG where for variables $V_1$, $V_2$, we add an edge
-whenever $V_1$ contains $V_2$ in its rhs. When sizes are compuated
-on a CFG before we modify it, if $V_1 \rightarrow V_2$ then $size(V_1) > size(V_2)$.
-Replacing a variable $V_1$ with
-$V_2$ has the same effect as redirecting all in-edges to $V_1$ onto $V_2$.
-
-Suppose for the sake of contradiction that the Similarity algorithm
-created a cycle at one step by redirecting an edge originally pointing 
-at $A$ to $B$. This would mean that there was a path going from 
-$B$ to $A$. However, by comparing the sizes of $A$ and $B$ we know that
-$A$ must have been bigger than $B$. Thus there cannot have been a path from $B$
-to $A$, and we have a contradiction.
-
-For the Cluster algorithm the analysis is a bit more involved, since
-to avoid recomputation we only calculate the sizes of the variables once
-before doing all of the replacements. First note that even as we modify
-the DAG, all of the edges we create in the DAG will respect the ordering of
-sizes that we calculated before we did any replacements. This is because
-each time we redirect an edge from $X\rightarrow A$ to $X\rightarrow B$
+whenever $V_1$ contains $V_2$ in its rhs. When sizes are compuated on a CFG
+before we modify it, if $V_1 \rightarrow V_2$ then $size(V_1) > size(V_2)$.
+Replacing a variable $V_1$ with $V_2$ has the same effect as redirecting all
+in-edges to $V_1$ onto $V_2$.
+
+Suppose for the sake of contradiction that the Similarity algorithm created a
+cycle at one step by redirecting an edge originally pointing at $A$ to $B$.
+This would mean that there was a path going from $B$ to $A$. However, by
+comparing the sizes of $A$ and $B$ we know that $A$ must have been bigger than
+$B$. Thus there cannot have been a path from $B$ to $A$, and we have a
+contradiction.
+
+For the Cluster algorithm the analysis is a bit more involved, since to avoid
+recomputation we only calculate the sizes of the variables once before doing
+all of the replacements. First note that even as we modify the DAG, all of the
+edges we create in the DAG will respect the ordering of sizes that we
+calculated before we did any replacements. This is because each time we
+redirect an edge from $X\rightarrow A$ to $X\rightarrow B$
 $size(X)>size(A)>size(B)$ and transitivity will preserve the ordering.
 
-Suppose then that we created a cycle during Cluster
-by redirecting an edge from $X \rightarrow A$ to $X \rightarrow B$.
-This means that there must have been a path from $B$ to $X$. However,
-even in its intermediate state we know that the DAG respects the
-$size$ ordering, and $size(X) > size(A) > size(B)$, so there cannot
-be a path from $B$ to $X$. This is a contradiction and we are done.
+Suppose then that we created a cycle during Cluster by redirecting an edge from
+$X \rightarrow A$ to $X \rightarrow B$.  This means that there must have been a
+path from $B$ to $X$. However, even in its intermediate state we know that the
+DAG respects the $size$ ordering, and $size(X) > size(A) > size(B)$, so there
+cannot be a path from $B$ to $X$. This is a contradiction and we are done.
 
 \subsection{Runtime Analysis}
-Given an input of length $n$, let $r$ be the number of rules generated
-by Sequitur from the input. 
-In the analysis of the Similarity algorithm, let $R$ be the total number of variable
-replacements we perform. For each replacement, we had to have searched though
-up to all pairs of variables, and for each pair of variables expanding
-them and calculating the levenshtein distance could take time linear in $n$.
-Performing the replacement only takes time linear in $r$.
+
+Given an input of length $n$, let $r$ be the number of rules generated by
+\Sequitur from the input.  In the analysis of the Similarity algorithm, let $R$
+be the total number of variable replacements we perform. For each replacement,
+we had to have searched though up to all pairs of variables, and for each pair
+of variables expanding them and calculating the levenshtein distance could take
+time linear in $n$.  Performing the replacement only takes time linear in $r$.
 Thus the runtime is $\boxed{O(R*r^2*n)}$.
 
-From the \Sequitur paper \cite{sequitur} we
-know that the number of rules in the grammar generated by Sequitur
-is between $O(1)$ and $O(n)$. Thus the worst case runtime on the similarity
-algorithm is $\boxed{O(n^4)}$. This is because we could have up to $r$ replacements.
-However, in practice we will see that the runtime is much better
-than $O(n^4)$. There are a variety of possible reasons for this:
-in our test data, $r$ didn't grow quite linearly in $n$, $R$ could grow much
-slower than $r$, and the use of native code in calculating the levenshtein distance 
-made the its contribution much less noticeable for the input sizes we tested.
-
-The inner loops of the Cluster are similar to those of the Similarity algorithm.
-We iterate the entire clustering and replacement procedure $C$ times. Going through
-each variable-cluster pair takes up to $O(r^2)$ time and calculating the levenshtein
-distance takes $O(n)$ time. The time taken to actually perform the replacements each
-iteration is absorbed into the $O(r^2)$ term. So, again we have a runtime of
-$\boxed{O(C*r^2*n)}$. Just as with the similarity algorithm,
-a variety of factors make the runtime much better than the
-$\boxed{O(n^4)}$ worst case scenario. 
-The big gain over the similarity algorithm is that $C$ turns out
-to be nearly constant, since there are very few huge chains of reductions that are
-only possible after other reductions have been performed, an in fact we are able to 
-perform a very large number of the remaining reductions each time we perform an iteration
-of the clustering algorithm. Thus a more reasonable upper bound for the runtime
-of cluster would be $\boxed{O(n^3)}$
+From the \Sequitur paper \cite{sequitur} we know that the number of rules in
+the grammar generated by \Sequitur is between $O(1)$ and $O(n)$. Thus the worst
+case runtime on the similarity algorithm is $\boxed{O(n^4)}$. This is because
+we could have up to $r$ replacements.  However, in practice we will see that
+the runtime is much better than $O(n^4)$. There are a variety of possible
+reasons for this: in our test data, $r$ didn't grow quite linearly in $n$, $R$
+could grow much slower than $r$, and the use of native code in calculating the
+levenshtein distance made the its contribution much less noticeable for the
+input sizes we tested.
+
+The inner loops of the Cluster are similar to those of the Similarity
+algorithm.  We iterate the entire clustering and replacement procedure $C$
+times. Going through each variable-cluster pair takes up to $O(r^2)$ time and
+calculating the levenshtein distance takes $O(n)$ time. The time taken to
+actually perform the replacements each iteration is absorbed into the $O(r^2)$
+term. So, again we have a runtime of $\boxed{O(C*r^2*n)}$. Just as with the
+similarity algorithm, a variety of factors make the runtime much better than
+the $\boxed{O(n^4)}$ worst case scenario.  The big gain over the similarity
+algorithm is that $C$ turns out to be nearly constant, since there are very few
+huge chains of reductions that are only possible after other reductions have
+been performed, an in fact we are able to perform a very large number of the
+remaining reductions each time we perform an iteration of the clustering
+algorithm. Thus a more reasonable upper bound for the runtime of cluster would
+be $\boxed{O(n^3)}$
 
 \section{Implementation}
+
 Overview
@@
 \subsection{Grammar Reductions}
-What is their purpose? Why do we want them? What are they?
-Grammar reductions are a way of taking a CFG, generated possibly through
-ad-hoc techniques, and systematically removing internal redundancies
-like duplicate bigrams in Sequitur. They make the CFG representation smaller
-but don't introduce any loss. We want them since our lossifier algorithms
-often make our grammars redundant, even though the each step we take
-in the lossifier makes the grammars smaller.
+
+What is their purpose? Why do we want them? What are they?  Grammar reductions
+are a way of taking a CFG, generated possibly through ad-hoc techniques, and
+systematically removing internal redundancies like duplicate bigrams in
+\Sequitur. They make the CFG representation smaller but don't introduce any
+loss. We want them since our lossifier algorithms often make our grammars
+redundant, even though the each step we take in the lossifier makes the
+grammars smaller.
 
 \subsection{Data Structures and Optimizations}
+
 Grammar Data Structure
@@
 \section{Results}
+
 \subsection{Illustrative Examples}
-It is useful to look at how the algorithms function on a small piece of
-data with both plenty of hierarchical structure, and also some messiness
-that most lossless compression schemes would not deal well with. Take
-the string below, similar to what one might come across in a math
-puzzle book.
+
+It is useful to look at how the algorithms function on a small piece of data
+with both plenty of hierarchical structure, and also some messiness that most
+lossless compression schemes would not deal well with. Take the string below,
+similar to what one might come across in a math puzzle book.
 
@@
 Disregarding its meaning, it consists of four very similar equations of the
-form $a + b = ?$, and on a smaller scale each summand consists of
-3 repetitions of approximately the same string of either $123,$ or $321,$.
-
-When comparing grammars, we chose to define the \emph{size} of a grammar
-as the sum of the lengths of the rhs's of the production rules.
-Sequitur yields a relatively complicated inferred grammar with 13 rules and
-a size of 51. Both the Similarity and Clustering Algorithms with $\epsilon = .4$
-, yield the following grammar after reduction. It has 5 rules and is almost
-half the size, at 30.
+form $a + b = ?$, and on a smaller scale each summand consists of 3 repetitions
+of approximately the same string of either $123,$ or $321,$.
+
+When comparing grammars, we chose to define the \emph{size} of a grammar as the
+sum of the lengths of the rhs's of the production rules.  \Sequitur yields a
+relatively complicated inferred grammar with 13 rules and a size of 51. Both
+the Similarity and Clustering Algorithms with $\epsilon = .4$ , yield the
+following grammar after reduction. It has 5 rules and is almost half the size,
+at 30.
 
@@
 
-The lossifier algorithm reduced the string to a pair of
-pairs of the same equations, where each equation consists of the
-sum of triples of "12" or "32". Smaller discrepencies
-in the numbers involved were dropped, while the filler
-commas and linebreaks also ended
-up being lost. Note that the structure described is easily
-observable from the grammar generated.
+The lossifier algorithm reduced the string to a pair of pairs of the same
+equations, where each equation consists of the sum of triples of "12" or "32".
+Smaller discrepencies in the numbers involved were dropped, while the filler
+commas and linebreaks also ended up being lost. Note that the structure
+described is easily observable from the grammar generated.
 
 \subsection{Similarity vs Clustering}
-For most of our benchmark tests, we focused on two extreme kinds of data:
-one with large amounts of repetition at all levels, and one with a word
-and sentence level structure but with little obvious large-scale repetition.
-One set of test data was constructed by repeating the same letter `a' n-times
-but then adding some
-texture to the data by changing 1/30 of the characters uniformly
-at random to `*'. Call these test-cases \emph{rep-n}
-Another set of test data was constructed from the first n characters of the
-book of Genesis from the KJV version of the Bible. 
-Call these test-cases \emph{gen-n}
+
+For most of our benchmark tests, we focused on two extreme kinds of data: one
+with large amounts of repetition at all levels, and one with a word and
+sentence level structure but with little obvious large-scale repetition.  One
+set of test data was constructed by repeating the same letter `a' n-times but
+then adding some texture to the data by changing 1/30 of the characters
+uniformly at random to `*'. Call these test-cases \emph{rep-n} Another set of
+test data was constructed from the first n characters of the book of Genesis
+from the KJV version of the Bible.  Call these test-cases \emph{gen-n}
 
 For our first suite of tests, we compare the effectiveness of the
-\emph{Similarity} and \emph{Cluster} algorithms. We might expect
-Similarity to achieve better quality at the same compression ratio
-since it updates the state of the grammar between each variable
-replacement, on the other hand since it must update the grammar
-it has the potential to run much more slowly than Cluster.
-
-We don't report the running times for just \Sequitur since its linear
-running time is negligible compared to our algorithms for all but the
-smallest input data.
+\emph{Similarity} and \emph{Cluster} algorithms. We might expect Similarity to
+achieve better quality at the same compression ratio since it updates the state
+of the grammar between each variable replacement, on the other hand since it
+must update the grammar it has the potential to run much more slowly than
+Cluster.
+
+We don't report the running times for just \Sequitur since its linear running
+time is negligible compared to our algorithms for all but the smallest input
+data.
 
@@
 
-In table \ref{simclus_r_t} we report the results on running Sequitur 
-and Cluster, without
-the grammar reductions, on the rep-n datasets. Time is the wall clock
-running as measured on a 2.26 Ghz Intel Core 2 Duo Macbook Pro.
-Ratio is the relative size of the grammar produced compared with
-Sequitur, where the size of the grammar is the sum of the lengths of the
-rhs's of the rules. Levdistance is the levenshtein distance between
-the output of the lossified grammar and the original input string, it
-is a kind of measure of the fidelity of our algorithms but doesn't
-take into account how well patterns were preserved. For the rep-n tests,
-proportion is the proportion of the final string produced that was
-`*'. The closer to 1/30~0.033, the more accurate the lossifier was
-in recreating the original texture of the input.
-
-The columns of the data are prefixed by S or T depending on whether
-they are measurements for Similarity or Cluster respectively.
-
-The compression ratios achieved for both algorithms are roughly equal for the same
-value for $\epsilon$, so it is fair to compare the algorithms for the same $\epsilon$.
-
-From plotting the data, we see that runtime of the algorithms are roughly polynomial
-in the size of the data, so we can perform a linear regression on the log-log transform
-of the data to find the exponents $\alpha$ for the runtime $r(n)=n^\alpha$ where $n$ is the
-length of the input string. In the chart below, all of the linear regressions had
-$R^2$ values of over $0.98$.
+In table \ref{simclus_r_t} we report the results on running \Sequitur and
+Cluster, without the grammar reductions, on the rep-n datasets. Time is the
+wall clock running as measured on a 2.26 Ghz Intel Core 2 Duo Macbook Pro.
+Ratio is the relative size of the grammar produced compared with \Sequitur,
+where the size of the grammar is the sum of the lengths of the rhs's of the
+rules. Levdistance is the levenshtein distance between the output of the
+lossified grammar and the original input string, it is a kind of measure of the
+fidelity of our algorithms but doesn't take into account how well patterns were
+preserved. For the rep-n tests, proportion is the proportion of the final
+string produced that was `*'. The closer to 1/30~0.033, the more accurate the
+lossifier was in recreating the original texture of the input.
+
+The columns of the data are prefixed by S or T depending on whether they are
+measurements for Similarity or Cluster respectively.
+
+The compression ratios achieved for both algorithms are roughly equal for the
+same value for $\epsilon$, so it is fair to compare the algorithms for the same
+$\epsilon$.
+
+From plotting the data, we see that runtime of the algorithms are roughly
+polynomial in the size of the data, so we can perform a linear regression on
+the log-log transform of the data to find the exponents $\alpha$ for the
+runtime $r(n)=n^\alpha$ where $n$ is the length of the input string. In the
+chart below, all of the linear regressions had $R^2$ values of over $0.98$.
 
@@
 
-For both values of $\epsilon$, 
-The clustering algorithm is nearly linear, while the similarity algorithm is roughly
-quadratic. On this dataset, the leveshtein distances were similar for both algorithms
-but fairly high since the length of the final string could change drastically.
-The proportion of '*'s in the input was usually the right order of magnitude,
-with both algorithms making big mistakes occassionally. In most cases,
-the two algorithms maintained similarly acceptable fidelity levels in recreating
-the texture of the original rep-n input file.
-
-The actual compression ratio was respectable, with both
-Similarity and Cluster coming in at nearly half the size of the Sequitur CFG.
-Setting higher values for $\epsilon$ had diminishing returns for compression
-for larger and larger file sizes, at the cost of drastically lowering the fidelity
-of the output.
+For both values of $\epsilon$, The clustering algorithm is nearly linear, while
+the similarity algorithm is roughly quadratic. On this dataset, the leveshtein
+distances were similar for both algorithms but fairly high since the length of
+the final string could change drastically.  The proportion of '*'s in the input
+was usually the right order of magnitude, with both algorithms making big
+mistakes occassionally. In most cases, the two algorithms maintained similarly
+acceptable fidelity levels in recreating the texture of the original rep-n
+input file.
+
+The actual compression ratio was respectable, with both Similarity and Cluster
+coming in at nearly half the size of the \Sequitur CFG.  Setting higher values
+for $\epsilon$ had diminishing returns for compression for larger and larger
+file sizes, at the cost of drastically lowering the fidelity of the output.
 
@@
 
-Repeating the above tests on the \emph{gen-n} datasets yields the results in table
-\ref{simclus_g_t}
+Repeating the above tests on the \emph{gen-n} datasets yields the results in
+table \ref{simclus_g_t}.
 
@@
 
-Just as in the rep-n results, the two algorithms have 
-almost identical compression ratios and
-their fidelity is about the same both subjectively and based on levenshtein
-distance. For an example of the text generated on one of the gen-n datasets,
-see the appendix.
-
-The runtimes are noticeably worse than for the rep-n dataset. The
-similarity algorithm runs in roughly $n^{2.5}$ time while cluster
-runs in roughly $n^2$ time.
-
-Changing epsilon in this case had a large effect on both the compression
-ratio and the fidelity of the final output string.
+Just as in the rep-n results, the two algorithms have almost identical
+compression ratios and their fidelity is about the same both subjectively and
+based on levenshtein distance. For an example of the text generated on one of
+the gen-n datasets, see the appendix.
+
+The runtimes are noticeably worse than for the rep-n dataset. The similarity
+algorithm runs in roughly $n^{2.5}$ time while cluster runs in roughly $n^2$
+time.
+
+Changing epsilon in this case had a large effect on both the compression ratio
+and the fidelity of the final output string.
 
@@
 
-Both of the algorithms were able to achieve great additional compression
-ratios on top of \Sequitur of up to about 40 percent for the highly
-structured rep-n data, and 30 percent for the more loosely structured
-gen-n data without using the grammar reducer. The fidelity between
-the original and lossified strings wasn't great as measured by
-levenshtein distance, but the texture in both the rep-n and
-gen-n datasets was preserved well. Decreasing the $\epsilon$ threshold
-improved fidelity greatly for larger very structured data at little
-cost to the compression ratio, but hurt the compression ratio
-a lot for less structured data.
+Both of the algorithms were able to achieve great additional compression ratios
+on top of \Sequitur of up to about 40 percent for the highly structured rep-n
+data, and 30 percent for the more loosely structured gen-n data without using
+the grammar reducer. The fidelity between the original and lossified strings
+wasn't great as measured by levenshtein distance, but the texture in both the
+rep-n and gen-n datasets was preserved well. Decreasing the $\epsilon$
+threshold improved fidelity greatly for larger very structured data at little
+cost to the compression ratio, but hurt the compression ratio a lot for less
+structured data.
 
 The differences between the two algorithms lie almost entirely in their
-runtime. Both algorithms perform much worse on the less structued
-gen-n data, we believe because simplifications are made more incrementally
-and the grammar does not get dramatically simpler after a few iterations
-of our algorithm like it does for the rep-n data.
-Unfortunately the Similarity algorithm tends to consistently incur
-an overhead of about $O(\sqrt{n})$ on top of the runtime for the
-Cluster algorithm. Because of this we dismiss the Similarity algorithm
-as unscalable and do not test it further.
+runtime. Both algorithms perform much worse on the less structued gen-n data,
+we believe because simplifications are made more incrementally and the grammar
+does not get dramatically simpler after a few iterations of our algorithm like
+it does for the rep-n data.  Unfortunately the Similarity algorithm tends to
+consistently incur an overhead of about $O(\sqrt{n})$ on top of the runtime for
+the Cluster algorithm. Because of this we dismiss the Similarity algorithm as
+unscalable and do not test it further.
 
 \subsection{Grammar Reduction Impact}
-Although the grammars generated by Sequitur are irreducible,
-our lossifiers produce grammars which could potentially be
-compressed further (losslessly) using Reduction Rules without
-changing the output string. To test the cost and
-effectiveness of applying reduction rules, we looked at the
-rep-n and gen-n test cases again, this time focusing on the
-clustering algorithm for $\epsilon = .2$. This time we give the ratio
-on top of the simplified grammar produced by the clustering algorithm,
-and the time taken to run just the reduction rules.
+
+Although the grammars generated by \Sequitur are irreducible, our lossifiers
+produce grammars which could potentially be compressed further (losslessly)
+using Reduction Rules without changing the output string. To test the cost and
+effectiveness of applying reduction rules, we looked at the rep-n and gen-n
+test cases again, this time focusing on the clustering algorithm for $\epsilon
+= .2$. This time we give the ratio on top of the simplified grammar produced by
+the clustering algorithm, and the time taken to run just the reduction rules.
 
 When plotted on a log-log scale, the runtime for our straightforward
-implementation of the reduction
-rules appears to be superpolynomial. While modest additional gains of about 30 percent
-are possible for highly self-similar data, very few additional gains are
-possible for more diverse datasets such as gen-n. This appears to be because
-the only simplifications that are possible on diverse datasets don't lead to
-the wide-ranging repetitive substitutions which allow reduction rules to be allowed.
-The simplifications in gen-n tend to be more local.
-
-Because we were unable to get reduction rules to scale to larger datasets,
-we will not consider them further.
+implementation of the reduction rules appears to be superpolynomial. While
+modest additional gains of about 30 percent are possible for highly
+self-similar data, very few additional gains are possible for more diverse
+datasets such as gen-n. This appears to be because the only simplifications
+that are possible on diverse datasets don't lead to the wide-ranging repetitive
+substitutions which allow reduction rules to be allowed.  The simplifications
+in gen-n tend to be more local.
+
+Because we were unable to get reduction rules to scale to larger datasets, we
+will not consider them further.
 
@@
 
-Though we were ultimately not able to come up with a good way to quantify how much of the
-structure of a text was preserved, we have included the final output strings we obtain
-after running the Clustering Algorithm with $\epsilon = .4$ in the Appendices. We
-believe these tests serve as a good (subjective) summary of the strengths and
-weaknesses of our
-algorithm in compressing data while preserving
-its character.
-
-The first piece of text is from the gen-n dataset again, the first 256000 characters
-of the book of Genesis. As listed in table \ref{seqclus_g_t}, the compressed grammar
-was size 5540 compared with Sequitur's size of 8060. Without knowledge of English 
-spelling, the Cluster algorithm produces output that is hard to read. However,
-key words are clearly recognizable, especially important words such as `God', and
-the overall structure in the book of Genesis is retained.
+Though we were ultimately not able to come up with a good way to quantify how
+much of the structure of a text was preserved, we have included the final
+output strings we obtain after running the Clustering Algorithm with $\epsilon
+= .4$ in the Appendices. We believe these tests serve as a good (subjective)
+summary of the strengths and weaknesses of our algorithm in compressing data
+while preserving its character.
+
+The first piece of text is from the gen-n dataset again, the first 256000
+characters of the book of Genesis. As listed in table \ref{simclus_g_t}, the
+compressed grammar was size 5540 compared with \Sequitur's size of 8060.
+Without knowledge of English spelling, the Cluster algorithm produces output
+that is hard to read. However, key words are clearly recognizable, especially
+important words such as `God', and the overall structure in the book of Genesis
+is retained.
 
 The second piece of text is one of the entries to this year's International
-Obfuscated C code contest, at \url{http://www.ioccc.org/2012/deckmyn/deckmyn.c} .
-The size of the lossified grammar is 1576 compared with the original size of
+Obfuscated C code contest, at \url{http://www.ioccc.org/2012/deckmyn/deckmyn.c}
+.  The size of the lossified grammar is 1576 compared with the original size of
 1973. The intricate whitespacing was lost, but the clear delineation into
@@
 \section{Conclusions}
+
 \subsection{Summary of Results}
+
 * General concept of simplifying grammars, two proposed algorithms

### 2012-12-11 tex/paper.tex
Fixing up intro and formatting title page

@@
 \newcommand{\Sequitur}{\textsc{Sequitur}\xspace}
+\newcommand{\Similarity}{\textsc{Similarity}\xspace}
+\newcommand{\Cluster}{\textsc{Cluster}\xspace}
 
@@
 
-\title{Squint: Lossy Hierarchical Compression on Symbol Streams}
+\title{%
+  \vspace{-.2in}%
+  Squint: Lossy Hierarchical Compression on Symbol Streams%
+  \vspace{-.1in}%
+}
 \author{Edward Gan \& Max Wang}
-\date{December 11, 2012}
+\date{%
+  \vspace{-.1in}%
+  December 11, 2012%
+}
+
 \maketitle
@@
 streams of symbolic data to achieve good compression ratios.  We develop a set
-of algorithms for improving upon Grammar-based compression schemes by allowing
-some loss in the compression. Although precise subsequences might be altered,
-in return we can simplify the grammars produced to obtain better ratios while
-still preserving important syntatic patterns at both the micro and macro
-levels.
+of algorithms for improving upon grammar-based compression schemes by allowing
+some lossiness in the compression.  Although precise subsequences might be
+altered, in return we can simplify the grammars produced to obtain better
+ratios while still preserving important syntatic patterns at both the micro and
+macro levels.
 
@@
 Many sources of data exhibit natural hierarchical structure.  For instance,
-english text often can be grouped well into sections, paragraphs, sentences,
-...  all the way down to words and phonemes. Music too, consists of repeated
-motifs, phrases, themes, and so on. Finally, computer code is so hierarchical
-that its syntax can usually be put in BNF. This allows grammar-based
-compression schemes to build up Context Free Grammars (CFGs) and achieve very
-good compression ratios.
-
-Taking a step back, in the image compression world there are both lossless and
-lossy compression schemes. In certain situations, such as for low-pixel count
-logos, one could use a lossless image compression scheme. In other situations
-however, such as for textures and some photographs, the exact pixel values are
-almost irrelevant compared to the overall gradients, colors, and structures in
-the image, and one could use a lossy compression scheme like jpeg.
-
-We believe it is valuable to bring some of the attitudes prevalent in image/
-video compression to the world of hierarchical text compression. Better
-compression ratios for text are possible.
-
-It may seem strange to allow a string such as "mind" to be lossily corrupted
-into ``rind'', but if one is more interested in recurrent structure in the text
-than in its precise meaning, these changes aren't important. For instance, if
-one is looking at a text in a foreign language, one is more likely to pick up
-the texture, the repeated sounds, some phrases that come up and the breaks
-between larger scale sections.  A change in a single letter probably wouldn't
-be noticed. Similarly, when skimming through huge files, perhaps execution
-logs, one usually first starts looking for patterns or things that stand out.
+English text often can be sensibly grouped into sections, paragraphs,
+sentences, and so on, all the way down to words and phonemes.  Likewise, music
+consists of repeated motifs, phrases, themes, etc., which make up musical
+structure.  Programming languages are so hierarchical that their syntax can
+usually be represented in BNF form.  It is natural, then, to represent these
+types of data using Context-Free Grammars (CFGs), which capture these
+hierarchical structures; this form of \emph{grammar-based compression} can be
+used to achieve very good compression ratios.
+
+In the parallel world of image compression, however, two different classes of
+compression schemes exist---\emph{lossless compression} and \emph{lossy
+compression}.  In lossy compression schemes, rather than representing data
+exactly, we allow a certain amount of corruption in order to achieve better
+compression ratios.  Some situations may demand lossless encoding---say, in the
+case of a corporation's trademarked graphical signature.  Many other
+situations, however, may have less stringent requirements.  For texture
+patterns and some photographs, the exact pixel values are far less important
+than the overall gradients, colors, and structures in the image.  In such
+cases, lossy compression schemes, such as JPEG for images \cite{jpeg}, can be
+valuable and effective.
+
+We believe it is worthwhile to introduce these notions of lossy compression
+that are prevalent in image/video compression into the world of hierarchical
+text compression.  Introducing lossiness to text compression opens up the
+possibility for better compression ratios.
+
+It may seem strange at first to allow a string such as "mind" to be lossily
+corrupted into, say, ``rind''.  If, however, one is more interested in
+recurrent structure in the text than in its precise semantic meaning, these
+corruptions are unimportant.  For instance, when looking at text in a foreign
+language, one is more likely to pick up the texture of the passage, the
+repeated sounds, common phrases, and breaks between larger-scale sections of
+text.  In this situation, a change of a single letter would probably go
+unnoticed.  Similarly, when skimming through a huge file, say, an execution
+log, an administrator usually begins by looking for patterns that stand out.
 Lorem ipsum is a perfect example of a piece of ``text'' whose value lies in its
-texture and sentence/paragraph structure, and we will come back to it again.
-
-In this paper we develop 2 algorithms and implement a system for lossily
-compressing text while preserving its hierarchical structures. The goal was to
-preserve the feel of a text at both micro and macro levels.  Since
-grammar-based compression schemes were designed to capture these structures, we
-built our system on top of the existing \Sequitur grammar inference algorithm.
-Starting with the CFG generated by a lossless grammar inference algorithm, we
-repeatedly simplify it, throwing out details along the way to emphasize
-repetitions. Our main contribution lies in the development, implementation, and
-evaluation of 2 algorithms for ``simplifying'' CFG grammars while preserving
-their structure.
+texture and sentence/paragraph structure, and we will return to it in later
+sections.
+
+In such situations, lossy hierarchical compression schemes may prove beneficial
+if they can provide better compression ratios.  Moreover, introducing lossiness
+to the CFGs which represent the input stream may accentuate more important
+parts of the stream's structure.  Compression of a musical passage might, for
+instance, capture important motives in favor of showing every rhythm and note.
+Similarly, compression of text in a particular language may highlight
+syntactical features.  We consider this a relevant potential use of lossy
+hierarchical compression.
+
+In this paper, we develop two algorithms and implement a broader system for
+lossily compressing text while preserving its hierarchical structures.  The
+goal was to preserve the feel of the text at both micro and macro levels.
+Since grammar-based compression schemes were designed to capture these
+structures, we built our system on top of the existing \Sequitur grammar
+inference algorithm.  Starting with the CFG generated by a lossless grammar
+inference algorithm, we repeatedly simplify it, throwing out details along the
+way to emphasize repetitions.  Our main contribution lies in the development,
+implementation, and evaluation of two algorithms for ``simplifying'' CFG
+grammars while attempting to preserve their structure.
 
@@
 algorithm would be to try to replace similar variables in parallel. This would
-avoid many of the repeat pair-comparisons done in \emph{Similarity}. In other
+avoid many of the repeat pair-comparisons done in \Similarity. In other
 words, the algorithm below puts similar variables into equivalence class
@@
 
-Suppose for the sake of contradiction that the Similarity algorithm created a
+Suppose for the sake of contradiction that the \Similarity algorithm created a
 cycle at one step by redirecting an edge originally pointing at $A$ to $B$.
@@
 
-For the Cluster algorithm the analysis is a bit more involved, since to avoid
+For the \Cluster algorithm the analysis is a bit more involved, since to avoid
 recomputation we only calculate the sizes of the variables once before doing
@@
 
-Suppose then that we created a cycle during Cluster by redirecting an edge from
+Suppose then that we created a cycle during \Cluster by redirecting an edge from
 $X \rightarrow A$ to $X \rightarrow B$.  This means that there must have been a
@@
 Given an input of length $n$, let $r$ be the number of rules generated by
-\Sequitur from the input.  In the analysis of the Similarity algorithm, let $R$
+\Sequitur from the input.  In the analysis of the \Similarity algorithm, let $R$
 be the total number of variable replacements we perform. For each replacement,
@@
 
-The inner loops of the Cluster are similar to those of the Similarity
+The inner loops of the \Cluster are similar to those of the \Similarity
 algorithm.  We iterate the entire clustering and replacement procedure $C$
@@
 relatively complicated inferred grammar with 13 rules and a size of 51. Both
-the Similarity and Clustering Algorithms with $\epsilon = .4$ , yield the
+the \Similarity and \Cluster algorithms with $\epsilon = .4$ , yield the
 following grammar after reduction. It has 5 rules and is almost half the size,
@@
 For our first suite of tests, we compare the effectiveness of the
-\emph{Similarity} and \emph{Cluster} algorithms. We might expect Similarity to
+\Similarity and \Cluster algorithms. We might expect \Similarity to
 achieve better quality at the same compression ratio since it updates the state
@@
 must update the grammar it has the potential to run much more slowly than
-Cluster.
+\Cluster.
 
@@
 In table \ref{simclus_r_t} we report the results on running \Sequitur and
-Cluster, without the grammar reductions, on the rep-n datasets. Time is the
+\Cluster, without the grammar reductions, on the rep-n datasets. Time is the
 wall clock running as measured on a 2.26 Ghz Intel Core 2 Duo Macbook Pro.
@@
 The columns of the data are prefixed by S or T depending on whether they are
-measurements for Similarity or Cluster respectively.
+measurements for \Similarity or \Cluster respectively.
 
@@
 
-The actual compression ratio was respectable, with both Similarity and Cluster
+The actual compression ratio was respectable, with both \Similarity and \Cluster
 coming in at nearly half the size of the \Sequitur CFG.  Setting higher values
@@
 does not get dramatically simpler after a few iterations of our algorithm like
-it does for the rep-n data.  Unfortunately the Similarity algorithm tends to
+it does for the rep-n data.  Unfortunately the \Similarity algorithm tends to
 consistently incur an overhead of about $O(\sqrt{n})$ on top of the runtime for
-the Cluster algorithm. Because of this we dismiss the Similarity algorithm as
+the \Cluster algorithm. Because of this we dismiss the \Similarity algorithm as
 unscalable and do not test it further.
@@
 much of the structure of a text was preserved, we have included the final
-output strings we obtain after running the Clustering Algorithm with $\epsilon
+output strings we obtain after running the \Cluster algorithm with $\epsilon
 = .4$ in the Appendices. We believe these tests serve as a good (subjective)
@@
 compressed grammar was size 5540 compared with \Sequitur's size of 8060.
-Without knowledge of English spelling, the Cluster algorithm produces output
+Without knowledge of English spelling, the \Cluster algorithm produces output
 that is hard to read. However, key words are clearly recognizable, especially

### 2012-12-12 tex/paper.tex
Various edits up through 3.2

@@
 \usepackage[letterpaper,text={7in,9in},centering]{geometry}
+\usepackage{tocloft}
 \usepackage{algorithm}
@@
 
+%\setlength{\cftbeforesecskip}{0pt}
+
 \newcommand{\Sequitur}{\textsc{Sequitur}\xspace}
@@
 In summary, established work on grammar inference algorithms provides a solid
-set of tools for experimenting with the kinds of local modifications we propose
-to make to improve grammar structure by introducing lossiness.
+set of tools for experimenting with the kinds of local lossy modifications we
+propose to make to simplify grammar structure.
 
@@
 format uses a similar technique in order to compress images lossily
-\cite{jpeg}. Fractal based compression takes this a step further and represents
-parts of a signal using self-similarity, encoding an image for instance as a
-fixed point of contraction maps. This is similar in spirit to what is done in
-MPEG-1, where parts of one frame can be transformed to encode parts of the next
-frame.
-
-Work has also been done regarding the lossy compression of particular input
-formats.  For text, methods such as replacing words with synonyms
-\cite{semanticlossy} or reordering the middle letters of a word (i.e., all
-letters save for the first and last) \cite{semilossless} have been explored.
-However, these are very specific tricks to optimize text, and not more broadly
-applicable to compressing hierarchical streams.
+\cite{jpeg}.  Fractal based compression takes this a step further and
+represents parts of a signal using self-similarity, encoding an image for
+instance as a fixed point of contraction maps.  This is similar in spirit to
+what is done in MPEG-1, where parts of one frame can be transformed to encode
+parts of the next frame.
+
+Work has also been done regarding the lossy compression of other input formats.
+For text, methods such as replacing words with synonyms \cite{semanticlossy} or
+reordering the middle letters of a word (i.e., all letters save for the first
+and last) \cite{semilossless} have been explored.  However, these are very
+specific tricks to optimize text, and not more broadly applicable to
+compressing hierarchical streams.
 
@@
 with a lossless grammar inference algorithm.  By building off of a grammar
-inference algorithm, we hope to take advantage of some of the partial
-structures that are visible in a CFG representation of a stream.  In this
-project, we chose \Sequitur, but any way of obtaining a grammar could be used.
-Then, transformed the grammar, introducing loss and perform compressions by
-simplifying the grammar. Thus our compression system is a multi-step process:
+inference algorithm, we hope to take advantage of some of the structures that
+are visible in a CFG representation of a stream.  In this project, we chose
+\Sequitur, but any method of inferring a grammar could be used.  After
+inference, we simplified the resulting grammar, introducing loss and obtaining
+greater compression.  Thus our full lossy grammar compression system is a
+multi-step process:
 
 \begin{itemize}
-  \item Infer a CFG representation of a stream (possibly using \Sequitur)
-  \item Simplify the CFG, possibly introducing loss
-  \item Output the CFG
+  \item Infer a CFG representation of a stream (we use \Sequitur).
+  \item Simplify the CFG, introducing loss.
+  \item Encode and output the CFG.
 \end{itemize}
 
-Our core new contribution comes in step two, in what we shall call a
-\emph{lossifier} algorithm, i.e. an algorithm that takes an admissible CFG and
-simplifies it. The output CFG should be smaller than the input CFG but generate
-a string which preserves the structure in the original string as much as
-possible.
+Our core contribution comes in step two, via what we refer to as a
+\emph{lossifier} algorithm---an algorithm that takes an admissible CFG and
+structurally simplifies it.  The output CFG should be smaller than the input
+CFG but generate a string which preserves the structure in the original string
+as much as possible.  In this paper, we focus on implementing and evaluating
+this step, and hence omit step 3 altogether.
 
@@
 closely as possible with an existing set of primitives.  In particular, just as
-with Mpeg and fractal compression, to approximate one part of the grammar we
-can use \emph{other parts of the grammar}. This is the seed for both of our
-algorithms. Given a rule $S \rightarrow rhs$, we will try to replace occurences
-of $S$ with other similar variables.
-
-For example, consider the grammars in figure \ref{sred}.
+with MPEG and fractal compression, to approximate one part of the grammar we
+can use \emph{other parts of the grammar}.  This is the core principle behind
+both of our algorithms.  Given a rule $A \rightarrow \alpha_1, \ldots,
+\alpha_n$, we will try to replace occurences of $A$ with other similar
+production rules.
+
+For example, consider the first grammar given in \autoref{sred}.  In this
+graphical representation, we omit the ordering of symbols on the right-hand
+side of production rules.  The picture makes it clear, however, that one easy
+route for simplifying the grammar is to simply replace the rule B with A.  They
+differ in their expansions by only a single character, `e', and they form
+almost identical halves of S.
 
@@
 
-To simplify the grammar, it would make sense to replace B with A, since they
-form two almost identical halves of S and are similar.
-
 \subsection{Pair-Similarity}
@@
 \rightarrow rhs_1$ and $V_2 \rightarrow rhs_2$ are similar when
-
 \[
@@
 \]
-
-where $levenshtein$ is the levenstein edit distance between two strings.  For
+where $levenshtein$ is the Levenstein edit distance between two strings.  For
 convenience we can define.
@@
 \[ lsim(s1,s2) = \frac{levenshtein(s1,s2)}{max(len(s1),len(s2))} \]
-
-The first algorithm takes the idea of variable similarity and uses it to
-simplify the grammar sequentially as much as possible.
+Our first algorithm takes the idea of variable similarity and uses it to
+iteratively simplify the grammar.
 

### 2012-12-12 tex/paper.tex
Future Work section

@@
       \If{$var_1$ hasn't been added to a cluster}
-        \State Add $var_1$ to a new cluster in $clusters$
+        \State add $var_1$ to a new cluster in $clusters$
       \EndIf
@@
 \subsection{Future Work}
-* Better ways to set epsilon to maximize value / cost for different size and types
-* Better ways to quantify preservation of hierarchies
-* How to find near `self-similarities' that aren't captured by a rule
-* Iterating lossify and reduce in a cycle
+
+There are many further directions to explore in the area of lossy grammar
+compression.  One immediate direction for future work is to develop additional
+lossifier algorithms, or to determine techniques for improving our existing
+ones.  One shortcoming of our \Cluster algorithm is that, by replacing clusters
+with the shortest rule, we lose both fidelity in the final expanded stream
+\emph{and} structure in the simplified grammar.  Replacing each cluster with
+the most frequently-appearing rule could improve this situation, at some small
+cost to the final compression ratio (small because all rules in a cluster are
+similar in size).  Unfortunately, we were unable to devise an efficient
+algorithm to perform this sort of replacement without accidentally introducing
+cycles.
+
+Investigating the intuition---or perhaps, providing some methodology---for
+choosing the parameter $\epsilon$ would also be useful.  Ideally, we could
+reliably choose $\epsilon$ in such a way that we maximize the ratio of
+compression to fidelity loss.  Investigating the effect that $\epsilon$ has for
+additional types of structured data might shed some light on this problem.
+
+Additional metrics for lossified grammars are also an important direction to
+consider.  Ultimately, our goal is to compress the \emph{hierarchical
+structure} of data, rather than the data itself, efficiently and with high
+fidelity.  However, we were only able to directly measure how well we preserved
+structure on contrived data, such as rep-n.  On real datasets, we instead made
+a first-order approximation by comparing the fully-expanded output streams.
+Some broadly-applicable metrics for comparing hierarchical structures would
+significantly improve the evaluability of lossy grammar compression techniques.
 

### 2012-12-12 tex/paper.tex
condensed genesis long output

--- 2012-12-12 tex/paper.tex
@@
 
-1:11 And God said, Let the earth bring forth grass, the herb yielding
-seed, and the fruit tree yielding fruit after his kind, whose seed is
-in itself, upon the earth: and it was so.
-
-1:12 And the earth brought forth grass, and herb yielding seed after
-his kind, and the tree yielding fruit, whose seed was in itself, after
-his kind: and God saw that it was good.
-
-1:13 And the evening and the morning were the third day.
-
-1:14 And God said, Let there be lights in the firmament of the heaven
-to divide the day from the night; and let them be for signs, and for
-seasons, and for days, and years: 1:15 And let them be for lights in
-the firmament of the heaven to give light upon the earth: and it was
-so.
-
-1:16 And God made two great lights; the greater light to rule the day,
-and the lesser light to rule the night: he made the stars also.
-
-1:17 And God set them in the firmament of the heaven to give light
-upon the earth, 1:18 And to rule over the day and over the night, and
-to divide the light from the darkness: and God saw that it was good.
-
-1:19 And the evening and the morning were the fourth day.
-
-1:20 And God said, Let the waters bring forth abundantly the moving
-creature that hath life, and fowl that may fly above the earth in the
-open firmament of heaven.
-
-1:21 And God created great whales, and every living creature that
-moveth, which the waters brought forth abundantly, after their kind,
-and every winged fowl after his kind: and God saw that it was good.
-
-1:22 And God blessed them, saying, Be fruitful, and multiply, and fill
-the waters in the seas, and let fowl multiply in the earth.
-
-1:23 And the evening and the morning were the fifth day.
-
-1:24 And God said, Let the earth bring forth the living creature after
-his kind, cattle, and creeping thing, and beast of the earth after his
-kind: and it was so.
-
-1:25 And God made the beast of the earth after his kind, and cattle
-after their kind, and every thing that creepeth upon the earth after
-his kind: and God saw that it was good.
-
-1:26 And God said, Let us make man in our image, after our likeness:
-and let them have dominion over the fish of the sea, and over the fowl
-of the air, and over the cattle, and over all the earth, and over
-every creeping thing that creepeth upon the earth.
-
 (Rest Omitted)
@@
 
-ine id, Lt e frargras, e ding
-seedn e rut dingrut fes inwhseseed i
-in tel, e frt wa
-
-n e frrurhtargras, n dingseedfes inn e dingrut , whseseeds in tel, fe
-s inine tht waod3e e frinwee e thir
-
-1e said, e behts ine mamene e
-todividthayfrm e htn e e e frsignd an
-seasond anayn ar: 1:n e e e frhts in
-e frmamene e  toinhte fr: n t wa
-
-6ine  tcrathts;e grHatoule thay
-nde estoule e hte e e e stars al
-
-7ine t e ine mamene e toinht
-e fr1:n toule e e ayn e e htd an
-todivide htfrm darknes: God tht waod9n e e n frinwee e fr
-
-0e said, binar aundanolye moilngcaledthhath in frwlthay flyabene frentoopen firmamenofmae
-
-1 n God Haedcratwhalesn enry lcalede
-moen, hiche wate brurhtar aundanoly, fe
-s inn enrys inedfrwlfes in: God th ts no
-
-n God es te m,saying, e rut fuln multiplyllwateine seasn t frwlmultiply in e fr
-
-3n e e n frinwee e fi
-
-e said, e frare lcaledfe
-s in, cattln crepn e frfehi
-int wa
-
-2n God Haede frfehis inn cattl
-fe
-s inn enry thinge frfes in: God th ts no
-
-6e said, usmae mae ur imae feur ienes:n e e mae dominon e e fish fe saan arfrwl
-e e n arcattln ar alle frn ar
-enrycrepe fr
-
 (Rest Omitted)

### 2012-12-12 tex/paper.tex
Make Appendix look passable

--- 2012-12-12 tex/paper.tex
@@
 
-\begin{table}[h]
-\subfloat[$\epsilon = .4$]{
-\begin{tabular}{l|l|l|l|l||l|l|l|l}
-Dataset & S-T(ime) & S-R(atio) & S-L(evdistance) & S-P(roportion) & C-T & C-R & C-L & C-P \\
-\hline
-rep-100   & 0.006 & 18/34    & 0.713 & 0.07 & 0.002 & 20/34    & 0.535 & 0.04\\
-rep-400   & 0.028 & 45/76    & 0.269 & 0.01 & 0.009 & 41/76    & 0.743 & 0.04\\
-rep-1600  & 0.355 & 103/194  & 0.873 & 0.03 & 0.042 & 103/194  & 0.804 & 0.02\\
-rep-6400  & 3.978 & 258/448  & 0.706 & 0.005& 0.203 & 256/448  & 0.869 & 0.50\\
-rep-12800 & 44.63 & 604/1064 & 0.791 & 0.01 & 1.136 & 600/1064 & 0.856 & 0.01\\
-rep-25600 & 135.0 & 900/1482 & 0.829 & 0.005& 2.906 & 898/1482 & 0.725 & 0.25\\
-\end{tabular}
+\begin{table}[H]
+  \centering
+  \subfloat[$\epsilon = .4$]{%
+  \begin{tabular}{l||l|l|l|l||l|l|l|l}
+    & \multicolumn{4}{c||}{\Similarity} & \multicolumn{4}{c}{\Cluster} \\ \hline
+    Dataset & Time & Ratio & Lev. & Prop. & Time & Ratio & Lev. & Prop. \\ \hline
+    rep-100   & 0.006 & 18/34    & 0.713 & 0.07 & 0.002 & 20/34    & 0.535 & 0.04 \\
+    rep-400   & 0.028 & 45/76    & 0.269 & 0.01 & 0.009 & 41/76    & 0.743 & 0.04 \\
+    rep-1600  & 0.355 & 103/194  & 0.873 & 0.03 & 0.042 & 103/194  & 0.804 & 0.02 \\
+    rep-6400  & 3.978 & 258/448  & 0.706 & 0.005& 0.203 & 256/448  & 0.869 & 0.50 \\
+    rep-12800 & 44.63 & 604/1064 & 0.791 & 0.01 & 1.136 & 600/1064 & 0.856 & 0.01 \\
+    rep-25600 & 135.0 & 900/1482 & 0.829 & 0.005& 2.906 & 898/1482 & 0.725 & 0.25 \\
+  \end{tabular}%
   }
 
-\subfloat[$\epsilon = .2$]{
-\begin{tabular}{l|l|l|l|l||l|l|l|l}
-Dataset & S-T(ime) & S-R(atio) & S-L(evdistance) & S-P(roportion) & C-T & C-R & C-L & C-P \\
-\hline
-rep-100   & 0.002 & 30/34    & 0.040 & 0.05 & 0.002 & 30/34    & 0.040 & 0.05\\
-rep-400   & 0.019 & 60/76    & 0.035 & 0.02 & 0.009 & 60/76    & 0.035 & 0.02\\
-rep-1600  & 0.350 & 123/194  & 0.355 & 0.03 & 0.055 & 123/194  & 0.206 & 0.01\\
-rep-6400  & 3.942 & 280/448  & 0.387 & 0.08 & 0.295 & 274/448  & 0.543 & 0.09\\
-rep-12800 & 42.21 & 638/1064 & 0.491 & 0.03 & 1.105 & 632/1064 & 0.457 & 0.01\\
-rep-25600 & 124.5 & 922/1482 & 0.565 & 0.12 & 2.597 & 926/1482 & 0.437 & 0.007\\
-\end{tabular}
+  \subfloat[$\epsilon = .2$]{%
+  \begin{tabular}{l||l|l|l|l||l|l|l|l}
+    & \multicolumn{4}{c||}{\Similarity} & \multicolumn{4}{c}{\Cluster} \\ \hline
+    Dataset & Time & Ratio & Lev. & Prop. & Time & Ratio & Lev. & Prop. \\ \hline
+    rep-100   & 0.002 & 30/34    & 0.040 & 0.05 & 0.002 & 30/34    & 0.040 & 0.05 \\
+    rep-400   & 0.019 & 60/76    & 0.035 & 0.02 & 0.009 & 60/76    & 0.035 & 0.02 \\
+    rep-1600  & 0.350 & 123/194  & 0.355 & 0.03 & 0.055 & 123/194  & 0.206 & 0.01 \\
+    rep-6400  & 3.942 & 280/448  & 0.387 & 0.08 & 0.295 & 274/448  & 0.543 & 0.09 \\
+    rep-12800 & 42.21 & 638/1064 & 0.491 & 0.03 & 1.105 & 632/1064 & 0.457 & 0.01 \\
+    rep-25600 & 124.5 & 922/1482 & 0.565 & 0.12 & 2.597 & 926/1482 & 0.437 & 0.007 \\
+  \end{tabular}%
   }
@@
 
-\begin{table}[h]
-\subfloat[$\epsilon = .4$]{
-\begin{tabular}{l|l|l|l||l|l|l}
-Dataset & S-T(ime) & S-R(atio) & S-L(evdistance) & C-T & C-R & C-L \\
-\hline
-gen-100   & 0.006 & 87/89    & 0.04 &0.006 & 87/89    & 0.04\\
-gen-400   & 0.175 & 224/258  & 0.12 &0.069 & 219/258  & 0.19\\
-gen-1600  & 4.863 & 539/697  & 0.31 &0.732 & 529/697  & 0.33\\
-gen-6400  & 216.9 & 1740/2348& 0.41 &9.127 & 1694/2348& 0.43\\
-gen-12800 & 1812  & 3296/4523& 0.40 &29.17 & 3181/4523& 0.44\\
-gen-25600 & 12989 & 5823/8060& 0.43 &176.7 & 5540/8060& 0.50\\
-\end{tabular}
+\begin{table}[H]
+  \centering
+  \subfloat[$\epsilon = .4$]{%
+  \begin{tabular}{l||l|l|l||l|l|l}
+    & \multicolumn{3}{c||}{\Similarity} & \multicolumn{3}{c}{\Cluster} \\ \hline
+    Dataset & Time & Ratio & Lev. & Time & Ratio & Lev. \\ \hline
+    gen-100   & 0.006 & 87/89    & 0.04 &0.006 & 87/89    & 0.04 \\
+    gen-400   & 0.175 & 224/258  & 0.12 &0.069 & 219/258  & 0.19 \\
+    gen-1600  & 4.863 & 539/697  & 0.31 &0.732 & 529/697  & 0.33 \\
+    gen-6400  & 216.9 & 1740/2348& 0.41 &9.127 & 1694/2348& 0.43 \\
+    gen-12800 & 1812  & 3296/4523& 0.40 &29.17 & 3181/4523& 0.44 \\
+    gen-25600 & 12989 & 5823/8060& 0.43 &176.7 & 5540/8060& 0.50 \\
+  \end{tabular}%
   }
 
-\subfloat[$\epsilon = .2$]{
-\begin{tabular}{l|l|l|l||l|l|l}
-Dataset & S-T(ime) & S-R(atio) & S-L(evdistance) & C-T & C-R & C-L \\
-\hline
+  \subfloat[$\epsilon = .2$]{%
+  \begin{tabular}{l||l|l|l||l|l|l}
+    & \multicolumn{3}{c||}{\Similarity} & \multicolumn{3}{c}{\Cluster} \\ \hline
+    Dataset & Time & Ratio & Lev. & Time & Ratio & Lev. \\ \hline
     gen-100   & 0.004 & 89/89    & 0.00 &0.002 & 89/89    & 0.00 \\
@@
     gen-1600  & 1.292 & 664/697  & 0.03 &0.874 & 664/697  & 0.03 \\
-gen-6400  & 73.83 & 2173/2348& 0.05 &14.02 & 2171/2348& 0.05\\
-gen-12800 & 549.2 & 4197/4253& 0.05 &45.64 & 4186/4253& 0.05\\
-gen-25600 & 3939  & 7405/8060& 0.06 &224.9 & 7374/8060& 0.06\\
-\end{tabular}
+    gen-6400  & 73.83 & 2173/2348& 0.05 &14.02 & 2171/2348& 0.05 \\
+    gen-12800 & 549.2 & 4197/4253& 0.05 &45.64 & 4186/4253& 0.05 \\
+    gen-25600 & 3939  & 7405/8060& 0.06 &224.9 & 7374/8060& 0.06 \\
+  \end{tabular}%
   }
@@
 
-\begin{table}[h]
+\begin{table}[H]
+  \centering
   \subfloat[rep-n tests]{
   \begin{tabular}{l|l|l}
-Dataset & Time & Ratio \\
-\hline
+    Dataset & Time & Ratio \\ \hline
     rep-100   & 0.008 & 29/30 \\
@@
     rep-1600  & 0.132 & 96/123 \\
-rep-6400  & 1.145 & 153/186\\
-rep-12800 & 19.43 & 403/632\\
+    rep-6400  & 1.145 & 153/186 \\
+    rep-12800 & 19.43 & 403/632 \\
   \end{tabular}
@@
     \begin{tabular}{l|l|l}
-Dataset & Time & Ratio \\
-\hline
-gen-100   & 0.016 & 89/89\\
-gen-400   & 0.126 & 245/245\\
-gen-1600  & 4.833 & 660/664\\
-gen-6400  & 308.7 & 2139/2171\\
+    Dataset & Time & Ratio \\ \hline
+    gen-100   & 0.016 & 89/89 \\
+    gen-400   & 0.126 & 245/245 \\
+    gen-1600  & 4.833 & 660/664 \\
+    gen-6400  & 308.7 & 2139/2171 \\
     \end{tabular}
   }
-
   \caption{Reducer Tests for $\epsilon = .2$}
@@
 
-\pagebreak
-
-\subsection*{Deckmyn.c Input / Output:}
+\subsection*{deckmyn.c}
+
 \subsubsection*{Input:}

### 2012-12-12 tex/paper.tex
messing with results

--- 2012-12-12 tex/paper.tex
@@
 with both plenty of hierarchical structure, and also some messiness that most
-lossless compression schemes would not deal well with. Take the string below,
+lossless compression schemes would not deal well with. As an example,
+the string below is
 similar to what one might come across in a math puzzle book.
@@
 then adding some texture to the data by changing 1/30 of the characters
-uniformly at random to `*'. Call these test-cases \emph{rep-n} Another set of
+uniformly at random to `*'. These are the \emph{rep-n} test-cases.
+Another set of
 test data was constructed from the first n characters of the book of Genesis
-from the KJV version of the Bible.  Call these test-cases \emph{gen-n}
+from the KJV version of the Bible. Call these test-cases \emph{gen-n}.
 
@@
 must update the grammar it has the potential to run much more slowly than
-\Cluster.
-
-We don't report the running times for just \Sequitur since its linear running
-time is negligible compared to our algorithms for all but the smallest input
-data.
-
-\subsubsection{Rep-n results}
+\Cluster. The running times for just \Sequitur aren't reported since its linear running
+time is negligible compared to our algorithms.
 
 In table \ref{simclus_r_t} we report the results on running \Sequitur and
-\Cluster, without the grammar reductions, on the rep-n datasets. Time is the
+\Cluster, without the grammar reductions, on the rep-n datasets. In table
+\ref{simclus_g_t} we do the same for the gen-n datasets.
+In these tests
+we wanted to evaluate the relative fidelity, compression ratios, and runtime 
+of the two algorithms.
+
+In the table, Time is the
 wall clock running as measured on a 2.26 Ghz Intel Core 2 Duo Macbook Pro.
@@
 lossifier was in recreating the original texture of the input.
-
 The columns of the data are prefixed by S or T depending on whether they are
@@
 
-The compression ratios achieved for both algorithms are roughly equal for the
+In both datasets, the 
+compression ratios achieved for both algorithms are roughly equal given the
 same value for $\epsilon$, so it is fair to compare the algorithms for the same
 $\epsilon$. 
-
 From plotting the data, we see that runtime of the algorithms are roughly
@@
 
-For both values of $\epsilon$, The clustering algorithm is nearly linear, while
+For both values of $\epsilon$, The runtime for the
+clustering algorithm is nearly linear, while
 the similarity algorithm is roughly quadratic. On this dataset, the leveshtein
@@
 
-The actual compression ratio was respectable, with both \Similarity and \Cluster
-coming in at nearly half the size of the \Sequitur CFG.  Setting higher values
-for $\epsilon$ had diminishing returns for compression for larger and larger
+As mentioned earlier,
+the actual compression ratio was respectable, with both \Similarity and \Cluster
+coming in at nearly half the size of the \Sequitur CFG. 
+However, setting higher values
+for $\epsilon$ had diminishing returns for compression for larger
 file sizes, at the cost of drastically lowering the fidelity of the output.
-
-\subsubsection{Gen-n results}
-
-Repeating the above tests on the \emph{gen-n} datasets yields the results in
-table \ref{simclus_g_t}.
 
@@
 
-Just as in the rep-n results, the two algorithms have almost identical
+In line he two algorithms have almost identical
 compression ratios and their fidelity is about the same both subjectively and

### 2012-12-12 tex/paper.tex
Number numbered things

@@
 
-\begin{itemize}
+\begin{enumerate}
   \item Infer a CFG representation of a stream (we use \Sequitur).
@@
   \item Encode and output the CFG.
-\end{itemize}
+\end{enumerate}
 
@@
 
-Overview
+In this section, we discuss some of the details of our implementation for
+lossifying and reducing CFGs.
+
+\subsection{Data Structures and Optimizations}
+
+Grammar Data Structure
+Cluster Data Structure
+Invariants Maintained
+Core utility methods??
+Linked Lists vs Arrays, Caching Expansions, Strings vs Symbols for expansion, Native Levenshtein
 
@@
 grammars smaller.
-
-\subsection{Data Structures and Optimizations}
-
-Grammar Data Structure
-Cluster Data Structure
-Invariants Maintained
-Core utility methods??
-Linked Lists vs Arrays, Caching Expansions, Strings vs Symbols for expansion, Native Levenshtein
 
@@
 
+\subsubsection{Methodology}
+
 For most of our benchmark tests, we focused on two extreme kinds of data: one
@@
 
-For our first suite of tests, we compare the effectiveness of the
+For our first suite of tests, we compared the effectiveness of the
 \Similarity and \Cluster algorithms. We might expect \Similarity to
@@
 must update the grammar it has the potential to run much more slowly than
-\Cluster. The running times for just \Sequitur aren't reported since its linear running
-time is negligible compared to our algorithms.
+\Cluster.
 
@@
 In these tests
-we wanted to evaluate the relative fidelity, compression ratios, and runtime 
+we were looking to evaluate the relative fidelity, compression ratios, and runtime 
 of the two algorithms.
 
-In the table, Time is the
+In the tables, Time is the
 wall clock running as measured on a 2.26 Ghz Intel Core 2 Duo Macbook Pro.
@@
 lossifier was in recreating the original texture of the input.
-The columns of the data are prefixed by S or T depending on whether they are
-measurements for \Similarity or \Cluster respectively.
+
+\subsubsection{Result Summary}
 
@@
 same value for $\epsilon$, so it is fair to compare the algorithms for the same
-$\epsilon$. 
+values of $\epsilon$. 
 From plotting the data, we see that runtime of the algorithms are roughly
@@
 runtime $r(n)=n^\alpha$ where $n$ is the length of the input string. In the
-chart below, all of the linear regressions had $R^2$ values of over $0.98$.
+charts below, all of the linear regressions had $R^2$ values of over $0.98$.
 
 \begin{tabular}{l|l|l|l|l}
-Data     & Sim: $\epsilon=.4$ & Sim: $\epsilon=.2$ & Clust: $\epsilon = .4$ & Clust: $\epsilon = .2$ \\
+rep-n runtime & Sim: $\epsilon=.4$ & Sim: $\epsilon=.2$ & Clust: $\epsilon = .4$ & Clust: $\epsilon = .2$ \\
 Exponent      & $1.859$ & $2.026$ & $1.306$ & $1.305$
@@
 
-For both values of $\epsilon$, The runtime for the
-clustering algorithm is nearly linear, while
-the similarity algorithm is roughly quadratic. On this dataset, the leveshtein
-distances were similar for both algorithms but fairly high since the length of
-the final string could change drastically.  The proportion of '*'s in the input
+\begin{tabular}{l|l|l|l|l}
+gen-n runtime & Sim: $\epsilon=.4$ & Sim: $\epsilon=.2$ & Clust: $\epsilon = .4$ & Clust: $\epsilon = .2$ \\
+Exponent      & $2.629$ & $2.493$ & $1.814$ & $2.061$
+\end{tabular}
+
+In general, \Cluster has a runtime which is slightly better than quadratic,
+while \Similarity has a runtime which is slightly worse than quadratic.
+Differant values of $\epsilon$ have comparatively smaller effects on the
+runtime, and both algorithms performed worse on the gen-n dataset. We think
+this is because there are more tiny substitutions one can perform on
+the gen-n dataset, whereas the rep-n dataset allows for much larger scale
+substitutions which eliminate the need for many small substitutions.
+As expected, \Cluster has a significantly better runtime than \Similarity,
+but it is interesting that both algorithms performed much better than our
+worst case analysis, we believe it is for the same reasons we proposed
+in our runtime analysis section.
+
+Across all of the tests, the leveshtein
+distances were comparable for both algorithms. Though they differed in
+individual cases, no algorithm did consistently better than the other.
+In general the levenshtein distances were high, except in the $\epsilon=.2$
+case for the gen-n data where very little compression was possible at all.
+We believe most of the distance could be explained by
+the fact that the algorithms could drastically shorten
+the strings, since it always replaces variables with \emph{smaller} variables
+as discussed earlier.
+The proportion of '*'s in the rep-n input
 was usually the right order of magnitude, with both algorithms making big
-mistakes occassionally. In most cases, the two algorithms maintained similarly
+mistakes occassionally. In most cases, we felt that the two algorithms maintained similarly
 acceptable fidelity levels in recreating the texture of the original rep-n
-input file.
-
-As mentioned earlier,
-the actual compression ratio was respectable, with both \Similarity and \Cluster
-coming in at nearly half the size of the \Sequitur CFG. 
+input file, and we discuss this subjective impression further in the
+Larger Tests section.
+
+The actual compression ratio was respectable in most tests, with both \Similarity and \Cluster
+coming in at nearly half the size of the \Sequitur CFG for the rep-n data,
+and 70 percent for the gen-n data. A notable exception was
+the gen-n test with $\epsilon=.2$, where $\epsilon$ was so small that barely
+any substitutions were possible. 
 However, setting higher values
@@
 file sizes, at the cost of drastically lowering the fidelity of the output.
-
-Just as with the rep-n data, log-log regression tests yield the following
-exponents for the runtime on the gen-n data, again with very high values of
-$R^2$.
-
-\begin{tabular}{l|l|l|l|l}
-Data     & Sim: $\epsilon=.4$ & Sim: $\epsilon=.2$ & Clust: $\epsilon = .4$ & Clust: $\epsilon = .2$ \\
-Exponent & $2.629$ & $2.493$ & $1.814$ & $2.061$
-\end{tabular}
-
-In line he two algorithms have almost identical
-compression ratios and their fidelity is about the same both subjectively and
-based on levenshtein distance. For an example of the text generated on one of
-the gen-n datasets, see the appendix.
-
-The runtimes are noticeably worse than for the rep-n dataset. The similarity
-algorithm runs in roughly $n^{2.5}$ time while cluster runs in roughly $n^2$
-time.
-
-Changing epsilon in this case had a large effect on both the compression ratio
-and the fidelity of the final output string.
-
-\subsubsection{Summary}
-
-Both of the algorithms were able to achieve great additional compression ratios
-on top of \Sequitur of up to about 40 percent for the highly structured rep-n
-data, and 30 percent for the more loosely structured gen-n data without using
-the grammar reducer. The fidelity between the original and lossified strings
-wasn't great as measured by levenshtein distance, but the texture in both the
-rep-n and gen-n datasets was preserved well. Decreasing the $\epsilon$
-threshold improved fidelity greatly for larger very structured data at little
-cost to the compression ratio, but hurt the compression ratio a lot for less
-structured data.
-
+Data with differing amuounts of repetition and noise
+appear to require different values of $\epsilon$ for the optimal fidelity/compression ratio.
+
+In summary, both of the algorithms were able to achieve moderate additional compression ratios
+on top of \Sequitur. The fidelity between the original and lossified strings
+wasn't great as measured by levenshtein distance, but we think the texture in both the
+rep-n and gen-n datasets was preserved well. Both fidelity and compression ratio were
+very sensitive to $\epsilon$ depending on the kind of data.
 The differences between the two algorithms lie almost entirely in their
-runtime. Both algorithms perform much worse on the less structued gen-n data,
-we believe because simplifications are made more incrementally and the grammar
-does not get dramatically simpler after a few iterations of our algorithm like
-it does for the rep-n data.  Unfortunately the \Similarity algorithm tends to
+runtime. Empirically the \Similarity algorithm tends to
 consistently incur an overhead of about $O(\sqrt{n})$ on top of the runtime for
@@
 produce grammars which could potentially be compressed further (losslessly)
-using Reduction Rules without changing the output string. To test the cost and
-effectiveness of applying reduction rules, we looked at the rep-n and gen-n
+using Reduction Rules without changing the output string. To test these rules, 
+we looked at the rep-n and gen-n
 test cases again, this time focusing on the clustering algorithm for $\epsilon
-= .2$. This time we give the ratio on top of the simplified grammar produced by
+= .2$. We report our results in table \ref{red_t}, where
+we give the ratio on top of the simplified grammar produced by
 the clustering algorithm, and the time taken to run just the reduction rules.
@@
 that are possible on diverse datasets don't lead to the wide-ranging repetitive
-substitutions which allow reduction rules to be allowed.  The simplifications
-in gen-n tend to be more local.
-
-Because we were unable to get reduction rules to scale to larger datasets, we
+substitutions which allow reduction rules to be allowed.
+
+Because we were unable to get our implementation of
+reduction rules to scale to larger datasets, we
 will not consider them further.
@@
 Though we were ultimately not able to come up with a good way to quantify how
-much of the structure of a text was preserved, we have included the final
-output strings we obtain after running the \Cluster algorithm with $\epsilon
+much of the structure of a text was preserved, we have included some final
+output strings obtained after running the \Cluster algorithm with $\epsilon
 = .4$ in the Appendices. We believe these tests serve as a good (subjective)
@@
 that is hard to read. However, key words are clearly recognizable, especially
-important words such as `God', and the overall structure in the book of Genesis
-is retained.
+important words such as `God', and the overall division of the book into
+verses was retained.
 
@@
 1973. The intricate whitespacing was lost, but the clear delineation into
-sections was retained, key words such as `define' and `for' are very visible,
-and the texture of the obfuscated C is very visible.
+sections was kept, key words such as `define' and `for' are very visible,
+and the texture of the obfuscated C is very obvious.
 

### 2012-12-12 tex/paper.tex
streamlined results section
Some Implementation section

@@
 \subsectionfont{\normalsize}
+\subsubsectionfont{\normalfont}
 
@@
 
+\setcounter{tocdepth}{2}
 \tableofcontents
@@
 
-In this section, we discuss some of the details of our implementation for
-lossifying and reducing CFGs.
+Our full grammar compression scheme involves the following sequence of
+operations and components:
+\begin{enumerate}
+  \item \Sequitur, which infers a CFG from the input stream.
+  \item Lossifier algorithms, which take the \Sequitur grammar and simplify it,
+    introducing lossiness.
+  \item Reduction engine, which applies Kieffer and Yang's reduction rules to
+    the lossified grammar.
+\end{enumerate}
+
+We implement all these components in the Ruby programming language.  Our
+implementation of \Sequitur exactly mirrors the data structures and algorithm
+described by Nevill-Manning and Witten \cite{sequitur}---and, empirically, it
+has the same asymptotic running time---so we do not discuss it further.
+However, we will examine in this section some of the techniques we used in
+implementing \Similarity, \Cluster, and the reduction engine.
 
@@
 
+\subsubsection{CFG Representation}
+
+Our representation of a context-free grammar is very similar to that used by
+\Sequitur.  All of our rules are represented by a linked list of symbols;
+however, unlike in \Sequitur, we omit refcounting and do not point from each
+instance of a nonterminal symbol to its corresponding rule.  Instead, our rules
+are kept in a hash table, mapping nonterminal symbols to a list representing
+its RHS.
+
+We chose to implement a list library rather than use Ruby's native arrays in
+order to achieve better runtime for the splice operations.  Both \Sequitur and
+our implementation of Kieffer and Yang's reductions require us to splice out a
+subsequence of terminals from a rule, or to insert a subsequence (e.g., as a
+result of inlining a singleton rule) into a rule.
+
 Grammar Data Structure
-Cluster Data Structure
-Invariants Maintained
-Core utility methods??
-Linked Lists vs Arrays, Caching Expansions, Strings vs Symbols for expansion, Native Levenshtein
+Linked Lists vs Arrays, Caching Expansions, Native Levenshtein
+
+\subsubsection{Lossifier Optimizations}
+
+Perhaps the most crucial optimization we implement deals with the recursive
+expansion of nonterminals into terminal strings.
 

### 2012-12-12 tex/paper.tex
Subsubsection format
Finishing Implementation section

@@
 \subsectionfont{\normalsize}
-\subsubsectionfont{\normalfont}
+\subsubsectionfont{\normalfont\itshape}
 
@@
 We chose to implement a list library rather than use Ruby's native arrays in
-order to achieve better runtime for the splice operations.  Both \Sequitur and
+order to achieve better runtime for the splice operation.  Both \Sequitur and
 our implementation of Kieffer and Yang's reductions require us to splice out a
 subsequence of terminals from a rule, or to insert a subsequence (e.g., as a
-result of inlining a singleton rule) into a rule.
-
-Grammar Data Structure
-Linked Lists vs Arrays, Caching Expansions, Native Levenshtein
+result of inlining a singleton rule) into a rule.  Moreover, our initial
+implementation of the \texttt{replace} operation---which replaces all
+occurrences of a nonterminal with another---initially employed rule inlining to
+prevent the creation of cycles.  Our first iteration implementation used native
+arrays, and we found that linked lists did result in an empirical improvement
+in runtime.
 
@@
 Perhaps the most crucial optimization we implement deals with the recursive
-expansion of nonterminals into terminal strings.
+expansion of nonterminals into terminal strings---the \texttt{expand}
+operation.  An extraordinarily na\"ive implementation, in which we simply
+recurse into every rule we find, could have runtime as bad as $O(\prod_{R \in
+G} |R|)$, the product of the lengths of all rules in the grammar $G$.  We can
+easily reduce this to $O(\sum_{R \in G} |R|)$ by caching the expansion of each
+nonterminal when we complete it, and using the cached result for later
+instances.
+
+We further leverage this caching optimization, however, by preserving the
+expansion cache across calls to \texttt{expand}.  This optimization is critical
+to achieving acceptable runtime for the \Similarity and \Cluster algorithms
+because of their heavy reliance on expanding.  Each iteration of the algorithm
+requires expanding all nonterminals.  Since \texttt{replac}ing a nonterminal A
+with B changes the expansion of all rules containing A, we clear the expand
+cache whenever replace is called.  This is the key reason for \Cluster's
+superior runtime relative to \Similarity---\Similarity needs to expand all
+nonterms for each individual \texttt{replace} operation, whereas \Cluster
+clusters \texttt{replace} operations, thereby reducing the number of calls and
+effective overhead of \texttt{expand}.
+
+In addition to this optimization, we also use a native C implementation of the
+Levenshtein function.  We do not compare this to any Ruby implementation, but
+we suspect it has some nontrivial impact on runtime due to the volume of calls
+to the function.
 
@@
 
-What is their purpose? Why do we want them? What are they?  Grammar reductions
-are a way of taking a CFG, generated possibly through ad-hoc techniques, and
-systematically removing internal redundancies like duplicate bigrams in
-\Sequitur. They make the CFG representation smaller but don't introduce any
-loss. We want them since our lossifier algorithms often make our grammars
-redundant, even though the each step we take in the lossifier makes the
-grammars smaller.
+Kieffer and Yang's grammar reductions are a collection of rules which, when
+applied iteratively to a grammar, systematically eliminate internal
+redundancies, such as duplicate symbol subsequences.  They often reduce the
+size of CFGs by factoring out repeated patterns as new rules and by eliminating
+unnecessary rules, but they introduce no loss.
+
+Since our lossification techniques make no guarantees about the form of the
+resultant grammar, unlike, say, \Sequitur, we produced a na\"ive implementation
+of these reduction rules in order to make our lossy grammars as small as
+possible.  As we shall see in the next section, this simple reduction engine
+has poor runtime and does not scale very well.
 

### 2012-12-12 tex/paper.tex
edits in intro

--- 2012-12-12 tex/paper.tex
@@
 compression schemes exist---\emph{lossless compression} and \emph{lossy
-compression}.  In lossy compression schemes, rather than representing data
-exactly, we allow a certain amount of corruption in order to achieve better
-compression ratios.  Some situations may demand lossless encoding---say, in the
-case of a corporation's trademarked graphical signature.  Many other
-situations, however, may have less stringent requirements.  For texture
-patterns and some photographs, the exact pixel values are far less important
+compression}.  In lossy compression schemes we allow a certain amount of 
+corruption in order to achieve better
+compression ratios.  Some situations may demand lossless encoding, but many others
+have less stringent requirements. With texture
+patterns and some photographs for instance, the exact pixel values are far less important
 than the overall gradients, colors, and structures in the image.  In such
@@
 that are prevalent in image/video compression into the world of hierarchical
-text compression.  Introducing lossiness to text compression opens up the
-possibility for better compression ratios.
+symbol-stream compression. Introducing lossiness to symbol-stream compression 
+opens up the possibility for better compression ratios.
 
@@
 
-The intution that grammar inference can support many detailed policies is made
+The intuition that grammar inference can support many detailed policies is made
 more formal in \cite{grammarcodes}.  The authors classify the properties a
@@
 
-There has also been very interesting work analyzing how grammar based
-algorithms compare with the theoretical best lossless grammar for a string, in
-the line of \cite{approximation}.  However, these analytic bounds for lossless
-grammars are outside the scope of our proposed research.
-
 In summary, established work on grammar inference algorithms provides a solid
-set of tools for experimenting with the kinds of local lossy modifications we
-propose to make to simplify grammar structure.
+set of tools for experimenting with the kinds of lossy modifications we
+propose to make to simplify grammar structure, and we will draw upon the kinds
+of grammar transformations used in \Sequitur and in Kieffer's reduction rules.
 
@@
 While we were unable to find work in the area of lossy grammar compression, a
-number of existing techniques for lossy hierarchical compression more broadly
-may be applicable.
+number of existing techniques for lossy compression more broadly serve
+as inspiration.
 
@@
 what is done in MPEG-1, where parts of one frame can be transformed to encode
-parts of the next frame.
+parts of the next frame. Many of these ideas, such as finding largest matching
+components of objects, make their way into our algorithms.
 
@@
 
-\begin{enumerate}
+\begin{itemize}
   \item Infer a CFG representation of a stream (we use \Sequitur).
@@
   \item Encode and output the CFG.
-\end{enumerate}
+\end{itemize}
 
@@
 We can view the CFG as a DAG where for variables $V_1$, $V_2$, we add an edge
-whenever $V_1$ contains $V_2$ in its rhs. When sizes are compuated on a CFG
-before we modify it, if $V_1 \rightarrow V_2$ then $size(V_1) > size(V_2)$.
+whenever $V_1$ contains $V_2$ in its rhs. When sizes are computed on a CFG
+before we modify it, if $V_1 \rightarrow V_2$ then $size(V_1) > size(V_2)$ since
+the rhs of $V_1$ contains $V_2$.
 Replacing a variable $V_1$ with $V_2$ has the same effect as redirecting all
@@
 $size(X)>size(A)>size(B)$ and transitivity will preserve the ordering.
-
 Suppose then that we created a cycle during \Cluster by redirecting an edge from
@@
 From the \Sequitur paper \cite{sequitur} we know that the number of rules in
-the grammar generated by \Sequitur is between $O(1)$ and $O(n)$. Thus the worst
+the grammar generated by \Sequitur is bounded by $O(n)$. Thus the worst
 case runtime on the similarity algorithm is $\boxed{O(n^4)}$. This is because
@@
 The inner loops of the \Cluster are similar to those of the \Similarity
-algorithm.  We iterate the entire clustering and replacement procedure $C$
-times. Going through each variable-cluster pair takes up to $O(r^2)$ time and
-calculating the levenshtein distance takes $O(n)$ time. The time taken to
-actually perform the replacements each iteration is absorbed into the $O(r^2)$
-term. So, again we have a runtime of $\boxed{O(C*r^2*n)}$. Just as with the
-similarity algorithm, a variety of factors make the runtime much better than
-the $\boxed{O(n^4)}$ worst case scenario.  The big gain over the similarity
-algorithm is that $C$ turns out to be nearly constant, since there are very few
-huge chains of reductions that are only possible after other reductions have
-been performed, an in fact we are able to perform a very large number of the
-remaining reductions each time we perform an iteration of the clustering
-algorithm. Thus a more reasonable upper bound for the runtime of cluster would
+algorithm, and an analogous analysis yields a runtime of $\boxed{O(C*r^2*n)}$. 
+Just as with the
+similarity algorithm however, a variety of factors make the runtime much better than
+the $\boxed{O(n^4)}$ worst case scenario.  The big gain over \Similarity
+is that $C$ turns out to be nearly constant, since there exist very few
+huge chains of reductions that must be performed in sequence
+so we are able to perform many reductions in parallel.
+Thus a more reasonable upper bound for the runtime of cluster would
 be $\boxed{O(n^3)}$
@@
 
-In this section, we discuss some of the details of our implementation for
-lossifying and reducing CFGs.
-
-\subsection{Data Structures and Optimizations}
-
-Grammar Data Structure
-Cluster Data Structure
-Invariants Maintained
-Core utility methods??
-Linked Lists vs Arrays, Caching Expansions, Strings vs Symbols for expansion, Native Levenshtein
+Overview
 
@@
 grammars smaller.
+
+\subsection{Data Structures and Optimizations}
+
+Grammar Data Structure
+Cluster Data Structure
+Invariants Maintained
+Core utility methods??
+Linked Lists vs Arrays, Caching Expansions, Strings vs Symbols for expansion, Native Levenshtein
 

### 2012-12-12 tex/paper.tex

@@
 \subsectionfont{\normalsize}
+\subsubsectionfont{\normalfont\itshape}
 
@@
 
-Overview
+Our full grammar compression scheme involves the following sequence of
+operations and components:
+\begin{enumerate}
+  \item \Sequitur, which infers a CFG from the input stream.
+  \item Lossifier algorithms, which take the \Sequitur grammar and simplify it,
+    introducing lossiness.
+  \item Reduction engine, which applies Kieffer and Yang's reduction rules to
+    the lossified grammar.
+\end{enumerate}
+
+We implement all these components in the Ruby programming language.  Our
+implementation of \Sequitur exactly mirrors the data structures and algorithm
+described by Nevill-Manning and Witten \cite{sequitur}---and, empirically, it
+has the same asymptotic running time---so we do not discuss it further.
+However, we will examine in this section some of the techniques we used in
+implementing \Similarity, \Cluster, and the reduction engine.
+
+\subsection{Data Structures and Optimizations}
+
+\subsubsection{CFG Representation}
+
+Our representation of a context-free grammar is very similar to that used by
+\Sequitur.  All of our rules are represented by a linked list of symbols;
+however, unlike in \Sequitur, we omit refcounting and do not point from each
+instance of a nonterminal symbol to its corresponding rule.  Instead, our rules
+are kept in a hash table, mapping nonterminal symbols to a list representing
+its RHS.
+
+We chose to implement a list library rather than use Ruby's native arrays in
+order to achieve better runtime for the splice operations.  Both \Sequitur and
+our implementation of Kieffer and Yang's reductions require us to splice out a
+subsequence of terminals from a rule, or to insert a subsequence (e.g., as a
+result of inlining a singleton rule) into a rule.
+
+Grammar Data Structure
+Linked Lists vs Arrays, Caching Expansions, Native Levenshtein
+
+\subsubsection{Lossifier Optimizations}
+
+Perhaps the most crucial optimization we implement deals with the recursive
+expansion of nonterminals into terminal strings.
 

### 2012-12-12 tex/paper.tex
Some formatting

@@
 compression}.  In lossy compression schemes we allow a certain amount of
-corruption in order to achieve better
-compression ratios.  Some situations may demand lossless encoding, but many others
-have less stringent requirements. With texture
-patterns and some photographs for instance, the exact pixel values are far less important
-than the overall gradients, colors, and structures in the image.  In such
-cases, lossy compression schemes, such as JPEG for images \cite{jpeg}, can be
-valuable and effective.
+corruption in order to achieve better compression ratios.  Some situations may
+demand lossless encoding, but many others have less stringent requirements.
+With texture patterns and some photographs for instance, the exact pixel values
+are far less important than the overall gradients, colors, and structures in
+the image.  In such cases, lossy compression schemes, such as JPEG for images
+\cite{jpeg}, can be valuable and effective.
 
@@
 In summary, established work on grammar inference algorithms provides a solid
-set of tools for experimenting with the kinds of lossy modifications we
-propose to make to simplify grammar structure, and we will draw upon the kinds
-of grammar transformations used in \Sequitur and in Kieffer's reduction rules.
+set of tools for experimenting with the kinds of lossy modifications we propose
+to make to simplify grammar structure, and we will draw upon the kinds of
+grammar transformations used in \Sequitur and in Kieffer's reduction rules.
 
@@
 While we were unable to find work in the area of lossy grammar compression, a
-number of existing techniques for lossy compression more broadly serve
-as inspiration.
+number of existing techniques for lossy compression more broadly serve as
+inspiration.
 
@@
 whenever $V_1$ contains $V_2$ in its rhs. When sizes are computed on a CFG
-before we modify it, if $V_1 \rightarrow V_2$ then $size(V_1) > size(V_2)$ since
-the rhs of $V_1$ contains $V_2$.
-Replacing a variable $V_1$ with $V_2$ has the same effect as redirecting all
-in-edges to $V_1$ onto $V_2$.
+before we modify it, if $V_1 \rightarrow V_2$ then $size(V_1) > size(V_2)$
+since the rhs of $V_1$ contains $V_2$.  Replacing a variable $V_1$ with $V_2$
+has the same effect as redirecting all in-edges to $V_1$ onto $V_2$.
 
@@
 redirect an edge from $X\rightarrow A$ to $X\rightarrow B$
-$size(X)>size(A)>size(B)$ and transitivity will preserve the ordering.
-Suppose then that we created a cycle during \Cluster by redirecting an edge from
-$X \rightarrow A$ to $X \rightarrow B$.  This means that there must have been a
+$size(X)>size(A)>size(B)$ and transitivity will preserve the ordering.  Suppose
+then that we created a cycle during \Cluster by redirecting an edge from $X
+\rightarrow A$ to $X \rightarrow B$.  This means that there must have been a
 path from $B$ to $X$. However, even in its intermediate state we know that the
@@
 we had to have searched though up to all pairs of variables, and for each pair
-of variables expanding them and calculating the levenshtein distance could take
+of variables expanding them and calculating the Levenshtein distance could take
 time linear in $n$.  Performing the replacement only takes time linear in $r$.
@@
 could grow much slower than $r$, and the use of native code in calculating the
-levenshtein distance made the its contribution much less noticeable for the
+Levenshtein distance made the its contribution much less noticeable for the
 input sizes we tested.
@@
 algorithm, and an analogous analysis yields a runtime of $\boxed{O(C*r^2*n)}$.
-Just as with the
-similarity algorithm however, a variety of factors make the runtime much better than
-the $\boxed{O(n^4)}$ worst case scenario.  The big gain over \Similarity
-is that $C$ turns out to be nearly constant, since there exist very few
-huge chains of reductions that must be performed in sequence
-so we are able to perform many reductions in parallel.
-Thus a more reasonable upper bound for the runtime of cluster would
-be $\boxed{O(n^3)}$
+Just as with the similarity algorithm however, a variety of factors make the
+runtime much better than the $\boxed{O(n^4)}$ worst case scenario.  The big
+gain over \Similarity is that $C$ turns out to be nearly constant, since there
+exist very few huge chains of reductions that must be performed in sequence so
+we are able to perform many reductions in parallel.  Thus a more reasonable
+upper bound for the runtime of cluster would be $\boxed{O(n^3)}$
 
@@
 with both plenty of hierarchical structure, and also some messiness that most
-lossless compression schemes would not deal well with. As an example,
-the string below is
-similar to what one might come across in a math puzzle book.
+lossless compression schemes would not deal well with. As an example, the
+string below is similar to what one might come across in a math puzzle book.
 
@@
 then adding some texture to the data by changing 1/30 of the characters
-uniformly at random to `*'. These are the \emph{rep-n} test-cases.
-Another set of
-test data was constructed from the first n characters of the book of Genesis
+uniformly at random to `*'. These are the \emph{rep-n} test-cases.  Another set
+of test data was constructed from the first n characters of the book of Genesis
 from the KJV version of the Bible. Call these test-cases \emph{gen-n}.
 
-For our first suite of tests, we compared the effectiveness of the
-\Similarity and \Cluster algorithms. We might expect \Similarity to
-achieve better quality at the same compression ratio since it updates the state
-of the grammar between each variable replacement, on the other hand since it
-must update the grammar it has the potential to run much more slowly than
-\Cluster.
+For our first suite of tests, we compared the effectiveness of the \Similarity
+and \Cluster algorithms. We might expect \Similarity to achieve better quality
+at the same compression ratio since it updates the state of the grammar between
+each variable replacement, on the other hand since it must update the grammar
+it has the potential to run much more slowly than \Cluster.
 
@@
 \Cluster, without the grammar reductions, on the rep-n datasets. In table
-\ref{simclus_g_t} we do the same for the gen-n datasets.
-In these tests
-we were looking to evaluate the relative fidelity, compression ratios, and runtime
+\ref{simclus_g_t} we do the same for the gen-n datasets.  In these tests we
+were looking to evaluate the relative fidelity, compression ratios, and runtime
 of the two algorithms.
 
-In the tables, Time is the
-wall clock running as measured on a 2.26 Ghz Intel Core 2 Duo Macbook Pro.
-Ratio is the relative size of the grammar produced compared with \Sequitur,
-where the size of the grammar is the sum of the lengths of the rhs's of the
-rules. Levdistance is the levenshtein distance between the output of the
-lossified grammar and the original input string, it is a kind of measure of the
-fidelity of our algorithms but doesn't take into account how well patterns were
-preserved. For the rep-n tests, proportion is the proportion of the final
-string produced that was `*'. The closer to 1/30~0.033, the more accurate the
-lossifier was in recreating the original texture of the input.
+In the tables, Time is the wall clock running as measured on a 2.26 Ghz Intel
+Core 2 Duo Macbook Pro.  Ratio is the relative size of the grammar produced
+compared with \Sequitur, where the size of the grammar is the sum of the
+lengths of the rhs's of the rules. Levdistance is the Levenshtein distance
+between the output of the lossified grammar and the original input string, it
+is a kind of measure of the fidelity of our algorithms but doesn't take into
+account how well patterns were preserved. For the rep-n tests, proportion is
+the proportion of the final string produced that was `*'. The closer to
+1/30~0.033, the more accurate the lossifier was in recreating the original
+texture of the input.
 
@@
 
-In both datasets, the
-compression ratios achieved for both algorithms are roughly equal given the
-same value for $\epsilon$, so it is fair to compare the algorithms for the same
-values of $\epsilon$.
-From plotting the data, we see that runtime of the algorithms are roughly
-polynomial in the size of the data, so we can perform a linear regression on
-the log-log transform of the data to find the exponents $\alpha$ for the
-runtime $r(n)=n^\alpha$ where $n$ is the length of the input string. In the
-charts below, all of the linear regressions had $R^2$ values of over $0.98$.
+In both datasets, the compression ratios achieved for both algorithms are
+roughly equal given the same value for $\epsilon$, so it is fair to compare the
+algorithms for the same values of $\epsilon$.  From plotting the data, we see
+that runtime of the algorithms are roughly polynomial in the size of the data,
+so we can perform a linear regression on the log-log transform of the data to
+find the exponents $\alpha$ for the runtime $r(n)=n^\alpha$ where $n$ is the
+length of the input string. In the charts below, all of the linear regressions
+had $R^2$ values of over $0.98$.
 
@@
 
-Across all of the tests, the leveshtein
-distances were comparable for both algorithms. Though they differed in
-individual cases, no algorithm did consistently better than the other.
-In general the levenshtein distances were high, except in the $\epsilon=.2$
-case for the gen-n data where very little compression was possible at all.
-We believe most of the distance could be explained by
-the fact that the algorithms could drastically shorten
-the strings, since it always replaces variables with \emph{smaller} variables
-as discussed earlier.
-The proportion of '*'s in the rep-n input
-was usually the right order of magnitude, with both algorithms making big
-mistakes occassionally. In most cases, we felt that the two algorithms maintained similarly
-acceptable fidelity levels in recreating the texture of the original rep-n
-input file, and we discuss this subjective impression further in the
+Across all of the tests, the Leveshtein distances were comparable for both
+algorithms. Though they differed in individual cases, no algorithm did
+consistently better than the other.  In general the Levenshtein distances were
+high, except in the $\epsilon=.2$ case for the gen-n data where very little
+compression was possible at all.  We believe most of the distance could be
+explained by the fact that the algorithms could drastically shorten the
+strings, since it always replaces variables with \emph{smaller} variables as
+discussed earlier.  The proportion of '*'s in the rep-n input was usually the
+right order of magnitude, with both algorithms making big mistakes
+occassionally. In most cases, we felt that the two algorithms maintained
+similarly acceptable fidelity levels in recreating the texture of the original
+rep-n input file, and we discuss this subjective impression further in the
 Larger Tests section.
 
-The actual compression ratio was respectable in most tests, with both \Similarity and \Cluster
-coming in at nearly half the size of the \Sequitur CFG for the rep-n data,
-and 70 percent for the gen-n data. A notable exception was
+The actual compression ratio was respectable in most tests, with both
+\Similarity and \Cluster coming in at nearly half the size of the \Sequitur CFG
+for the rep-n data, and 70 percent for the gen-n data. A notable exception was
 the gen-n test with $\epsilon=.2$, where $\epsilon$ was so small that barely
-any substitutions were possible.
-However, setting higher values
-for $\epsilon$ had diminishing returns for compression for larger
-file sizes, at the cost of drastically lowering the fidelity of the output.
-Data with differing amuounts of repetition and noise
-appear to require different values of $\epsilon$ for the optimal fidelity/compression ratio.
-
-In summary, both of the algorithms were able to achieve moderate additional compression ratios
-on top of \Sequitur. The fidelity between the original and lossified strings
-wasn't great as measured by levenshtein distance, but we think the texture in both the
-rep-n and gen-n datasets was preserved well. Both fidelity and compression ratio were
-very sensitive to $\epsilon$ depending on the kind of data.
-The differences between the two algorithms lie almost entirely in their
-runtime. Empirically the \Similarity algorithm tends to
+any substitutions were possible.  However, setting higher values for $\epsilon$
+had diminishing returns for compression for larger file sizes, at the cost of
+drastically lowering the fidelity of the output.  Data with differing amuounts
+of repetition and noise appear to require different values of $\epsilon$ for
+the optimal fidelity/compression ratio.
+
+In summary, both of the algorithms were able to achieve moderate additional
+compression ratios on top of \Sequitur. The fidelity between the original and
+lossified strings wasn't great as measured by Levenshtein distance, but we
+think the texture in both the rep-n and gen-n datasets was preserved well. Both
+fidelity and compression ratio were very sensitive to $\epsilon$ depending on
+the kind of data.  The differences between the two algorithms lie almost
+entirely in their runtime. Empirically the \Similarity algorithm tends to
 consistently incur an overhead of about $O(\sqrt{n})$ on top of the runtime for
@@
 using Reduction Rules without changing the output string. To test these rules,
-we looked at the rep-n and gen-n
-test cases again, this time focusing on the clustering algorithm for $\epsilon
-= .2$. We report our results in table \ref{red_t}, where
-we give the ratio on top of the simplified grammar produced by
-the clustering algorithm, and the time taken to run just the reduction rules.
+we looked at the rep-n and gen-n test cases again, this time focusing on the
+clustering algorithm for $\epsilon = .2$. We report our results in table
+\ref{red_t}, where we give the ratio on top of the simplified grammar produced
+by the clustering algorithm, and the time taken to run just the reduction
+rules.
 
@@
 
-Because we were unable to get our implementation of
-reduction rules to scale to larger datasets, we
-will not consider them further.
+Because we were unable to get our implementation of reduction rules to scale to
+larger datasets, we will not consider them further.
 
@@
 much of the structure of a text was preserved, we have included some final
-output strings obtained after running the \Cluster algorithm with $\epsilon
-= .4$ in the Appendices. We believe these tests serve as a good (subjective)
+output strings obtained after running the \Cluster algorithm with $\epsilon =
+.4$ in the Appendices. We believe these tests serve as a good (subjective)
 summary of the strengths and weaknesses of our algorithm in compressing data
@@
 that is hard to read. However, key words are clearly recognizable, especially
-important words such as `God', and the overall division of the book into
-verses was retained.
+important words such as `God', and the overall division of the book into verses
+was retained.
 
@@
 1973. The intricate whitespacing was lost, but the clear delineation into
-sections was kept, key words such as `define' and `for' are very visible,
-and the texture of the obfuscated C is very obvious.
+sections was kept, key words such as `define' and `for' are very visible, and
+the texture of the obfuscated C is very obvious.
 
@@
 
-\subsection*{deckmyn.c}
+\appendix
+
+\section{Large Examples}
+
+\subsection*{deckmyn.c, Cluster, $\epsilon=.4$:}
 
@@
 
+\pagebreak
+
 \subsection*{Output:}
@@
 
-\pagebreak
-
-\subsection*{gen256000 Output String, Cluster, $\epsilon=.4$:}
+\vspace{0.2in}
+\subsection*{gen256000, Cluster, $\epsilon=.4$:}
 

### 2012-12-12 tex/paper.tex
slight additions to implementation section

@@
 In addition to this optimization, we also use a native C implementation of the
-Levenshtein function.  We do not compare this to any Ruby implementation, but
-we suspect it has some nontrivial impact on runtime due to the volume of calls
-to the function.
+Levenshtein function. In early versions of our program,
+since every pair of variable expansions must be compared,
+our Ruby implementation of a longest-common-subsequence algorithm was by far the
+largest bottleneck. Replacing it with a native C implementation of levenshtein
+distance improved performance dramatically. Though both implementations have the
+same $O(n^2)$ asymptotic runtime, the constants in the native implementation
+are so small that calls to levenshtein take nearly negligible time on the data
+sizes we were able to test.
 

### 2012-12-12 tex/paper.tex
Moar table fixin'

@@
       \If{$lsim(str_{1},str_{2})<\epsilon$}
-        \State replace the longer var with the smaller throughout $g$
+        \State replace the larger var with the smaller throughout $g$
         \State break
@@
 go through comparing all pairs again after the change.  An alternative
-algorithm would be to try to replace similar variables in parallel. This would
+algorithm would be to try to replace as many similar variables as possible
+all in one sweep. This would
 avoid many of the repeat pair-comparisons done in \Similarity. In other
@@
 had $R^2$ values of over $0.98$.
-
-\begin{tabular}{l|l|l|l|l}
-rep-n runtime & Sim: $\epsilon=.4$ & Sim: $\epsilon=.2$ & Clust: $\epsilon = .4$ & Clust: $\epsilon = .2$ \\
-Exponent      & $1.859$ & $2.026$ & $1.306$ & $1.305$
-\end{tabular}
-
-\begin{tabular}{l|l|l|l|l}
-gen-n runtime & Sim: $\epsilon=.4$ & Sim: $\epsilon=.2$ & Clust: $\epsilon = .4$ & Clust: $\epsilon = .2$ \\
-Exponent      & $2.629$ & $2.493$ & $1.814$ & $2.061$
-\end{tabular}
+\begin{align*}
+  \text{rep-n:}\quad
+  \begin{array}{l|l|l||l|l}
+    & \multicolumn{2}{c||}{\Similarity} & \multicolumn{2}{c}{\Cluster} \\ \hline
+    \epsilon & .4 & .2 & .4 & .2 \\
+    \alpha   & 1.859 & 2.026 & 1.306 & 1.305
+  \end{array}
+  &&
+  \text{gen-n:}\quad
+  \begin{array}{l|l|l||l|l}
+    & \multicolumn{2}{c||}{\Similarity} & \multicolumn{2}{c}{\Cluster} \\ \hline
+    \epsilon & .4 & .2 & .4 & .2 \\
+    \alpha   & 2.629 & 2.493 & 1.814 & 2.061
+  \end{array}
+\end{align*}
 
@@
 Obfuscated C code contest, at \url{http://www.ioccc.org/2012/deckmyn/deckmyn.c}
-.  The size of the lossified grammar is 1576 compared with the original size of
+. We believe this is a particularly good example since its structure is foreign,
+but still discernable. Whitespace is particularly important in standard english
+but our lossifiers don't treat it any differently from any other symbol.
+The size of the lossified grammar is 1576 compared with the original size of
 1973. The intricate whitespacing was lost, but the clear delineation into
 sections was kept, key words such as `define' and `for' are very visible, and
-the texture of the obfuscated C is very obvious.
+the texture of the obfuscated C is very obvious. For instance, after the
+distinct `#define' headers, the first sections has a bunch of control statements
+like `for' and `if', while the middle sections have very generous spacing
+and many repeated pointer derence operators such as `***'. Both of these
+kinds of textures are retained after lossifying.
 
@@
 \end{table}
-
-\appendix
 
@@
 
-\subsection*{Output:}
+\subsubsection*{Output:}
 

### 2012-12-12 tex/paper.tex
Concussion

@@
 
-\subsection{Summary of Results}
-
-* General concept of simplifying grammars, two proposed algorithms
-* reasonably efficient implementation with runtime better than expected
-* compared algorithms, evaluated effectiveness of reducer
-* looked at effect of epsilon
-
-\subsection{Ratios and Fidelity}
-
-* How useful are these algorithms??
-* What is the ratio we get
-* What kind of fidelity to we get
+In this paper, we present the design and implementation of original algorithms
+for performing lossy compression of hierarchical data.  To our knowledge, the
+idea of lossifying context-free grammar as a means of simplifying and
+compressing hierarchical structure has not been explored in the past.  We
+believe, however, that it is a technique worth further investigation.
+
+We were able to produce a reasonably efficient implementation of two basic
+grammar lossification algorithms, and in particular, the runtime for our
+\Cluster was better than we expected.  Our techniques resulted in good
+compression ratios.  The fidelity of the output strings from our compressed
+grammars, as compared to the original inputs, were lacking---however, our
+techniques were intended not to maintain the fidelity of outputs, but rather,
+of hierarchical structure.
+
+Unfortunately, we were unable to establish numerical metrics for the structural
+similarities between the lossless and lossy grammars; however, subjectively, we
+find that lossification maintains interesting invariants, such as the word
+``God'' in a lossy compression of the Bible, or the texture and block structure
+of an obfuscated C program.
 

### 2012-12-12 tex/paper.tex
minor changes to larger tests

--- 2012-12-12 tex/paper.tex
@@
 had $R^2$ values of over $0.98$.
-\begin{align*}
-  \text{rep-n:}\quad
-  \begin{array}{l|l|l||l|l}
-    & \multicolumn{2}{c||}{\Similarity} & \multicolumn{2}{c}{\Cluster} \\ \hline
-    \epsilon & .4 & .2 & .4 & .2 \\
-    \alpha   & 1.859 & 2.026 & 1.306 & 1.305
-  \end{array}
-  &&
-  \text{gen-n:}\quad
-  \begin{array}{l|l|l||l|l}
-    & \multicolumn{2}{c||}{\Similarity} & \multicolumn{2}{c}{\Cluster} \\ \hline
-    \epsilon & .4 & .2 & .4 & .2 \\
-    \alpha   & 2.629 & 2.493 & 1.814 & 2.061
-  \end{array}
-\end{align*}
+
+\begin{tabular}{l|l|l|l|l}
+rep-n runtime & Sim: $\epsilon=.4$ & Sim: $\epsilon=.2$ & Clust: $\epsilon = .4$ & Clust: $\epsilon = .2$ \\
+Exponent      & $1.859$ & $2.026$ & $1.306$ & $1.305$
+\end{tabular}
+
+\begin{tabular}{l|l|l|l|l}
+gen-n runtime & Sim: $\epsilon=.4$ & Sim: $\epsilon=.2$ & Clust: $\epsilon = .4$ & Clust: $\epsilon = .2$ \\
+Exponent      & $2.629$ & $2.493$ & $1.814$ & $2.061$
+\end{tabular}
 
@@
 \end{table}
+
+\appendix
 
@@
 
-\subsubsection*{Output:}
+\subsection*{Output:}
 

### 2012-12-12 tex/paper.tex
Bugfix

--- 2012-12-12 tex/paper.tex
@@
 Obfuscated C code contest, at \url{http://www.ioccc.org/2012/deckmyn/deckmyn.c}
-. We believe this is a particularly good example since its structure is foreign,
-but still discernable. Whitespace is particularly important in standard english
-but our lossifiers don't treat it any differently from any other symbol.
-The size of the lossified grammar is 1576 compared with the original size of
-1973. The intricate whitespacing was lost, but the clear delineation into
-sections was kept, key words such as `define' and `for' are very visible, and
-the texture of the obfuscated C is very obvious. For instance, after the
-distinct `#define' headers, the first sections has a bunch of control statements
-like `for' and `if', while the middle sections have very generous spacing
-and many repeated pointer derence operators such as `***'. Both of these
-kinds of textures are retained after lossifying.
+. We believe this is a particularly good example since its structure is
+foreign, but still discernable. Whitespace is particularly important in
+standard english but our lossifiers don't treat it any differently from any
+other symbol.  The size of the lossified grammar is 1576 compared with the
+original size of 1973. The intricate whitespacing was lost, but the clear
+delineation into sections was kept, key words such as `define' and `for' are
+very visible, and the texture of the obfuscated C is very obvious. For
+instance, after the distinct `\#define' headers, the first sections has a bunch
+of control statements like `for' and `if', while the middle sections have very
+generous spacing and many repeated pointer derence operators such as `***'.
+Both of these kinds of textures are retained after lossifying.
 

### 2012-12-12 tex/paper.tex
Typos

@@
 for performing lossy compression of hierarchical data.  To our knowledge, the
-idea of lossifying context-free grammar as a means of simplifying and
+idea of lossifying context-free grammars as a means of simplifying and
 compressing hierarchical structure has not been explored in the past.  We
@@
 \Cluster was better than we expected.  Our techniques resulted in good
-compression ratios.  The fidelity of the output strings from our compressed
+compression ratios.  The fidelity of the output streams from our compressed
 grammars, as compared to the original inputs, were lacking---however, our
@@
 
-Unfortunately, we were unable to establish numerical metrics for the structural
+We were unfortunately unable to establish numerical metrics for the structural
 similarities between the lossless and lossy grammars; however, subjectively, we

### 2012-12-12 tex/paper.tex
typos

@@
 
-It may seem strange at first to allow a string such as "mind" to be lossily
+It may seem strange at first to allow a string such as ``mind'' to be lossily
 corrupted into, say, ``rind''.  If, however, one is more interested in
@@
 compression: wavelet-based and fractal-based schemes \cite{tutorial}.  In
-wavelet-based schemas, a signal is compressed by recording the largest
+wavelet-based schemes, a signal is compressed by recording the largest
 coefficients of its wavelet decomposition, in effect trying to represent a
@@
 \end{algorithmic}
+\label{s_alg}
 \end{algorithm}
@@
 avoid many of the repeat pair-comparisons done in \Similarity. In other
-words, the algorithm below puts similar variables into equivalence class
+words, the \Cluster puts similar variables into equivalence class
 clusters, and then replaces each variable in a cluster them with the smallest
-representative all at once each iteration.
+representative all at once each iteration. More detailed pseudocode is given
+in Algorithm \ref{c_alg} below.
 
@@
 \end{algorithmic}
+\label{c_alg}
 \end{algorithm}
 
-Both of the algorithms depend on the concept of the "smallest / smaller"
+Both of the algorithms depend on the concept of the ``smallest / smaller''
 variable. We can define the size of a variable to be the length of the
@@
 whenever $V_1$ contains $V_2$ in its rhs. When sizes are computed on a CFG
-before we modify it, if $V_1 \rightarrow V_2$ then $size(V_1) > size(V_2)$
+before we modify it, if $V_1 \hookrightarrow V_2$ is an edge then $size(V_1) > size(V_2)$
 since the rhs of $V_1$ contains $V_2$.  Replacing a variable $V_1$ with $V_2$
@@
 calculated before we did any replacements. This is because each time we
-redirect an edge from $X\rightarrow A$ to $X\rightarrow B$
+redirect an edge from $X\hookrightarrow A$ to $X\hookrightarrow B$
 $size(X)>size(A)>size(B)$ and transitivity will preserve the ordering.  Suppose
 then that we created a cycle during \Cluster by redirecting an edge from $X
-\rightarrow A$ to $X \rightarrow B$.  This means that there must have been a
+\hookrightarrow A$ to $X \hookrightarrow B$.  This means that there must have been a
 path from $B$ to $X$. However, even in its intermediate state we know that the
@@
 could grow much slower than $r$, and the use of native code in calculating the
-Levenshtein distance made the its contribution much less noticeable for the
+Levenshtein distance made its contribution much less noticeable for the
 input sizes we tested.
@@
 The lossifier algorithm reduced the string to a pair of pairs of the same
-equations, where each equation consists of the sum of triples of "12" or "32".
+equations, where each equation consists of the sum of triples of ``12'' or ``32''.
 Smaller discrepencies in the numbers involved were dropped, while the filler
@@
 
-In table \ref{simclus_r_t} we report the results on running \Sequitur and
+In table \ref{simclus_r_t} in the appendix we report the results on running \Sequitur and
 \Cluster, without the grammar reductions, on the rep-n datasets. In table
-\ref{simclus_g_t} we do the same for the gen-n datasets.  In these tests we
+\ref{simclus_g_t} in the appendix we do the same for the gen-n datasets.  In these tests we
 were looking to evaluate the relative fidelity, compression ratios, and runtime
@@
 compared with \Sequitur, where the size of the grammar is the sum of the
-lengths of the rhs's of the rules. Levdistance is the Levenshtein distance
+lengths of the rhs's of the rules. \emph{Lev.} is the Levenshtein distance
 between the output of the lossified grammar and the original input string, it
 is a kind of measure of the fidelity of our algorithms but doesn't take into
-account how well patterns were preserved. For the rep-n tests, proportion is
+account how well patterns were preserved. For the rep-n tests, \emph{Prop.} is
 the proportion of the final string produced that was `*'. The closer to
-1/30~0.033, the more accurate the lossifier was in recreating the original
+$1/30 \approx 0.033$, the more accurate the lossifier was in recreating the original
 texture of the input.
@@
 algorithms for the same values of $\epsilon$.  From plotting the data, we see
-that runtime of the algorithms are roughly polynomial in the size of the data,
+that the runtimes of the algorithms are roughly polynomial in the size of the data,
 so we can perform a linear regression on the log-log transform of the data to
@@
 \Similarity and \Cluster coming in at nearly half the size of the \Sequitur CFG
-for the rep-n data, and 70 percent for the gen-n data. A notable exception was
+for the rep-n data, and 70 percent for the gen-n data. This is promising because
+in \cite{sequitur}, the authors show that \Sequitur on its own achieves better
+ratios than most standard compression algorithms.
+A notable exception was
 the gen-n test with $\epsilon=.2$, where $\epsilon$ was so small that barely
@@
 clustering algorithm for $\epsilon = .2$. We report our results in table
-\ref{red_t}, where we give the ratio on top of the simplified grammar produced
+\ref{red_t} in the appendix, where we give the ratio on top of the simplified grammar produced
 by the clustering algorithm, and the time taken to run just the reduction
@@
 that are possible on diverse datasets don't lead to the wide-ranging repetitive
-substitutions which allow reduction rules to be allowed.
+substitutions which allow reduction rules to be applied.
 
@@
 very visible, and the texture of the obfuscated C is very obvious. For
-instance, after the distinct `\#define' headers, the first sections has a bunch
+instance, after the distinct `\#define' headers, the first section has a bunch
 of control statements like `for' and `if', while the middle sections have very
@@
 
-1:9 And God said, Let the waters under the heaven be gathered together
-unto one place, and let the dry land appear: and it was so.
-
-1:10 And God called the dry land Earth; and the gathering together of
-the waters called he Seas: and God saw that it was good.
-
 (Rest Omitted)
@@
 
+ine id, Lt e frargras, e ding
+seedn e rut dingrut fes inwhseseed i
+in tel, e frt wa
+
+n e frrurhtargras, n dingseedfes inn e dingrut , whseseeds in tel, fe
+s inine tht waod3e e frinwee e thir
+
+1e said, e behts ine mamene e 
+todividthayfrm e htn e e e frsignd an
+seasond anayn ar: 1:n e e e frhts in
+e frmamene e  toinhte fr: n t wa
+
+6ine  tcrathts;e grHatoule thay
+nde estoule e hte e e e stars al
+
+7ine t e ine mamene e toinht
+e fr1:n toule e e ayn e e htd an
+todivide htfrm darknes: God tht waod9n e e n frinwee e fr
+
+0e said, binar aundanolye moilngcaledthhath in frwlthay flyabene frentoopen firmamenofmae
+
+1 n God Haedcratwhalesn enry lcalede
+
 (Rest Omitted)
